---
title: "Notes 1"
author: "Jefferson Ong"
date: "1/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Use number in naming if can 01, 02, 03,..

0-download.R
1-parse.R
2-explore.R

Use _ in naming

# Good
day_one
day_1

# Bad
first_day_of_the_month
DayOne
dayone
djm1


Don't use T or F

```{r}
# Bad
T <- FALSE
c <- 10
mean <- function(x) sum(x)
```

Space properly

# Good
average <- mean(feet / 12 + inches, na.rm = TRUE)

# Bad
average<-mean(feet/12+inches,na.rm=TRUE)


---

() argument for the function

[] location of vector, [i] one dimension, [i,j] two dimensions

{} multiple lines of codes

```{r}
long_function_name <- function(a = "a long argument", 
                               b = "another argument",
                               c = "another long argument") {
  # As usual code is indented by two spaces.
}
```

----

<- makes object

= set parameters inside functions

-------------

```{r}
answer <- 2 + 3
```


The answer is `r answer`.

```{r}
random_num <- rnorm(1, 0, 1)

```

The random number is `r random_num`

The second random number is `r rnorm(1, 0, 1)`














