---

title: "Challenge #1 Examples"
author: "Author: Jill Thomley"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document:
    theme: readable
    highlight: textmate
---

***

#### Instructions

Using the `flights` and `airlines` datasets in `nycflights13`...

* Create a barplot of airports from which flights originated.
* Create a barplot of the airlines associated with the flights.
* For each plot, add a title, better labels on the axes, and a caption source: nycflights13.
* Can you add color or other adjustments to the plots to make them more attractive?
* Can you change the labels on the individual bars from their default variable values?
* Can you left_join flights and airlines so you can use airline names instead of codes?
* Can you flip the plot so that the bars are displayed horizontally rather than vertically?


#### Set `R` code chunk options and load packages

```{r global_setup}
# Set options for code chunks
# Suppress non-warning messages 
# Remove default ## from output
knitr::opts_chunk$set(message = FALSE, comment = "")
```

```{r load_packages}
library(nycflights13)
library(ggplot2)
library(dplyr)
```


#### Create a basic barplot for the `origin` of flights

```{r origin_barplot_basic}
ggplot(flights, aes(x = origin)) +
  geom_bar()
```


#### Format the plot to make it more attractive/useful

```{r origin_barplot_formatted}
ggplot(flights, aes(x = origin)) +
  geom_bar(fill = "lightgreen", color = "darkgray", width = -1) +
  labs(title = "Number of flights departing NYC metropolitan airports in 2013",
       x = "airport of origin",
       y = "number of flights",
       caption = "source: nycflights13") +
  scale_x_discrete(labels = c("Newark", "JFK", "LaGuardia")) +
  scale_y_continuous(labels = scales::comma)
```


#### Create a basic barplot for the `carrier` of flights

```{r carrier_barplot_basic}
ggplot(flights, aes(x = carrier)) +
  geom_bar()
```


#### Format the plot to make it more attractive/useful

```{r carrier_barplot_formatted}
ggplot(flights, aes(x = carrier)) +
  geom_bar(fill = "lightblue", color = "darkgray", width = .75) +
  labs(title = "Number of flights per carrier departing NYC metro airports in 2013",
       x = "airline code",
       y = "number of flights",
       caption = "source: nycflights13") +
  scale_y_continuous(labels = scales::comma)
```


#### Merge `flights` and `airlines` so we can use carrier names

```{r lj_flights_airlines}
flightsM <- left_join(flights, airlines, key = carrier)
```


#### Recreate the barplot using `name` instead of `carrier`

```{r airname_barplot_formatted}
ggplot(flightsM, aes(x = name)) +
  geom_bar(fill = "lightblue", color = "darkgray", width = .75) +
  labs(title = "Flights per carrier departing NYC metro airports in 2013",
       x = "airline name",
       y = "number of flights",
       caption = "source: nycflights13") +
  scale_y_continuous(labels = scales::comma) +
  coord_flip()
```


#### To arrange bar order, we need to use summarized data

```{r count_airname}
carrierCounts <- flightsM %>% 
  group_by(name) %>% 
  summarize(freq = n())
carrierCounts
```

```{r sorted_barplot_formatted}
ggplot(carrierCounts, aes(x = reorder(name, -freq), y = freq)) +
  geom_bar(stat = "identity", fill = "violet", color = "black", width = .75) +
  labs(title = "Flights per carrier departing NYC metro airports in 2013",
       x = "airline name",
       y = "number of flights",
       caption = "source: nycflights13") +
  scale_y_continuous(labels = scales::comma) +
  coord_flip()
```


***

```{r}
sessionInfo()
```