---
title: "Chapter 2 Homework Analysis 2"
author: "Jefferson Ong"
date: "1/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ISLR)
library(tidyverse)
college <- College
auto <- Auto
```

Please complete problems 8 and 9 from chapter 2 on pp 54-56 of the text.  Due date is Friday, Jan 31st.  Please write up your answers using R Markdown.  For this assignment, please e-mail me your final submission.

You will be allowed to work in groups of two if you prefer to do that for this homework.

If you have any questions please let me know.

Ross Gosky


# 8 

(a) Use the read.csv() function to read the data into R. Call the
loaded data college. Make sure that you have the directory set
to the correct location for the data.

(b) Look at the data using the fix() function. You should notice
that the first column is just the name of each university. We don’t
really want R to treat this as data. However, it may be handy to
have these names for later. Try the following commands:

2.4 Exercises 55

```{r}
rownames(college)=college[,1]
fix(college)
```



You should see that there is now a row.names column with the
name of each university recorded. This means that R has given
each row a name corresponding to the appropriate university. R
will not try to perform calculations on the row names. However,
we still need to eliminate the first column in the data where the
names are stored. Try



```{r}
college =college[,-1]
fix(college)
```

Now you should see that the first data column is Private. Note
that another column labeled row.names now appears before the
Private column. However, this is not a data column but rather
the name that R is giving to each row.

(c) i. Use the summary() function to produce a numerical summary
of the variables in the data set.

ii. Use the pairs() function to produce a scatterplot matrix of
the first ten columns or variables of the data. Recall that
you can reference the first ten columns of a matrix A using
A[,1:10].

iii. Use the plot() function to produce side-by-side boxplots of
Outstate versus Private.

iv. Create a new qualitative variable, called Elite, by binning
the Top10perc variable. We are going to divide universities
into two groups based on whether or not the proportion
of students coming from the top 10 % of their high school
classes exceeds 50 %.

> Elite=rep("No",nrow(college ))
> Elite[college$Top10perc >50]=" Yes"
> Elite=as.factor(Elite)
> college=data.frame(college , Elite)
Use the summary() function to see how many elite universities there are. Now use the plot() function to produce
side-by-side boxplots of Outstate versus Elite.
v. Use the hist() function to produce some histograms with
differing numbers of bins for a few of the quantitative variables. You may find the command par(mfrow=c(2,2)) useful:
it will divide the print window into four regions so that four
plots can be made simultaneously. Modifying the arguments
to this function will divide the screen in other ways.
vi. Continue exploring the data, and provide a brief summary
of what you discover.













































