---
title: "CleaningDatainR"
author: "Jefferson Ong"
date: "4/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(infer)
library(readr)
library(resampledata)

bmi <- read.csv("https://assets.datacamp.com/production/repositories/34/datasets/a0a569ebbb34500d11979eba95360125127e6434/bmi_clean.csv")

census <- read.csv("https://assets.datacamp.com/production/repositories/34/datasets/f82ab0a3ccb95fe40e18c6eac5644d288cd126ea/census-retail.csv")

students <- read.csv("https://assets.datacamp.com/production/repositories/34/datasets/f75a87dbbdf2cf79e2286f97b2af22146cb717b1/students_with_dates.csv")

# weather <- readRDS("weather.RDS")

```


## Introduction and exploring raw data

1. Exploring raw data
2. Tidying data
3. Preparing data for analysis

```{r}
# View the first 6 rows of data
# head(weather)
# 
# # View the last 6 rows of data
# tail(weather)
# 
# # View a condensed summary of the data
# str(weather)


# Check the structure of bmi, the dplyr way
glimpse(bmi)

# View a summary of bmi
summary(bmi)

# Print bmi to the console
bmi

# View the first 6 rows
head(bmi)

# View the first 15 rows
head(bmi, 15)

# View the last 6 rows
tail(bmi)

# View the last 10 rows
tail(bmi, 10)
```




● Understand the structure of your data
● Look at your data
● Visualize your data





Understanding the structure of your data
● class() - Class of data object
● dim() - Dimensions of data
● names() - Column names
● str() - Preview of data with helpful details
● glimpse() - Be!er version of str() from dplyr
● summary() - Summary of data


Looking at your data
● head() - View top of dataset
● tail() - View bo!om of dataset
● print() - View entire dataset (not recommended!)

```{r}
# Histogram of BMIs from 2008
hist(bmi$Y2008)

# Scatter plot comparing BMIs from 1980 to those from 2008
plot(bmi$Y1980, bmi$Y2008)
```


Visualizing your data
● hist() - View histogram of a single variable
● plot() - View plot of two variables




## Tidying data

Summary of key tidyr functions
● gather() - Gather columns into key-value pairs
● spread() - Spread key-value pairs into columns
● separate() - Separate one column into multiple
● unite() - Unite multiple columns into one

```{r}
# Apply gather() to bmi and save the result as bmi_long
bmi_long <- gather(bmi, year, bmi_val, -Country)

# View the first 20 rows of the result
head(bmi_long, 20)

# Apply spread() to bmi_long
bmi_wide <- spread(bmi_long, year, bmi_val)

# View the head of bmi_wide
head(bmi_wide)
```

```{r}
# # Apply separate() to bmi_cc
# bmi_cc_clean <- separate(bmi_cc, col = Country_ISO, into = c("Country", "ISO"), sep = # "/")
# 
# # View the head of the result
# head(bmi_cc_clean)
# 
# # Apply unite() to bmi_cc_clean
# bmi_cc <- unite(bmi_cc_clean, Country_ISO, Country, ISO, sep = "-")
# 
# # View the head of the result
# head(bmi_cc)
```

```{r}
# View the head of census
head(census)

# Gather the month columns
census2 <- gather(census, month, amount, -YEAR)

# Arrange rows by YEAR using dplyr's arrange
census2_arr <- arrange(census2, YEAR)

# # View first 20 rows of census2
# head(census2_arr, 20)
# 
# # View first 50 rows of census_long
# head(census_long, 50)
# 
# # Spread the type column
# census_long2 <- spread(census_long, type, amount)
# 
# # View first 20 rows of census_long2
# head(census_long2, 20)
# 
# # View the head of census_long3
# head(census_long3)
# 
# # Separate the yr_month column into two
# census_long4 <- separate(census_long3, yr_month, c("year", "month"))
# 
# # View the first 6 rows of the result
# head(census_long4)
```









## Preparing data for analysis

Types of variables in R
● character: "treatment", "123", "A"
● numeric: 23.44, 120, NaN, Inf
● integer: 4L, 1123L
● factor: factor("Hello"), factor(8)
● logical: TRUE, FALSE, NA


```{r}
# Make this evaluate to "character"
class("TRUE")

# Make this evaluate to "numeric"
class(8484.00)

# Make this evaluate to "integer"
class(99L)

# Make this evaluate to "factor"
class(factor("factor"))

# Make this evaluate to "logical"
class(FALSE)
```


Key functions in stringr for cleaning data
● str_trim() - Trim leading and trailing white space
● str_pad() - Pad with additional characters
● str_detect() - Detect a pa!ern
● str_replace() - Find and replace a pa!ern

```{r}
# Preview students with str()
str(students)

# Coerce Grades to character
students$Grades <- as.character(students$Grades)

# Coerce Medu to factor
students$Medu <- as.factor(students$Medu)

# Coerce Fedu to factor
students$Fedu <- as.factor(students$Fedu)

# Look at students once more with str()
str(students)
```



● tolower() - Make all lowercase
● toupper() - Make all uppercase

```{r}
# Preview students2 with str()
# str(students2)
# 
# # Load the lubridate package
# library(lubridate)
# 
# # Parse as date
# dmy("17 Sep 2015")
# 
# # Parse as date and time (with no seconds!)
# mdy_hm("July 15, 2012 12:56")
# 
# # Coerce dob to a date (with no time)
# students2$dob <- ymd(students2$dob)
# 
# # Coerce nurse_visit to a date and time
# students2$nurse_visit <- ymd_hms(students2$nurse_visit)
# 
# # Look at students2 once more with str()
# str(students2)
```


Missing values
● May be random, but dangerous to assume
● Sometimes associated with variable/outcome of interest
● In R, represented as NA
● May appear in other forms
● #N/A (Excel)
● Single dot (SPSS, SAS)
● Empty string

```{r}
# Load the stringr package
library(stringr)

# Trim all leading and trailing whitespace
str_trim(c("   Filip ", "Nick  ", " Jonathan"))

# Pad these strings with leading zeros
str_pad(c("23485W", "8823453Q", "994Z"), width = 9, side = "left", pad = "0")
```

```{r}
# # Print state abbreviations
# states
# 
# # Make states all uppercase and save result to states_upper
# states_upper <- toupper(states)
# 
# # Make states_upper all lowercase again
# tolower(states_upper)
```

```{r}
# # Copy of students2: students3
# students3 <- students2
# 
# # Look at the head of students3
# head(students3)
# 
# # Detect all dates of birth (dob) in 1997
# str_detect(students3$dob, "1997")
# 
# # In the sex column, replace "F" with "Female" ...
# students3$sex <- str_replace(students3$sex, "F", "Female")
# 
# # ... and "M" with "Male"
# students3$sex <- str_replace(students3$sex, "M", "Male")
# 
# # View the head of students3
# head(students3)
```



Special values
● Inf - "Infinite value" (indicative of outliers?)
● 1/0
● 1/0 + 1/0
● 33333^33333
● NaN - "Not a number" (rethink a variable?)
● 0/0
● 1/0 - 1/0


Outliers
● Extreme values distant from other values
● Several causes
● Valid measurements
● Variability in measurement
● Experimental error
● Data entry error
● May be discarded or retained depending on cause


```{r}
# # Call is.na() on the full social_df to spot all NAs
# is.na(social_df)
# 
# # Use the any() function to ask whether there are any NAs in the data
# any(is.na(social_df))
# 
# # View a summary() of the dataset
# summary(social_df)
# 
# # Call table() on the status column
# table(social_df$status)
```

```{r}
# # Replace all empty strings in status with NA
# social_df$status[social_df$status == ""] <- NA
# 
# # Print social_df to the console
# social_df
# 
# # Use complete.cases() to see which rows have no missing values
# complete.cases(social_df)
# 
# # Use na.omit() to remove all rows with any missing values
# na.omit(social_df)
```

```{r}
# # Look at a summary() of students3
# summary(students3)
# 
# # View a histogram of the age variable
# hist(students3$age)
# 
# # View a histogram of the absences variable
# hist(students3$absences)
# 
# # View a histogram of absences, but force zeros to be bucketed to the right of zero
# hist(students3$absences, right = FALSE)
# 
# # View a boxplot of age
# boxplot(students3$age)
# 
# # View a boxplot of absences
# boxplot(students3$absences)
```







## Putting it all together




Understanding the structure of your data
● class() - Class of data object
● dim() - Dimensions of data
● names() - Column names
● str() - Preview of data with helpful details
● glimpse() - Be!er version of str() from dplyr
● summary() - Summary of data

● head() - View top of dataset
● tail() - View bo!om of dataset
● print() - View entire dataset (not recommended!)

hist() - View histogram of a single variable
● plot() - View plot of two variables

```{r}
# Load the tidyr package
# library(tidyr)
# 
# # Gather the columns
# weather2 <- gather(weather, day, value, X1:X31, na.rm = TRUE)
# 
# # View the head
# head(weather2)

# # First remove column of row names
# without_x <- weather2[, -1]
# 
# # Spread the data
# weather3 <- spread(without_x, measure, value)
# 
# # View the head
# head(weather3)
```

```{r}
# Load the stringr and lubridate packages
# library(stringr)
# library(lubridate)
# 
# # Remove X's from day column
# weather3$day <- str_replace(weather3$day, "X", "")
# 
# # Unite the year, month, and day columns
# weather4 <- unite(weather3, date, year, month, day, sep = "-")
# 
# # Convert date column to proper date format using lubridates's ymd()
# weather4$date <- ymd(weather4$date)
# 
# # Rearrange columns using dplyr's select()
# weather5 <- select(weather4, date, Events, CloudCover:WindDirDegrees)
# 
# # View the head of weather5
# head(weather5)
```

```{r}
# # View the structure of weather5
# str(weather5)
# 
# # Examine the first 20 rows of weather5. Are most of the characters numeric?
# head(weather5, 20)
# 
# # See what happens if we try to convert PrecipitationIn to numeric
# as.numeric(weather5$PrecipitationIn)
```

```{r}
# # Replace "T" with "0" (T = trace)
# weather5$PrecipitationIn <- str_replace(weather5$PrecipitationIn, "T", "0")
# 
# # Convert characters to numerics
# weather6 <- mutate_at(weather5, vars(CloudCover:WindDirDegrees), funs(as.numeric))
# 
# # Look at result
# str(weather6)
```

```{r}
# # Count missing values
# sum(is.na(weather6))
# 
# # Find missing values
# summary(weather6)
# 
# # Find indices of NAs in Max.Gust.SpeedMPH
# ind <- which(is.na(weather6$Max.Gust.SpeedMPH))
# 
# # Look at the full rows for records missing Max.Gust.SpeedMPH
# weather6[ind, ]
#  head(weather)
```


```{r}
# # Review distributions for all variables
# summary(weather6)
# 
# # Find row with Max.Humidity of 1000
# ind <- which(weather6$Max.Humidity == 1000)
# 
# # Look at the data for that day
# weather6[ind, ]
# head(weather)
# 
# # Change 1000 to 100
# weather6$Max.Humidity[ind] <- 100
# weather6 %>%
#   mutate(Max.Humidity == 100)
```


```{r}
# # Look at summary of Mean.VisibilityMiles
# summary(weather6$Mean.VisibilityMiles)
# 
# # Get index of row with -1 value
# ind <- which(weather6$Mean.VisibilityMiles == -1)
# 
# # Look at full row
# weather6[ind, ]
# 
# # Set Mean.VisibilityMiles to the appropriate value
# weather6$Mean.VisibilityMiles[ind] <- 10
```

```{r}
# # Review summary of full data once more
# summary(weather6)
# 
# # Look at histogram for MeanDew.PointF
# hist(weather6$MeanDew.PointF)
# 
# # Look at histogram for Min.TemperatureF
# hist(weather6$Min.TemperatureF)
# 
# # Compare to histogram for Mean.TemperatureF
# hist(weather6$Mean.TemperatureF)
```


```{r}
# # Clean up column names
# names(weather6) <- new_colnames
# 
# # Replace empty cells in events column
# weather6$events[weather6$events == ""] <- "None"
# 
# # Print the first 6 rows of weather6
# head(weather6)
# glimpse(weather6)
# summary(weather6)
```













Summary of your accomplishments
● Inspected the data
● Tidied the data
● Improved date representations
● Dealt with incorrect variable codings
● Found and dealt with missing data
● Identified and corrected errors
● Visualized the result




















