---
title: "Data Visualization with ggplot2 (Part 3)"
author: "Jefferson Ong"
date: "3/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(lattice)
library(readr)
```


```{r}
UScities <- read_csv("https://assets.datacamp.com/production/repositories/414/datasets/24739149e0dbdbdc84dcbf275b68616cb2481005/US_Cities.txt")

Usstates <- read_csv("https://assets.datacamp.com/production/repositories/414/datasets/7eef36579d107fefbcb38d0c314c963e608c9609/US_States.txt")

Germanyunem <- read_csv("https://assets.datacamp.com/production/repositories/414/datasets/bdedafb52d7060a90f9bf320cf11a274ce02bcfd/germany_unemployment.txt")

JapPop <- read_csv("https://assets.datacamp.com/production/repositories/414/datasets/f2efc9d1f2f07a22843aabef510094c6e5474616/japanPOP.txt")

ParisWea <- read_csv("https://assets.datacamp.com/production/repositories/414/datasets/df61e885cc58b88db51968a13ca7827897b098e8/FRPARIS.txt")

ReyWea <- read_csv("https://assets.datacamp.com/production/repositories/414/datasets/45d984ccc4d2afa2023b7139824116040aac3a54/ILREYKJV.txt")


NewyrkWea <- read.csv("https://assets.datacamp.com/production/repositories/414/datasets/c37fafe15bfa05a338f8c835e79ee5e242400438/NYNEWYOR.txt")

LondonWea <- read_csv("https://assets.datacamp.com/production/repositories/414/datasets/89250a654c2f83331a90e6538f89d501aa966181/UKLONDON.txt")
```

```{r}
mammals <- readRDS("mammals.RDS")

```

# Statistical plots


```{r}
# Create movies_small
library(ggplot2movies)
set.seed(123)
movies_small <- movies[sample(nrow(movies), 1000), ]
movies_small$rating <- factor(round(movies_small$rating))

# Explore movies_small with str()
str(movies_small)

# Build a scatter plot with mean and 95% CI
ggplot(movies_small, aes(x = rating, y = votes)) +
  geom_point() +
  stat_summary(fun.data = "mean_cl_normal",
               geom = "crossbar",
               width = 0.2,
               col = "red") +
  scale_y_log10()

# Reproduce the plot
ggplot(diamonds, aes(x = carat, y = price, col = color)) +
  geom_point(alpha = 0.5, size = 0.5, shape = 16) +
  scale_x_log10(expression(log[10](Carat)), limits = c(0.1,10)) +
  scale_y_log10(expression(log[10](Price)), limits = c(100,100000)) +
  scale_color_brewer(palette = "YlOrRd") +
  coord_equal() +
  theme_classic()

# Reproduce the plot
ggplot(diamonds, aes(x = carat, y = price, col = color)) +
  geom_point(alpha = 0.5, size = 0.5, shape = 16) +
  scale_x_log10(expression(log[10](Carat)), limits = c(0.1,10)) +
  scale_y_log10(expression(log[10](Price)), limits = c(100,100000)) +
  scale_color_brewer(palette = "YlOrRd") +
  coord_equal() +
  theme_classic()
```

```{r}
# movies_small is available

# Add a boxplot geom
d <- ggplot(movies_small, aes(x = rating, y = votes)) +
  geom_point() +
  geom_boxplot() +
  stat_summary(fun.data = "mean_cl_normal",
               geom = "crossbar",
               width = 0.2,
               col = "red")

# Untransformed plot
d

# Transform the scale
d + scale_y_log10()

# Transform the coordinates
# d + coord_trans(y = "log10")




# Plot object p
p <- ggplot(diamonds, aes(x = carat, y = price))

# Use cut_interval
p + geom_boxplot(aes(group = cut_interval(carat, n = 10)))

# Use cut_number
p + geom_boxplot(aes(group = cut_number(carat, n = 10)))

# Use cut_width
p + geom_boxplot(aes(group = cut_width(carat, width = 0.25)))
```



## Box Plots

Box plot
● John Tukey - Exploratory Data Analysis
● Visualizing the 5 number summary


```{r}
# test_data is available

# Calculating density: d
# d <- density(test_data$norm)
# 
# # Use which.max() to calculate mode
# mode <- d$x[which.max(d$y)]
# 
# # Finish the ggplot call
# ggplot(test_data, aes(x = norm)) +
#   geom_rug() +
#   geom_density() +
#  geom_vline(xintercept = mode, col = "red")
```

```{r}
# test_data is available

# Arguments you'll need later on
# fun_args <- list(mean = mean(test_data$norm), sd = sd(test_data$norm))

# Finish the ggplot
# ggplot(test_data, aes(x = norm)) +
#   geom_histogram(aes(y = ..density..)) +
#   geom_density(col = "red") +
#   stat_function(fun = dnorm, args = fun_args, col = "blue")
```

```{r}
# small_data is available

# Get the bandwith
# get_bw <- density(small_data$x)$bw
# 
# # Basic plotting object
# p <- ggplot(small_data, aes(x = x)) +
#   geom_rug() +
#   coord_cartesian(ylim = c(0,0.5))
# 
# # Create three plots
# p + geom_density()
# p + geom_density(adjust = 0.25)
# p + geom_density(bw = 0.25 * get_bw)
# 
# # Create two plots
# p + geom_density(kernel = "r")
# p + geom_density(kernel = "e")
```
```{r}
# Individual densities
ggplot(mammals[mammals$vore == "Insectivore",], aes(x = sleep_total, fill = vore)) +
  geom_density(col = NA, alpha = 0.35) +
  scale_x_continuous(limits = c(0, 24)) +
  coord_cartesian(ylim = c(0, 0.3))

# With faceting
ggplot(mammals, aes(x = sleep_total, fill = vore)) +
  geom_density(col = NA, alpha = 0.35) +
  scale_x_continuous(limits = c(0, 24)) +
  coord_cartesian(ylim = c(0, 0.3)) +
  facet_wrap( ~ vore, nrow = 2)

# By default, the x ranges fill the scale
ggplot(mammals, aes(x = sleep_total, fill = vore)) +
  geom_density(col = NA, alpha = 0.35) +
  scale_x_continuous(limits = c(0, 24)) +
  coord_cartesian(ylim = c(0, 0.3))

# Trim each density plot individually
ggplot(mammals, aes(x = sleep_total, fill = vore)) +
  geom_density(col = NA, alpha = 0.35, trim = TRUE) +
  scale_x_continuous(limits = c(0, 24)) +
  coord_cartesian(ylim = c(0, 0.3))
```


```{r}
# Unweighted density plot from before
ggplot(mammals, aes(x = sleep_total, fill = vore)) +
  geom_density(col = NA, alpha = 0.35) +
  scale_x_continuous(limits = c(0, 24)) +
  coord_cartesian(ylim = c(0, 0.3))

# Unweighted violin plot
ggplot(mammals, aes(x = vore, y = sleep_total, fill = vore)) +
  geom_violin()

# Calculate weighting measure
library(dplyr)
mammals2 <- mammals %>%
  group_by(vore) %>%
  mutate(n = n() / nrow(mammals)) -> mammals

# Weighted density plot
ggplot(mammals2, aes(x = sleep_total, fill = vore)) +
  geom_density(aes(weight = n), col = NA, alpha = 0.35) +
  scale_x_continuous(limits = c(0, 24)) +
  coord_cartesian(ylim = c(0, 0.3))

# Weighted violin plot
ggplot(mammals2, aes(x = vore, y = sleep_total, fill = vore)) +
  geom_violin(aes(weight = n), col = NA)
```

```{r}
# Base layers
p <- ggplot(faithful, aes(x = waiting, y = eruptions)) +
  scale_y_continuous(limits = c(1, 5.5), expand = c(0, 0)) +
  scale_x_continuous(limits = c(40, 100), expand = c(0, 0)) +
  coord_fixed(60 / 4.5)

# 1 - Use geom_density_2d()
p + geom_density_2d()

# 2 - Use stat_density_2d() with arguments
p + stat_density_2d(aes(col = ..level..), h = c(5, 0.5))
```
```{r}
# Load in the viridis package
library(viridis)

# Add viridis color scale
ggplot(faithful, aes(x = waiting, y = eruptions)) +
  scale_y_continuous(limits = c(1, 5.5), expand = c(0,0)) +
  scale_x_continuous(limits = c(40, 100), expand = c(0,0)) +
  coord_fixed(60/4.5) +
  stat_density_2d(geom = "tile", aes(fill = ..density..), h=c(5,.5), contour = FALSE) +
  scale_fill_viridis()
```


## Density plot

 Distribution of univariate data
● Statistics
● Probability Density Function
● Theoretical: based on formula
● Empirical: based on data

## Kernel Density Estimate (KDE)

A sum of 'bumps' placed at the observations.
The kernel function determines the shape of the bumps
while the window width, h, determines their width.


# Plots for specific data types (Part 1)

Defining largeness

● Many observations
● High-resolution time series
● Large surveys
● Website analytics
● Many variables
● Multidimensional data
● Combination

```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
 geom_point(shape = 1)

ggplot(diamonds, aes(x = carat, y = price)) +
 geom_point(shape = ".")

 ggplot(diamonds, aes(x = carat, y = price)) +
 geom_point(shape = 16, alpha = 0.2)
 
  ggplot(diamonds, aes(x = carat, y = price)) +
 geom_point(shape = 16, alpha = 0.2) +
 stat_density2d(col = "blue")
  
   ggplot(diamonds, aes(x = carat, y = price)) +
 geom_point(shape = 16, alpha = 0.2) +
 stat_density2d(aes(colour=..level..))
   
   
    ggplot(diamonds, aes(x = carat, y = price)) +
 stat_density2d(geom = "tile",
 aes(fill = ..density..),
 contour = FALSE)
```
```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
 geom_bin2d(bins = 100)


ggplot(diamonds, aes(x = carat, y = price)) +
 geom_hex()

ggplot(diamonds, aes(x = carat, y = price)) +
 geom_hex(bins = 100)

```
Many observations

● Reducing over-plo!ing
● Reducing amount of information that is plo!ed
● Aggregates


Many variables

● Multi-variate or high-dimensional data
● Data reduction methods
● Previously: facets
● Two plot types


### SPLOM - Sca!er PLOt Matrix

```{r}
pairs(iris[-5])
```

### Correlation matrix

```{r}
library(PerformanceAnalytics)
 chart.Correlation(iris[-5])
```

### ggAlly::ggpairs()

```{r}
library(GGally)
 ggpairs(mtcars)
```


### Parallel coordinate plot

```{r}
ggparcoord(iris, columns = 1:4,
 groupColumn = 5,
 scale = "globalminmax",
 order = "anyClass", alphaLines = 0.4)
```

## Ternary Plots

● Triangle plot
● Compositional trivariate data
● Add up to 100%
● Typical example: soil composition
● Sand
● Silt
● Clay


```{r}
# africa.sample

# library(ggtern)
#  ggtern(africa, aes(Sand, Silt, Clay)) +
#  geom_density_tern()
```

## Network Plots

Blood types
● Four main types
● A, B, AB, O
● Rho negative or positive
● 8 different types
● Not all compatible
● Which blood type can be used as a donor?

```{r}
# ggplot(Blood_donors, aes(x = to, y = from)) +
#  geom_tile()
```

## Diagnostic Plots

● Quantitative data
● Ordinary least squares model
● Assess how the model fits the data

```{r}
 res <- lm(Volume ~ Girth, data = trees)
 plot(Volume ~ Girth, data = trees)
 abline(res, col = "red")
```

```{r}
# pairs
pairs(iris[1:4])

# chart.Correlation
library(PerformanceAnalytics)
chart.Correlation(iris[1:4])

# ggpairs
library(GGally)
# ggpairs(mtcars_fact[1:3])
```
```{r}
library(ggplot2)
library(reshape2)

cor_list <- function(x) {
  L <- M <- cor(x)
  
  M[lower.tri(M, diag = TRUE)] <- NA
  M <- melt(M)
  names(M)[3] <- "points"
  
  L[upper.tri(L, diag = TRUE)] <- NA
  L <- melt(L)
  names(L)[3] <- "labels"
  
  merge(M, L)
}

# Calculate xx with cor_list
library(dplyr)
xx <- iris %>%
  group_by(Species) %>%
  do(cor_list(.[1:4])) 

# Finish the plot
ggplot(xx, aes(x = Var1, y = Var2)) +
  geom_point(aes(col = points, size = abs(points)), shape = 16) +
  geom_text(aes(col = labels,  size = abs(labels), label = round(labels, 2))) +
  scale_size(range = c(0, 6)) +
  scale_color_gradient("r", limits = c(-1, 1)) +
  scale_y_discrete("", limits = rev(levels(xx$Var1))) +
  scale_x_discrete("") +
  guides(size = FALSE) +
  geom_abline(slope = -1, intercept = nlevels(xx$Var1) + 1) +
  coord_fixed() +
  facet_grid(. ~ Species) +
  theme(axis.text.y = element_text(angle = 45, hjust = 1),
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_blank())
```

```{r}
# # Load ggtern
# library(ggtern)
# 
# # Build ternary plot
# ggtern(africa, aes(x = Sand, y = Silt, z = Clay)) +
#   geom_point(shape = 16, alpha = 0.2)
```


```{r}
library(geomnet)
str(madmen)

# Merge edges and vertices
mmnet <- merge(madmen$edges, madmen$vertices,
               by.x = "Name1", by.y = "label",
               all = TRUE)

# Examine structure of mmnet
str(mmnet)
```

```{r}
# geomnet is pre-loaded

# Merge edges and vertices
mmnet <- merge(madmen$edges, madmen$vertices,
               by.x = "Name1", by.y = "label",
               all = TRUE)

# Finish the ggplot command
ggplot(mmnet, aes(from_id = Name1, to_id = Name2)) +
  geom_net(aes(col = Gender),
           size = 6,
           linewidth = 1,
           labelon = TRUE,
           fontsize = 3,
           labelcolour = "black")
```

```{r}
# geomnet is pre-loaded and mmnet is defined
head(mmnet)

# Node colors
pink_and_blue <- c(female = "#FF69B4", male = "#0099ff")

# Tweak the network plot
ggplot(data = mmnet, aes(from_id = Name1, to_id = Name2)) +
  geom_net(aes(col = Gender),
           size = 6,
           linewidth = 1,
           labelon = TRUE,
           fontsize = 3,
           labelcolour = "black",
           # Make the graph directed
           directed = TRUE) +
  # Add manual color scale
  scale_color_manual(values = pink_and_blue) + 
  # Set x-axis limits
  xlim(c(-0.05, 1.05)) +
  # Set void theme
  theme_void()
```


```{r}
# Create linear model: res
res <- lm(Volume ~ Girth, data = trees)

# Plot res
plot(res)

# Import ggfortify and use autoplot()
library(ggfortify)
autoplot(res, ncol = 2)
```

```{r}
# ggfortify and Canada are available

# Inspect structure of Canada
# str(Canada)
# 
# # Call plot() on Canada
# plot(Canada)
# 
# # Call autoplot() on Canada
# autoplot(Canada)

# ggfortify and eurodist are available
# Autoplot + ggplot2 tweaking
autoplot(eurodist) +
  coord_fixed()
  
# Autoplot of MDS
autoplot(cmdscale(eurodist, eig = TRUE),
         label = TRUE, 
         label.size = 3, 
         size = 0)
```





```{r}
# Perform clustering
iris_k <- kmeans(iris[-5], 3)

# Autplot: color according to cluster
autoplot(iris_k, data = iris, frame = TRUE)

# Autoplot: above, plus shape according to species
autoplot(iris_k, data = iris, frame = TRUE, shape = 'Species')
```








# Plots for specific data types (Part 2)


```{r}
# maps, ggplot2, and ggmap are pre-loaded
# Use map_data() to create usa
library(purrr)
usa <- map_data("usa")
str(usa)

# Build the map
ggplot(usa, aes(x = long, y = lat, group = group)) +
  geom_polygon() +
  coord_map() 
```
```{r}
# usa, cities, and all required packages are available

# Finish plot 1
# ggplot(usa, aes(x = long, y = lat, group = group)) +
#   geom_polygon() +
#   geom_point(data = cities, aes(group = State, size = Pop_est),
#              col = "red", shape = 16, alpha = 0.6) +
#   coord_map() +
#   theme_map()

# Arrange cities
# library(dplyr)
# cities_arr <- arrange(cities, Pop_est)
# 
# # Copy-paste plot 1 and adapt
# ggplot(usa, aes(x = long, y = lat, group = group)) +
#   geom_polygon(fill = "grey90") +
#   geom_point(data = cities_arr, aes(group = State, col = Pop_est),
#              shape = 16, size = 2) +
#   coord_map() +
#   theme_map() +
#  scale_color_viridis()

# pop and all required packages are available

# Use map_data() to create state
state <- map_data("state")

# Map of states
ggplot(state, aes(x = long, y = lat, fill = region, group = group)) +
  geom_polygon(col = "white") +
  coord_map() 
#  theme_nothing()

# Merge state and pop: state2
# state2 <- merge(state, pop)

# Map of states with populations
# ggplot(state2, aes(x = long, y = lat, fill = Pop_est, group = group)) +
#   geom_polygon(col = "white") +
#   coord_map() +
#   theme_map()
```


```{r}
# pop and all required packages are available

# Use map_data() to create state
# state <- map_data("state")
# 
# # Map of states
# ggplot(state, aes(x = long, y = lat, fill = region, group = group)) +
#   geom_polygon(col = "white") +
#   coord_map() +
#   theme_nothing()
# 
# # Merge state and pop: state2
# state2 <- merge(state, pop)
# 
# # Map of states with populations
# ggplot(state2, aes(x = long, y = lat, fill = Pop_est, group = group)) +
#   geom_polygon(col = "white") +
#   coord_map() +
#  theme_map()
```


```{r}
# All required packages are available

# # Import shape information: germany
# germany <- readOGR(dsn = "shapes", layer = "DEU_adm1")
# 
# # fortify germany: bundes
# bundes <- fortify(germany)
# 
# # Plot map of germany
# ggplot(bundes, aes(x = long, y = lat, group = group)) +
#     geom_polygon(fill = "blue", col = "white") +
#     coord_map() +
#    theme_nothing()
```


```{r}
library(ggplot2)
usa <- map_data("usa")
ggplot(usa, aes(long, lat, group = group)) +
 geom_polygon() +
 coord_map()
# 
# library(ggalt)
#  usa <- map_data("usa")
#  ggplot(usa, aes(long, lat, group = group)) +
#  geom_polygon() +
# coord_proj("+proj=wintri")
```
### Cartographic map

Drawn
● Topographical maps
● Altitude, infrastructure ...
● Photographic
● Satellite images
● Hybrid
● ggmap

```{r}
# Default style - zoom = 3
 library(ggmap)
# def_03 <- get_map(location = "Berlin, Germany", zoom = 3)
# ggmap(def_03, extent = "device")

#  library(ggmap)
#  def_13 <- get_map(location = "Berlin, Germany", zoom = 13)
# ggmap(def_13, extent = "device")
# 
# def_20 <- get_map(location = "Berlin, Germany", zoom = 20)
# ggmap(def_20, extent = "device")
# 
#  sat_13 <- get_map(location = "Berlin, Germany", zoom = 13,
#  source = "google", maptype = "satellite")
#ggmap(ton_13, extent = "device")
```

## Get coordinates

```{r}
berlin_sites <- c("Brandenburger Tor", "Potsdamer Platz",
 "Victory Column Berlin", "Checkpoint Charlie",
 "Reichstag Berlin", "Alexander Platz")
# xx <- geocode(berlin_sites)
# Information from URL : http://maps.googleapis.com/maps/...
# Information from URL : ...
 # Add column with cleaned up names
 xx$location <- sub(" Berlin",
"", berlin_sites)
 str(xx) 
```
## Animations

● Dense temporal data
● Great exploratory tool
● Several ways
● for loop to produce gif
● animation
● gganimate

## Motion Chart

Hans Rosling
● Karolinska Institute in Stockholm
● Founder of Gapminder
● UN data
● Life expectancy, GDP ...

```{r}
# gapminder <- read.delim("gapminder.tsv",
#  stringsAsFactors = FALSE) 

```

## Static plot

```{r}
library(gapminder)
ggplot(gapminder, aes(x = gdpPercap,
 y = lifeExp,
 colour = continent,
 size = pop)) +
 geom_point(alpha = 0.6) # details omitted
```

```{r}
p <- ggplot(gapminder, aes(x = gdpPercap,
 y = lifeExp,
 colour = continent,
 size = pop,
 frame = year)) +
 geom_point(alpha = 0.6) # details omitted
# gg_animate(p, "chart1.gif")
```

```{r}
# Inspect structure of japan
#str(japan)
#
## Finish the code inside saveGIF
#saveGIF({
#
#  # Loop through all time points
#  for (i in unique(japan$time)) {
#
#    # Subset japan: data
#    data <- subset(japan, time == i)
#
#    # Finish the ggplot command
#    p <- ggplot(data, aes(x = AGE, y = POP, fill = SEX, width = 1)) +
#      coord_flip() +
#      geom_bar(data = data[data$SEX == "Female",], stat = "identity") +
#      geom_bar(data = data[data$SEX == "Male",], stat = "identity") +
#      ggtitle(i)
#
#    print(p)
#
#  }
#
# }, movie.name = "pyramid.gif", interval = 0.1)
```

```{r}
# Vocab, gganimate and ggplot2 are available

# Update the static plot
# p <- ggplot(Vocab, aes(x = education, y = vocabulary,
#                        color = year, group = year,
#                        frame = year, cumulative = TRUE)) +
#     stat_smooth(method = "lm", se = FALSE, size = 3)
# 
# # Call gg_animate on p
# x <- gg_animate(p, filename = "vocab.gif", interval = 1.0)
```




# ggplot2 Internals

## ggplot2 internals

Explore grid graphics
● Elements of ggplot2 plot
● How do graphics work in R?
● 2 plo!ing systems
● base package
● grid graphics

```{r}
plot(mtcars$wt, mtcars$mpg, pch = 16, col = "#00000080")

plot(mtcars$wt, mtcars$mpg, pch = 16, col = "#C3212766",
 cex = 2, xlab = "Weight", ylab = "mpg (US gallons)")
 abline(lm(mpg ~ wt, data = mtcars), col = "red")
 
  plot(mtcars$wt, mtcars$mpg, pch = 16, col = "#00000080")
 abline(lm(mpg ~ wt, data = mtcars), col = "red")
 points(mtcars$wt, mtcars$mpg, pch = 16,
 col = "#C3212766", cex = 2)
```

## Graphic output

```{r}
# grid.rect()
# grid.lines()

# grid.rect(x = 0.5, y = 0.5,
#  width = 0.5, height = 0.5,
#  just = "center",
#  gp = gpar(fill = "darkgreen"))

#  grid.rect(gp = gpar(fill = "grey80"))
# ...
#  grid.points(mtcars$wt,mtcars$mpg, pch = 16,
#  gp = gpar(col = "#00000080"),
#  name = "datapoints")
#  grid.edit("datapoints",
#  gp = gpar(col = "#C3212766", cex = 4))
```

## Grobs

● Graphical objects = grobs
● ggplot2 object = collection of grobs

```{r}
p <- ggplot(iris, aes(x = Sepal.Length,
 y = Sepal.Width,
 col = Species)) +
 geom_point(alpha = 0.3, size = 5, shape = 16) +
 geom_smooth(method = "lm", se = FALSE) +
 scale_y_continuous("Width", limits = c(2, 4.5), expand = c(0,0)) +
 scale_x_continuous("Length", limits = c(4, 8), expand = c(0,0)) +
 coord_equal() +
 ggtitle("Iris Sepals") +
 theme(rect = element_blank())
```

```{r}
 library(grid)
g <- ggplotGrob(p)
 g

library(gtable)
# gtable_show_layout(g$grob[[8]])
```
```{r}
# Populate null viewport
grid.rect(gp = gpar(fill = "grey90"))
grid.text("null viewport")
grid.lines(x = c(0,0.75), y = c(0.25, 1),
           gp = gpar(lty = 2, col = "red"))

# Create new viewport: vp
vp <- viewport(x = 0.5, y = 0.5, width = 0.5, height = 0.5, just = "center")

# Push vp
pushViewport(vp)

# Populate new viewport with rectangle
grid.rect(gp = gpar(fill = "blue"))
```

```{r}
# 1 - Create plot viewport: pvp
mar <- c(5, 4, 2, 2)
pvp <- plotViewport(mar)

# 2 - Push pvp
pushViewport(pvp)

# 3 - Add rectangle
grid.rect(gp = gpar(fill = "grey80"))

# Create data viewport: dvp
dvp <- dataViewport(xData = mtcars$wt, yData = mtcars$mpg)

# 4 - Push dvp
pushViewport(dvp)

# Add two axes
grid.xaxis()
grid.yaxis()
```


```{r}
# Work from before
pushViewport(plotViewport(c(5, 4, 2, 2)))
grid.rect(gp = gpar())
pushViewport(dataViewport(xData = mtcars$wt, yData = mtcars$mpg))
grid.xaxis()
grid.yaxis()

# Work from before - add names
grid.text("Weight", y = unit(-3, "lines"), name = "xaxis")
grid.text("MPG", x = unit(-3, "lines"), rot = 90, name = "yaxis")
grid.points(x = mtcars$wt, y = mtcars$mpg, pch = 16, name = "datapoints")

# Edit "xaxis"
grid.edit("xaxis", label = "Weight (1000 lbs)")

# Edit "yaxis"
grid.edit("yaxis", label = "Miles/(US) gallon")

# Edit "datapoints"
grid.edit("datapoints", gp = gpar(col = "#C3212766", cex = 2))
```

```{r}
# Code from before
p <- ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) + geom_point()
gtab <- ggplotGrob(p)
g <- gtab$grobs
legend_index <- which(vapply(g, inherits, what = "gtable", logical(1)))
grid.draw(g[[legend_index]])

# 1 - Show layout of legend grob
gtable_show_layout(g[[legend_index]])

# Create text grob
my_text <- textGrob(label = "Motor Trend, 1974", gp = gpar(fontsize = 7, col = "gray25"))

# 2  -Use gtable_add_grob to modify original gtab
new_legend <- gtable_add_grob(gtab$grobs[[legend_index]], my_text, 3, 2)

# 3 - Update in gtab
gtab$grobs[[legend_index]] <- new_legend

# 4 - Draw gtab
grid.draw(gtab)
```

```{r}
# Box plot of mtcars: p
p <- ggplot(mtcars, aes(x = factor(cyl), y = wt)) + geom_boxplot()

# Build pdata
pdata <- ggplot_build(p)$data

# Access the first element of the list, a data frame
class(pdata[[1]])

# Isolate this data frame
my_df <- pdata[[1]]

# The x labels
my_df$group <- c("4", "6", "8")

# Print out specific variables
my_df[c(1:6, 11)]
```


```{r}
# # Definitions of g1 and g2
# g1 <- ggplot(mtcars, aes(wt, mpg, col = cyl)) +
#   geom_point() +
#   theme(legend.position = "bottom")
# 
# g2 <- ggplot(mtcars, aes(disp, fill = cyl)) +
#   geom_histogram(binwidth = 20) +
#   theme(legend.position = "none")
# 
# # Extract the legend from g1
# my_legend <- ggplotGrob(g1)$grobs[[legend_index]]  
# 
# # Create g1_noleg
# g1_noleg <- g1 + theme(legend.position = "none")
# 
# # Calculate the height: legend_height
# legend_height <- sum(my_legend$heights)
# 
# # Arrange g1_noleg, g2 and my_legend
# grid.arrange(g1_noleg, g2, my_legend,
#              layout_matrix = matrix(c(1, 3, 2, 3), ncol = 2),
#              heights = unit.c(unit(1, "npc") - legend_height, legend_height))
```


# Data Munging and Visualization Case Study

## Case Study I
Bag Plot

```{r}
# df$ID <- seq_len(nrow(df) / 2)
# ggplot(df, aes(x = type, Value, group = ID)) +
# geom_line(alpha = 0.3)
```


```{r}
# library(viridis)
#  ggplot(dat, aes(x = group1, y = group2)) +
#  stat_density_2d(geom = "tile", aes(fill = ..density..),
#  contour = FALSE) +
# scale_fill_viridis()
```
```{r}
library(aplpack)
# bagplot(dat[1:2])
```
```{r}
# ggplot(past, aes(x = new_day, y = temp, group = year)) +
# geom_line(alpha = 0.4) +
# geom_line(data = present, aes(group = 1), col = "red")
```


```{r}
# ggplot(my_data, aes(x = new_day, y = temp, fill = year)) +
#  stat_historical()
```

```{r}
# Call bagplot() on test_data
# bagplot(test_data)
# 
# # Call compute.bagplot on test_data: bag
# bag <- compute.bagplot(test_data)
# 
# # Display information
# bag$hull.loop
# bag$hull.bag
# bag$pxy.outlier
# 
# # Highlight components
# points(bag$hull.loop, col = "green", pch = 16)
# points(bag$hull.bag, col = "orange", pch = 16)
# points(bag$pxy.outlier, col = "purple", pch = 16)


# # Call bagplot() on test_data
# bagplot(test_data)
# 
# # Call compute.bagplot on test_data: bag
# bag <- compute.bagplot(test_data)
# 
# # Display information
# bag$hull.loop
# bag$hull.bag
# bag$pxy.outlier
# 
# # Highlight components
# points(bag$hull.loop, col = "green", pch = 16)
# points(bag$hull.bag, col = "orange", pch = 16)
# points(bag$pxy.outlier, col = "purple", pch = 16)
```

```{r}
# ggproto for StatLoop (hull.loop)
# StatLoop <- ggproto("StatLoop", Stat,
#                     required_aes = c("x", "y"),
#                     compute_group = function(data, scales) {
#                       bag <- compute.bagplot(x = data$x, y = data$y)
#                       data.frame(x = bag$hull.loop[,1], y = bag$hull.loop[,2])
#                     })
# 
# # ggproto for StatBag (hull.bag)
# StatBag <- ggproto("StatBag", Stat,
#                    required_aes = c("x", "y"),
#                    compute_group = function(data, scales) {
#                      bag <- compute.bagplot(x = data$x, y = data$y)
#                      data.frame(x = bag$hull.bag[,1], y = bag$hull.bag[,2])
#                    })
# 
# # ggproto for StatOut (pxy.outlier)
# StatOut <- ggproto("StatOut", Stat,
#                    required_aes = c("x", "y"),
#                    compute_group = function(data, scales) {
#                      bag <- compute.bagplot(x = data$x, y = data$y)
#                      data.frame(x = bag$pxy.outlier[,1], y = bag$pxy.outlier[,2])
#                   })
```

```{r}
# StatLoop, StatBag and StatOut are available

# # Combine ggproto objects in layers to build stat_bag()
# stat_bag <- function(mapping = NULL, data = NULL, geom = "polygon",
#                      position = "identity", na.rm = FALSE, show.legend = NA,
#                      inherit.aes = TRUE, loop = FALSE, ...) {
#   list(
#     # StatLoop layer
#     layer(
#       stat = StatLoop, data = data, mapping = mapping, geom = geom, 
#       position = position, show.legend = show.legend, inherit.aes = inherit.aes,
#       params = list(na.rm = na.rm, alpha = 0.35, col = NA, ...)
#     ),
#     # StatBag layer
#     layer(
#       stat = StatBag, data = data, mapping = mapping, geom = geom, 
#       position = position, show.legend = show.legend, inherit.aes = inherit.aes,
#       params = list(na.rm = na.rm, alpha = 0.35, col = NA, ...)
#     ),
#     # StatOut layer
#     layer(
#       stat = StatOut, data = data, mapping = mapping, geom = "point", 
#       position = position, show.legend = show.legend, inherit.aes = inherit.aes,
#       params = list(na.rm = na.rm, alpha = 0.7, col = NA, shape = 21, ...)
#     )
#   )
# }
```

```{r}
# hull.loop, hull.bag and pxy.outlier are available
# stat_bag, test_data and test_data2 are available

# Previous method
# ggplot(test_data, aes(x = x, y = y)) +
#   geom_polygon(data = hull.loop, fill = "green") +
#   geom_polygon(data = hull.bag, fill = "orange") +
#   geom_point(data = pxy.outlier, col = "purple", pch = 16, cex = 1.5)
# 
# # stat_bag
# ggplot(test_data, aes(x = x, y = y)) +
#   stat_bag(fill = 'black')
# 
# # stat_bag on test_data2
# ggplot(test_data2, aes(x = x, y = y, fill = treatment)) +
#  stat_bag()
```
```{r}
# Import weather data
# weather <- read.fwf("NYNEWYOR.txt",
#                     header = FALSE,
#                     col.names = c("month", "day", "year", "temp"),
#                     widths = c(14, 14, 13, 4))
# 
# # Check structure of weather
# str(weather)
# 
# # Create past with two filter() calls
# past <- weather %>%
#   filter(!(month == 2 & day == 29)) %>%
#   filter(year != max(year))
#   
# # Check structure of past
# str(past)
```


```{r}
# Finish the clean_weather function
clean_weather <- function(file) {
  weather <- read.fwf(file,
                      header = FALSE,
                      col.names = c("month", "day", "year", "temp"),
                      widths = c(14, 14, 13, 4))
  weather %>%
    filter(!(month == 2 & day == 29)) %>%
    group_by(year) %>%
    mutate(yearday = 1:length(day)) %>%
    ungroup() %>%
    filter(temp != -99)
}

# Import NYNEWYOR.txt: my_data
# my_data <- clean_weather("NYNEWYOR.txt")
```

```{r}
# Create the stats object
StatHistorical <- ggproto("StatHistorical", Stat,
                    compute_group = function(data, scales, params) {
                      data <- data %>%
                        filter(year != max(year)) %>%
                        group_by(x) %>%
                        mutate(ymin = Hmisc::smean.cl.normal(y)[3],
                               ymax = Hmisc::smean.cl.normal(y)[2]) %>%
                        ungroup()
                    },
                    required_aes = c("x", "y", "year"))

# Create the layer
stat_historical <- function(mapping = NULL, data = NULL, geom = "point",
                            position = "identity", na.rm = FALSE, show.legend = NA, 
                            inherit.aes = TRUE, ...) {
  list(
    layer(
      stat = "identity", data = data, mapping = mapping, geom = geom,
      position = position, show.legend = show.legend, inherit.aes = inherit.aes,
      params = list(na.rm = na.rm, col = "#EED8AE", alpha = 0.3, shape = 16, ...)
    ),
    layer(
      stat = StatHistorical, data = data, mapping = mapping, geom = "linerange",
      position = position, show.legend = show.legend, inherit.aes = inherit.aes,
      params = list(na.rm = na.rm, col = "#8B7E66", ...)
    )
  )
}

# Build the plot
# my_data <- clean_weather("NYNEWYOR.txt")
# ggplot(my_data, aes(x = yearday, y = temp, year = year)) +
#   stat_historical()
```
```{r}
# Create the stats object
StatPresent <- ggproto("StatPresent", Stat,
                       compute_group = function(data, scales, params) {
                         data <- filter(data, year == max(year))
                       },
                       required_aes = c("x", "y", "year"))

# Create the layer
stat_present <- function(mapping = NULL, data = NULL, geom = "line",
                       position = "identity", na.rm = FALSE, show.legend = NA, 
                       inherit.aes = TRUE, ...) {
  layer(
    stat = StatPresent, data = data, mapping = mapping, geom = geom,
    position = position, show.legend = show.legend, inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}

# # Build the plot
# my_data <- clean_weather("NYNEWYOR.txt")
# ggplot(my_data, aes(x = yearday, y = temp, year = year)) +
#   stat_historical() +
#  stat_present()
```


```{r}
# Create the stats object
StatExtremes <- ggproto("StatExtremes", Stat,
                        compute_group = function(data, scales, params) {
                          
                          present <- data %>%
                            filter(year == max(year)) 
                          
                          past <- data %>%
                            filter(year != max(year)) 
                          
                          past_extremes <- past %>%
                            group_by(x) %>%
                            summarise(past_low = min(y),
                                      past_high = max(y))
                          
                          # transform data to contain extremes
                          data <- present %>%
                            left_join(past_extremes) %>%
                            mutate(record = ifelse(y < past_low, 
                                                   "#0000CD", 
                                                   ifelse(y > past_high, 
                                                          "#CD2626", 
                                                          "#00000000")))

                        },
                        required_aes = c("x", "y", "year"))

# Create the layer
stat_extremes <- function(mapping = NULL, data = NULL, geom = "point",
                          position = "identity", na.rm = FALSE, show.legend = NA, 
                          inherit.aes = TRUE, ...) {
  layer(
    stat = StatExtremes, data = data, mapping = mapping, geom = geom,
    position = position, show.legend = show.legend, inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}

# Build the plot
# my_data <- clean_weather("NYNEWYOR.txt")
# ggplot(my_data, aes(x = yearday, y = temp, year = year)) +
#   stat_historical() +
#   stat_present() +
#   stat_extremes(aes(col = ..record..)) +
#  scale_color_identity()
```

```{r}
# File paths of all datasets
my_files <- c("NYNEWYOR.txt","FRPARIS.txt", "ILREYKJV.txt", "UKLONDON.txt")

# Build my_data with a for loop
# my_data <- NULL
# for (file in my_files) {
#   temp <- clean_weather(file)
#   temp$id <- sub(".txt", "", file)
#   my_data <- rbind(my_data, temp)
# }
# 
# # Build the final plot, from scratch!
# ggplot(my_data, aes(x = yearday, y = temp, year = year)) +
#   stat_historical() +
#   stat_present() +
#   stat_extremes(aes(col = ..record..)) +
#   scale_color_identity() +
# facet_wrap(~id, ncol = 2)
```











































