---
title: "Chapter 2 Homework Analysis 2"
author: "Jefferson Ong and Céline Prunet"
date: "1/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# (A)
library(ISLR)
library(tidyverse)
college <- College
auto <- Auto
attach(college)
```

Please complete problems 8 and 9 from chapter 2 on pp 54-56 of the text.  Due date is Friday, Jan 31st.  Please write up your answers using R Markdown.  For this assignment, please e-mail me your final submission.

You will be allowed to work in groups of two if you prefer to do that for this homework.

If you have any questions please let me know.

Ross Gosky


# 8 

This exercise relates to the College data set, which can be found in
the file College.csv. It contains a number of variables for 777 different
universities and colleges in the US. The variables are
• Private : Public/private indicator
• Apps : Number of applications received
• Accept : Number of applicants accepted
• Enroll : Number of new students enrolled
• Top10perc : New students from top 10 % of high school class
• Top25perc : New students from top 25 % of high school class
• F.Undergrad : Number of full-time undergraduates
• P.Undergrad : Number of part-time undergraduates
• Outstate : Out-of-state tuition
• Room.Board : Room and board costs
• Books : Estimated book costs
• Personal : Estimated personal spending
• PhD : Percent of faculty with Ph.D.’s
• Terminal : Percent of faculty with terminal degree
• S.F.Ratio : Student/faculty ratio
• perc.alumni : Percent of alumni who donate
• Expend : Instructional expenditure per student
• Grad.Rate : Graduation rate

(a) Use the read.csv() function to read the data into R. Call the
loaded data college. Make sure that you have the directory set
to the correct location for the data.

(b) Look at the data using the fix() function. You should notice
that the first column is just the name of each university. We don’t
really want R to treat this as data. However, it may be handy to
have these names for later. Try the following commands:

2.4 Exercises 55

```{r}
str(college)
```





```{r}
#rownames(college) = make.names(nams, unique=TRUE)
#rownames(college)=college[,1]
# fix(college)
```



You should see that there is now a row.names column with the
name of each university recorded. This means that R has given
each row a name corresponding to the appropriate university. R
will not try to perform calculations on the row names. However,
we still need to eliminate the first column in the data where the
names are stored. Try



```{r}
#college =college[,-1]
#fix(college)
```

Now you should see that the first data column is Private. Note
that another column labeled row.names now appears before the
Private column. However, this is not a data column but rather
the name that R is giving to each row.

(c) i. Use the summary() function to produce a numerical summary
of the variables in the data set.

```{r}
summary(college)
```


ii. Use the pairs() function to produce a scatterplot matrix of
the first ten columns or variables of the data. Recall that
you can reference the first ten columns of a matrix A using
A[,1:10].

```{r}
pairs(college)
pairs(college[,1:10])
```


iii. Use the plot() function to produce side-by-side boxplots of
Outstate versus Private.

```{r}
 boxplot(Outstate ~Private)

```


iv. Create a new qualitative variable, called Elite, by binning
the Top10perc variable. We are going to divide universities
into two groups based on whether or not the proportion
of students coming from the top 10 % of their high school
classes exceeds 50 %.


```{r}
# college %>%
#   mutate(Elite = Top10perc > 50)
```


```{r}

Elite = rep("No",nrow(college ))
Elite[college$Top10perc > 50] = "Yes"
Elite = as.factor(Elite)
college = data.frame(college , Elite)
str(college)
```

```{r}
summary(college)
```


Use the summary() function to see how many elite universities there are. Now use the plot() function to produce
side-by-side boxplots of Outstate versus Elite.

```{r}
#plot(college$Outstate, college$Elite) 
boxplot(college$Outstate~ college$Elite, col = c("blue", "green"), main = "Outstate versus Elite", 
    xlab = "Elite", ylab = "Outstate")
#fivenum(college$Outstate, college$Elite)
```

v. Use the hist() function to produce some histograms with
differing numbers of bins for a few of the quantitative variables. You may find the command par(mfrow=c(2,2)) useful:
it will divide the print window into four regions so that four
plots can be made simultaneously. Modifying the arguments
to this function will divide the screen in other ways.

```{r}

par(mfrow=c(2,2))
hist(Personal)
hist(Books)
hist(Outstate)
hist( Apps)
```

vi. Continue exploring the data, and provide a brief summary
of what you discover.

11% of all colleges are considered elite since those colleges has a student base that comes from the top 10% of their high school as the vast majority of their population. The rest of the colleges, 89%, do not have over 50% of their student population coming from the top 10% percent of their high school. 

The boxplot shows this disparity where non elite colleges has a higher IQR. 

# 9 

This exercise involves the Auto data set studied in the lab. Make sure
that the missing values have been removed from the data.

(a) Which of the predictors are quantitative, and which are qualitative?
```{r}
auto2 <- Auto
str(auto2)
```



```{r}
attach(auto)
auto <- auto[ -c(9) ]
str(auto)
```

quantitative: mpg, cylinders, displacement, horsepower, weight, acceleration, year, origin

qualitative: name

(b) What is the range of each quantitative predictor? You can answer this using the range() function. range()

```{r}
range(mpg)
range(cylinders)
range(displacement)
range(horsepower)
range(weight)
range(acceleration)
range(year)
range(origin)
```


(c) What is the mean and standard deviation of each quantitative
predictor?

```{r}
mean(mpg)
sd(mpg)

mean(cylinders)
sd(cylinders)

mean(displacement)
sd(displacement)

mean(horsepower)
sd(horsepower)

mean(weight)
sd(weight)

mean(acceleration)
sd(acceleration)

mean(year)
sd(year)

mean(origin)
sd(origin)

colMeans(auto)
```


(d) Now remove the 10th through 85th observations. What is the
range, mean, and standard deviation of each predictor in the
subset of the data that remains?



```{r}
#str(auto)



auto %>%
 summarise(mean = mean(mpg), sd = sd(mpg), min = min(mpg), max = max(mpg))


auto %>%
 summarise(mean = mean(cylinders), sd = sd(cylinders), min = min(cylinders), max = max(cylinders))
```


```{r}
 output <- sapply(auto[-c(10:85), ], function(x) round(c(range(x), mean(x), sd(x)), 2))
 rownames(output) <- c("min", "max", "mean", "sd")
 output


```


(e) Using the full data set, investigate the predictors graphically,
using scatterplots or other tools of your choice. Create some plots
highlighting the relationships among the predictors. Comment
on your findings.

```{r}
#ggplot(auto, aes(x = weight, y = mpg)) +
 #geom_point()

plot(auto2$mpg ~ auto2$year)

#lm(mpg ~ year, auto2)

ggplot(auto2, aes(x = horsepower, y = acceleration)) +
  geom_point() +
  geom_smooth(method = 'lm', se = F)
```



(f) Suppose that we wish to predict gas mileage (mpg) on the basis
of the other variables. Do your plots suggest that any of the
other variables might be useful in predicting mpg? Justify your
answer.

```{r}
pot <-lm(mpg ~ cylinders +displacement+ horsepower+ weight+ acceleration+ year+ origin, auto2)
summary(pot)
```
```{r}
pot2 <-lm(mpg ~ weight + year+ origin, auto2)
summary(pot2)
```

We did a linear fit of all the other quantitative variable to predict mpg, we only kept weight, year, and origin because of their t-statistic's low p- value. 



## Questions

What is sapply/lapply doing to the dataset
Why beginning code for 8 did not work?
how to create the elite tab using dplyr functions
How to interpret 8 vi (histograms)
Why 9 a) auto does not work anymore (does not have the right number of observations)



