---
title: "Data Visualization with ggplot2 (Part 2)"
author: "Jefferson Ong"
date: "3/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(lattice)
```

# Statistics

Two categories of functions
● Called from within a geom
● Called independently 

```{r}
 p <- ggplot(iris, aes(x = Sepal.Width))
 p + geom_histogram()
 p + geom_bar()
 p + stat_bin()
```
```{r}
 ggplot(mtcars, aes(x = mpg)) +
 geom_histogram(fill = "skyblue") +
 geom_freqpoly(col = "red") 
```

```{r}
 ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) +
 geom_point() +
 geom_smooth(method = "lm", fullrange = TRUE)
```

## Calculating Statistics

```{r}
set.seed(123)
 xx <- rnorm(100)
 mean(xx)

 mean(xx) + (sd(xx) * c(-1, 1))
 library(Hmisc)
 smean.sdl(xx, mult = 1)
```

```{r}
ggplot(iris, aes(x = Species, y = Sepal.Length)) +
 stat_summary(fun.y = mean, geom = "point") +
 stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1),
 geom = "errorbar", width = 0.1)

ggplot(iris, aes(x = Species, y = Sepal.Length)) +
 stat_summary(fun.y = mean, geom = "bar", fill = "skyblue") +
 stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1),
 geom = "errorbar", width = 0.1)
```

## Normal distribution

```{r}
library(MASS)
 mam.new <- data.frame(body = log10(mammals$body))
 ggplot(mam.new, aes(x = body)) +
 geom_histogram(aes( y = ..density..)) +
 geom_rug() +
 stat_function(fun = dnorm, colour = "red",
 arg = list(mean = mean(mam.new$body),
 sd = sd(mam.new$body)))
```

## QQ plot

```{r}
mam.new$slope <- diff(quantile(mam.new$body, c(0.25, 0.75))) /
 diff(qnorm(c(0.25, 0.75)))
mam.new$int <- quantile(mam.new$body, 0.25) -
 mam.new$slope * qnorm(0.25)
ggplot(mam.new, aes(sample = body)) +
 stat_qq() +
 geom_abline(aes(slope = slope, intercept = int), col = "red")
```

## Practice 1

```{r}
# Plot 1: Jittered scatter plot, add a linear model (lm) smooth
#ggplot(Vocab, aes(x = education, y = vocabulary)) +
#  geom_jitter(alpha = 0.2) +
#  stat_smooth(method = "lm", se = FALSE) # smooth
#
## Plot 2: points, colored by year
#ggplot(Vocab, aes(x = education, y = vocabulary, col = year)) +
#  geom_jitter(alpha = 0.2)
#  
## Plot 3: lm, colored by year
#ggplot(Vocab, aes(x = education, y = vocabulary, col = factor(year))) +
#  stat_smooth(method = "lm", se = FALSE) # smooth
#
## Plot 4: Set a color brewer palette
#ggplot(Vocab, aes(x = education, y = vocabulary, col = factor(year))) +
#  stat_smooth(method = "lm", se = FALSE) +  # smooth
#  scale_color_brewer()  # colors
#
## Plot 5: Add the group aes, specify alpha and size
#ggplot(Vocab, aes(x = education, y = vocabulary, col = year, group = factor(year))) +
#  stat_smooth(method = "lm", se = FALSE, alpha = 0.6, size = 2) +
#  scale_color_gradientn(colors = brewer.pal(9, "YlOrRd"))
```

```{r}
## Plot 1: Jittering only
#p <- ggplot(Vocab, aes(x = education, y = vocabulary)) +
#  geom_jitter(alpha = 0.2)
#
## Plot 2: Add stat_sum
#p +
#  stat_sum() # sum statistic
#
## Plot 3: Set size range
#p +
#  stat_sum() + # sum statistic
#  scale_size(range = c(1, 10)) # set size scale
```

```{r}
# Display structure of mtcars
str(mtcars)

# Convert cyl and am to factors
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$am <- as.factor(mtcars$am)

# Define positions
posn.d <- position_dodge(0.1)
posn.jd <- position_jitterdodge(jitter.width = 0.1, dodge.width = 0.2)
posn.j <- position_jitter(0.2)

# Base layers
wt.cyl.am <- ggplot(mtcars, aes(x = cyl,
                                y = wt,
                                col = am,
                                fill = am,
                                group = am))
```

## Custom Functions

```{r}
# Play vector xx is available

# Function to save range for use in ggplot
gg_range <- function(x) {
  # Change x below to return the instructed values
  data.frame(ymin = min(x), # Min
             ymax = max(x)) # Max
}

gg_range(xx)
# Required output
#   ymin ymax
# 1    1  100

# Function to Custom function:
med_IQR <- function(x) {
  # Change x below to return the instructed values
  data.frame(y =  median(x), # Median
             ymin = quantile(x)[2], # 1st quartile
             ymax = quantile(x)[4])  # 3rd quartile
}

med_IQR(xx)
# Required output
#        y  ymin  ymax
# 25% 50.5 25.75 75.25
```

```{r}
# The base ggplot command; you don't have to change this
#wt.cyl.am <- ggplot(mtcars, aes(x = cyl,y = wt, col = am, fill = am, group = am))
#
## Add three stat_summary calls to wt.cyl.am
#wt.cyl.am +
#  stat_summary(geom = "linerange", fun.data = med_IQR,
#               position = posn.d, size = 3) +
#  stat_summary(geom = "linerange", fun.data = gg_range,
#               position = posn.d, size = 3,
#               alpha = 0.4) +
#  stat_summary(geom = "point", fun.y = median,
#               position = posn.d, size = 3,
#               col = "black", shape = "X")
```




# Coordinates and Facets

## Zooming in


scale_x_continuous(limits = ...)
● xlim()
● coord_cartesian(xlim = ...)



```{r}
 iris.smooth <- ggplot(iris, aes(x = Sepal.Length,
 y = Sepal.Width,
 col = Species)) +
 geom_point(alpha = 0.7) + geom_smooth()
 iris.smooth
```

```{r}
 iris.smooth + scale_x_continuous(limits = c(4.5, 5.5))

```


```{r}
 iris.smooth + xlim(c(4.5, 5.5))
```

```{r}
 iris.smooth + coord_cartesian(xlim = c(4.5, 5.5))
```


## Aspect Ratio

● Height-to-width ratio
● Deception!
● Standardization a!empts
● Typically 1:1

```{r}
 library(reshape2); library(zoo)
 sunspots.m <- data.frame(year = index(sunspot.month),
 value = melt(sunspot.month)$value)
 ggplot(sunspots.m, aes(x = year, y = value)) +
 geom_line() +
 coord_equal() # a 1:1 aspect ratio
```


```{r}
ggplot(sunspots.m, aes(x = year, y = value)) +
 geom_line() +
 coord_fixed(0.055)
```

## Facets

● Straight-forward yet useful
● Concept of Small Multiples
● Edward Tu!e
● Visualization of Quantitative Information, 1983


## iris.tidy

```{r}
#ggplot(iris.tidy, aes(x = Measure, y = Value, col = Part)) +
# geom_jitter() +
# facet_grid(. ~ Species)
```


## Basic facet

```{r}
# Basic scatter plot
p <- ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point()

# 1 - Separate rows according to transmission type, am
p +
  facet_grid(am ~ .)

# 2 - Separate columns according to cylinders, cyl
p +
  facet_grid(. ~ cyl)

# 3 - Separate by both columns and rows 
p +
  facet_grid(am ~ cyl)
```


```{r}
## Code to create the cyl_am col and myCol vector
#mtcars$cyl_am <- paste(mtcars$cyl, mtcars$am, sep = "_")
#myCol <- rbind(brewer.pal(9, "Blues")[c(3,6,8)],
#               brewer.pal(9, "Reds")[c(3,6,8)])
#
## Map cyl_am onto col
#ggplot(mtcars, aes(x = wt, y = mpg, col = cyl_am)) +
#  geom_point() +
#  # Add a manual colour scale
#  scale_color_manual(values = myCol)
#  
## Grid facet on gear vs. vs
#ggplot(mtcars, aes(x = wt, y = mpg, col = cyl_am)) +
#  geom_point() +
#  scale_color_manual(values = myCol) +
#  facet_grid(gear ~ vs)
#
## Also map disp to size
#ggplot(mtcars, aes(x = wt, y = mpg, col = cyl_am, size = disp)) +
#  geom_point() +
#  scale_color_manual(values = myCol) +
#  facet_grid(gear ~ vs)
```


```{r}
## Basic scatter plot
#p <- ggplot(mamsleep, aes(x = time, y = name, col = sleep)) +
#  geom_point()
#  
## Execute to display plot
#p
#
## Facet rows accoding to vore
#p +
#  facet_grid(vore ~ .)
#
## Specify scale and space arguments to free up rows
#p +
#  facet_grid(vore ~ ., scale = "free_y", space = "free_y")
```




```{r}
# Basic ggplot() command, coded for you
p <- ggplot(mtcars, aes(x = wt, y = hp, col = am)) + geom_point() + geom_smooth()

# Add scale_x_continuous()
p + 
  scale_x_continuous(limits = c(3, 6), expand = c(0, 0))

# Add coord_cartesian(): the proper way to zoom in
p +
  coord_cartesian(xlim = c(3, 6))
```

```{r}
## Extend z with theme() function and 3 args
#z + 
#  theme(panel.grid = element_blank(),
#        axis.line = element_line(color = "red"),
#        axis.ticks = element_line(color = "red"))
```

```{r}
## Original plot, color provided
#z
#myRed
#
## Extend z with theme() function and 3 args
#z +
#  theme(strip.text = element_text(size = 16, color = myRed),
#        axis.title = element_text(color = myRed, hjust = 0, face = "italic"),
#        axis.text = element_text(color = "black"))
```

```{r}
## Move legend by position
#z +
#  theme(legend.position = c(0.85, 0.85))
#
## Change direction
#z +
#  theme(legend.direction = "horizontal")
#
## Change location by name
#z +
#  theme(legend.position = "bottom")
#
## Remove legend entirely
#z +
#  theme(legend.position = "none")
#
## Increase spacing between facets
#library(grid)
#z + theme(panel.spacing.x = unit(2, "cm"))
#
## Adjust the plot margin
#z + theme(panel.spacing.x = unit(2, "cm"),
#          plot.margin = unit(c(1,2,1,1), "cm"))
```


# Themes

## Themes Layer

● All the non-data ink
● Visual elements not part of data
● Three types
● text
● line
● rectangle
element_text()
element_line()
element_rect()

```{r}
## Starting point
#z
#
## Plot 1: Change the plot background fill to myPink
#z + 
#  theme(plot.background = element_rect(fill = myPink))
#
## Plot 2: Adjust the border to be a black line of size 3
#z + 
#  theme(plot.background = element_rect(fill = myPink, color = "black", size = 3)) # expanded from plot 1
#
## Theme to remove all rectangles
#no_panels <- theme(rect = element_blank())
#
## Plot 3: Combine custom themes
#z +
#  no_panels +
#  theme(plot.background = element_rect(fill = myPink, color = "black", size = 3))  # from plot 2
```




## Recycling Themes

Many plots
● Consistency in style
● Apply specific theme everywhere

```{r}
z <- ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) +
 geom_jitter(alpha = 0.7) +
 scale_color_brewer("Species",
 palette = "Dark2",
 labels = c("Setosa",
 "Versicolor",
"Virginica")) +
 scale_y_continuous("Width (cm)", limits = c(2, 4.5), expand = c(0, 0)) +
 scale_x_continuous("Length (cm)", limits = c(4, 8), expand = c(0, 0)) +
 ggtitle("Sepals") +
 coord_fixed(1)
```

```{r}
# z + theme(panel.background = element_blank(),
# legend.background = element_blank(),
# legend.key = element_blank(),
# panel.grid = element_blank(),
# axis.text = element_text(colour = "black"),
# axis.line = element_line(colour = "black"))
#
# z + theme_iris
```

```{r}
# m <- ggplot(iris.wide, aes(x = Length, y = Width, col = Part)) +
# geom_point() +
# facet_grid(. ~ Species)

# m + theme_iris
# 
# theme_iris <- theme_iris +
# theme(strip.background = element_blank()) 

#  m + theme_iris

```

## Discrete x-axis

```{r}
#p <- ggplot(iris.tidy, aes(x = Measure, y = Value, col = Part)) +
# geom_point(position = position_jitter(0.1), alpha = 0.6,
# width = 0.4) +
# scale_y_continuous("Value (cm)", limits = c(0, 8),
# expand = c(0, 0)) +
# facet_grid(. ~ Species)
# p
```


```{r}
# # Original plot
# z2
# 
# # Load ggthemes
# library(ggthemes)
# 
# # Apply theme_tufte, plot additional modifications
# custom_theme <- theme_tufte() +
#   theme(legend.position = c(0.9, 0.9),
#         legend.title = element_text(face = "italic", size = 12),
#         axis.title = element_text(face = "bold", size = 14))
# 
# # Draw the customized plot
# z2 + custom_theme
#  
# # Use theme set to set custom theme as default
# theme_set(custom_theme)
# 
# # Plot z2 again
# z2
```




# Best Practices

## Bar Plots

● Two types
● Absolute values
● Distribution


```{r}
# d <- ggplot(sleep, aes(vore, total)) +
# scale_y_continuous("Total sleep time (h)",
# limits = c(0, 24),
# breaks = seq(0, 24, 3),
# expand = c(0, 0)) +
# scale_x_discrete("Eating habits") +
# theme_classic()
# d +
# stat_summary(fun.y = mean, geom = "bar",
# fill = "grey50") +
# stat_summary(fun.data = mean_sdl, mult = 1,
# geom = "errorbar", width = 0.2)
```

```{r}
# Base layers
#m <- ggplot(mtcars.cyl, aes(x = cyl, y = wt.avg))
#
## Plot 1: Draw bar plot with geom_bar
#m + geom_bar(stat = "identity", fill = "skyblue")
#
## Plot 2: Draw bar plot with geom_col
#m + geom_col(fill = "skyblue")
#
## Plot 3: geom_col with variable widths.
#m + geom_col(fill = "skyblue", width = mtcars.cyl$prop)
#
## Plot 4: Add error bars
#m +
#  geom_col(fill = "skyblue", width = mtcars.cyl$prop) +
#  geom_errorbar(aes(ymin = wt.avg - sd, ymax = wt.avg + sd), width = 0.1)
```


## Heat Maps

```{r}
# # Create color palette
# myColors <- brewer.pal(9, "Reds")
# 
# # Build the heat map from scratch
# ggplot(barley, aes(x = year, y = variety, fill = yield)) +
#   geom_tile() + # Geom layer
#   facet_wrap( ~ site, ncol = 1) + # Facet layer
#  scale_fill_gradientn(colors = myColors) # Adjust colors
```

```{r}
# The heat map we want to replace
# Don't remove, it's here to help you!
# myColors <- brewer.pal(9, "Reds")
# ggplot(barley, aes(x = year, y = variety, fill = yield)) +
#   geom_tile() +
#   facet_wrap( ~ site, ncol = 1) +
#   scale_fill_gradientn(colors = myColors)
# 
# # Line plot; set the aes, geom and facet
# ggplot(barley, aes(x = year, y = yield, color = variety, group = variety)) + 
#   geom_line() +
#  facet_wrap( ~ site, nrow = 1)
```

```{r}
# Create overlapping ribbon plot from scratch
ggplot(barley, aes(x = year, y = yield, col = site, group = site, fill = site)) +
  stat_summary(fun.y = mean, geom = "line") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), geom = "ribbon", alpha = 0.1, col = NA)
```


# Case Study

## California Health Information Survey
Descriptive Statistics

```{r}
#ggplot(adult, aes(x = SRAGE_P)) +
# geom_histogram()
# diff(range(adult$SRAGE_P)) / 30
```
```{r}
#ggplot(adult, aes(x = SRAGE_P, y = BMI_P, col = factor(RBMI))) +
# geom_point(alpha = 0.4, position = position_jitter(width = 0.5))
```


```{r}
# Keep adults younger than or equal to 84
#adult <- adult[adult$SRAGE_P <= 84, ]
#
## Keep adults with BMI at least 16 and less than 52
#adult <- adult[adult$BMI_P >= 16 & adult$BMI_P < 52, ]
#
## Relabel the race variable
#adult$RACEHPR2 <- factor(adult$RACEHPR2, labels = c("Latino",
#                                                    "Asian",
#                                                    "African American",
#                                                    "White"))
#
## Relabel the BMI categories variable
#adult$RBMI <- factor(adult$RBMI, labels = c("Under-weight",
#                                            "Normal-weight",
#                                            "Over-weight",
#                                            "Obese"))
```


```{r}
# The color scale used in the plot
#BMI_fill <- scale_fill_brewer("BMI Category", palette = "Reds")
#
## Theme to fix category display in faceted plot
#fix_strips <- theme(strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.1, size = #14),
#                    strip.background = element_blank(),
#                    legend.position = "none")
#
## Histogram, add BMI_fill and customizations
#ggplot(adult, aes (x = SRAGE_P, fill= RBMI)) +
#  geom_histogram(binwidth = 1) +
#  fix_strips +
#  BMI_fill +
#  facet_grid(RBMI ~ .) +
#  theme_classic()
```


```{r}
# The initial contingency table
#DF <- as.data.frame.matrix(table(adult$SRAGE_P, adult$RBMI))
#
## Create groupSum, xmax and xmin columns
#DF$groupSum <- rowSums(DF)
#DF$xmax <- cumsum(DF$groupSum)
#DF$xmin <- DF$xmax - DF$groupSum
## The groupSum column needs to be removed; don't remove this line
#DF$groupSum <- NULL
#
## Copy row names to variable X
#DF$X <- row.names(DF)
#
## Melt the dataset
#library(reshape2)
#DF_melted <- melt(DF, id.vars = c("X", "xmin", "xmax"), variable.name = "FILL")
#
## dplyr call to calculate ymin and ymax - don't change
#library(dplyr)
#DF_melted <- DF_melted %>%
#  group_by(X) %>%
#  mutate(ymax = cumsum(value/sum(value)),
#         ymin = ymax - value/sum(value))
#
## Plot rectangles - don't change
#library(ggthemes)
#ggplot(DF_melted, aes(ymin = ymin,
#                 ymax = ymax,
#                 xmin = xmin,
#                 xmax = xmax,
#                 fill = FILL)) +
#  geom_rect(colour = "white") +
#  scale_x_continuous(expand = c(0,0)) +
#  scale_y_continuous(expand = c(0,0)) +
#  BMI_fill +
#  theme_tufte()
```



```{r}
# Perform chi.sq test (RBMI and SRAGE_P)
#results <- chisq.test(table(adult$RBMI, adult$SRAGE_P))
#
## Melt results$residuals and store as resid
#resid <- melt(results$residuals)
#
## Change names of resid
#names(resid) <- c("FILL", "X", "residual")
#
## merge the two datasets:
#DF_all <- merge(DF_melted, resid)

# Update plot command
#library(ggthemes)
#ggplot(DF_all, aes(ymin = ymin,
#                   ymax = ymax,
#                   xmin = xmin,
#                   xmax = xmax,
#                   fill = residual)) +
#  geom_rect() +
#  scale_fill_gradient2() +
#  scale_x_continuous(expand = c(0,0)) +
#  scale_y_continuous(expand = c(0,0)) +
#  theme_tufte()
```



```{r}
# Load all packages
library(ggplot2)
library(reshape2)
library(dplyr)
library(ggthemes)

## Script generalized into a function
#mosaicGG
#
## BMI described by age
#mosaicGG(adult, X = "SRAGE_P", FILL = "RBMI")
#
## Poverty described by age
#mosaicGG(adult, X = "SRAGE_P", FILL = "POVLL")
#
## mtcars: am described by cyl
#mosaicGG(mtcars, X = "cyl", FILL = "am")
#
## Vocab: vocabulary described by education
#library(carData)
#mosaicGG(Vocab, X = "education", FILL = "vocabulary")
```




































































































