---
title: "hw2prob2020(6.2)"
author: "Jefferson Ong"
date: "1/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Sec 6.2 - 2,4,6,8,10

# 2

When you purchase “1-pound bags” of carrots, you
can buy either “baby” carrots or regular carrots. We shall
compare the weights of 75 bags of each of these types of
carrots. The following table gives the weights of the bags
of baby carrots:


## (a) 

Calculate the five-number summary of each set of
weights

```{r}
baby <- c(1.03, 1.03, 1.06, 1.02, 1.03, 1.03, 1.03, 1.02, 1.03, 1.03,
1.06, 1.04, 1.05, 1.03, 1.04, 1.03, 1.05, 1.06, 1.04, 1.04,
1.03, 1.04, 1.04, 1.06, 1.03, 1.04, 1.05, 1.04, 1.04, 1.02,
1.03, 1.05, 1.05, 1.03, 1.04, 1.03, 1.04, 1.04, 1.03, 1.04,
1.03, 1.04, 1.04, 1.04, 1.05, 1.04, 1.04, 1.03, 1.03, 1.05,
1.04, 1.04, 1.05, 1.04, 1.03, 1.03, 1.05, 1.03, 1.04, 1.05,
1.04, 1.04, 1.04, 1.05, 1.03, 1.04, 1.04, 1.04, 1.04, 1.03,
1.05, 1.05, 1.05, 1.03, 1.04) 

baby <- sort(baby)
range(baby)
length(baby)
```
length() to get lowest and largest of dataset
0th percentile = 1.02
25th percentile = 1.03
50th percentile = 1.04
75th percentile = 1.04
100th percentile = 1.06


```{r}
baby[19]
baby[38]
baby[57]
p <- c(.25, .5, .75)
n <- 75
(n+1)*p
fivenum(baby)
```


```{r}
reg <- c(1.29, 1.10, 1.28, 1.29, 1.23, 1.20, 1.31, 1.25, 1.13, 1.26,
1.19, 1.33, 1.24, 1.20, 1.26, 1.24, 1.11, 1.14, 1.15, 1.15,
1.19, 1.26, 1.14, 1.20, 1.20, 1.20, 1.24, 1.25, 1.28, 1.24,
1.26, 1.20, 1.30, 1.23, 1.26, 1.16, 1.34, 1.10, 1.22, 1.27,
1.21, 1.09, 1.23, 1.03, 1.32, 1.21, 1.23, 1.34, 1.19, 1.18,
1.20, 1.20, 1.13, 1.43, 1.19, 1.05, 1.16, 1.19, 1.07, 1.21,
1.36, 1.21, 1.00, 1.23, 1.22, 1.13, 1.24, 1.10, 1.18, 1.26,
1.12, 1.10, 1.19, 1.10, 1.24)

reg <- sort(reg)
range(reg)
length(reg)
```

0th percentile = 1.00
25th percentile = 1.15
50th percentile = 1.21
75th percentile = 1.26
100th percentile = 1.43

```{r}
reg[19]
reg[38]
reg[57]
```

## (b)

On the same graph, construct box plots for each set of
weights.

```{r}
boxplot(baby, reg)
```

## (b)

If the carrots are the same price per package, which
is the better buy? Which type of carrots would you
select?

Its clear from the boxplot that it is generally better to take the regular carrots if one wants more carrots for their dollar.


# 4 

An insurance company experienced the following
mobile home losses in 10,000’s of dollars for 50 catastrophic events:

```{r}
mobile <- sort(c(1,2,2,3,3,4,4,5,5,5,
5, 6, 7, 7, 9, 9, 9, 10, 11, 12,
22, 24, 28, 29, 31, 33, 36, 38, 38, 38,
39, 41, 48, 49, 53, 55, 74, 82, 117, 134,
192, 207, 224, 225, 236 ,280, 301, 308, 351, 527))

length(mobile)
```


## (a) Find the five-number summary of the data and draw
a box-and-whisker diagram.

```{r}
range(mobile) # 0th and 100th
p = c(.25,.5,.75)
n = 50
(n + 1)*p

```

$\tilde{\pi_1} = (1 - a/b)* y_r + (a/b) * y_{r+1} $


```{r}
mobile[c(12,13,25,26,38,39)]

(1 - 3/4)*(mobile[12]) + (3/4)*mobile[13] # 25th
(1 - 1/2)*(mobile[25]) + (1/2)*mobile[26] # 50th  
(1 - 1/4)*(mobile[38]) + (1/4)*mobile[39] # 75th
fivenum(mobile)
```

```{r}

boxplot(mobile, horizontal = TRUE)
```



## (b) Calculate the IQR and the locations of the inner and
outer fences.

IQR is q4 - q2 

```{r}

90.75 - 6.75 # IQR
IQR(mobile)
```

```{r}
q1 <- 6.75
q3 <- 90.75
# inner fence
inn <- (1.5 * 84)
q1 - inn
q3 + inn

# outer fence
out <- (3 * 84)
q1 - out 
q3 + out 
```



## (c) Draw a box plot that shows the fences, suspected
outliers, and outliers.

```{r}
boxplot(mobile, horizontal = T)
```


## (d) Describe the distribution of losses. 

heavily skewed right

# 6

In the casino game roulette, if a player bets $1 on
red (or on black or on odd or on even), the probability
of winning $1 is 18/38 and the probability of losing $1 is
20/38. Suppose that a player begins with $5 and makes
successive $1 bets. Let Y equal the player’s maximum capital before losing the $5. One hundred observations of
Y were simulated on a computer, yielding the following
data:

```{r}
roulette <- sort(c(25, 9, 5, 5, 5, 9, 6, 5, 15, 45,
55, 6, 5, 6 ,24, 21, 16 ,5 ,8 ,7,
7 ,5, 5, 35, 13, 9, 5, 18, 6, 10,
19, 16, 21, 8 ,13, 5, 9, 10 ,10, 6,
23, 8 ,5, 10, 15, 7, 5, 5 ,24, 9,
11, 34, 12, 11 ,17 ,11, 16, 5, 15, 5,
12, 6, 5, 5 ,7 ,6, 17, 20 ,7 ,8,
8 ,6 ,10, 11, 6, 7 ,5 ,12 ,11, 18,
6 ,21, 6, 5 ,24, 7, 16 ,21, 23, 15,
11, 8, 6 ,8 ,14, 11 ,6, 9, 6 ,10))


range(roulette)
length(roulette)

```


## (a) Construct an ordered stem-and-leaf display.

```{r}
stem(roulette, scale = 1, width = 80, atom = 1e-08)
```



## (b) Find the five-number summary of the data and draw
a box-and-whisker diagram.

```{r}

range(roulette) # 0th and 100th
p = c(.25,.5,.75)
n = 100
(n + 1)*p

roulette[c(25,26,50,51,75,76)]

(1 - 1/4)*(roulette[25]) + (1/4)*roulette[26] # 25th  
(1 - 1/2)*(roulette[50]) + (1/2)*roulette[51] # 50th  
(1 - 3/4)*(roulette[75]) + (3/4)*roulette[76] # 75th
fivenum(roulette)
```

five sum 5, 6, 9, 15, 55




## (c) Calculate the IQR and the locations of the inner and
outer fences.


```{r}
q1 = 6
q2 = 9
q3 = 15
q3 - q1 # IQR
IQR(roulette)
iqr = 9
inn <- (1.5 * 9)
q1 - inn
q3 + inn

# outer fence
out <- (3 * 9)
q1 - out 
q3 + out 
```
-7.5 to 28.5 for inner

-21 to 42 for outer

## (d) Draw a box plot that shows the fences, suspected
outliers, and outliers.

```{r}
boxplot(roulette)
```
The suspected outliers are the tails of the boxplot while the outleiers are past the tail up to 42 and below the graph to -21




## (e) Find the 90th percentile.

```{r}
p = .9
n = 100
(n + 1)*p

(1 - 9/10)*(roulette[90]) + (9/10)*roulette[91] # 90th 
```
22.8



# 8

. The weights (in grams) of 25 indicator housings
used on gauges are as follows:


```{r}
house <- sort(c(102.0, 106.3, 106.6 ,108.8 ,107.7,
106.1 ,105.9 ,106.7 ,106.8, 110.2,
101.7 ,106.6 ,106.3 ,110.2, 109.9,
102.0 ,105.8 ,109.1 ,106.7, 107.3,
102.0 ,106.8 ,110.0 ,107.9 ,109.3))


range(house)
length(house)
```

## (a) Construct an ordered stem-and-leaf display, using
integers as the stems and tenths as the leaves.

```{r}
stem(house, scale = 1, width = 80, atom = 1e-08)
```


## (b) Find the five-number summary of the data and draw
a box plot.

```{r}
range(house) # 0th and 100th
p = c(.25,.5,.75)
n = 25
(n + 1)*p

house[c(6,7,13,19,20)]

(1 - 1/2)*(house[6]) + (1/2)*house[7] # 25th  
#(1 - 1/2)*(house[50]) + (1/2)*house[51] 
house[13]# 50th  
(1 - 1/2)*(house[19]) + (1/2)*house[20] # 75th
fivenum(house)
```

101.7, 106, 106.7, 108.95, 110.2

```{r}
boxplot(house)
```


## (c) Are there any suspected outliers? Are there any outliers?

```{r}
q1 = 106
q2 = 106.7
q3 = 108.95
q3 - q1 # IQR
IQR(house)
iqr =  2.95
inn <- (1.5 *  2.95)
q1 - inn
q3 + inn

# outer fence
out <- (3 *  2.95)
q1 - out 
q3 + out 
```
 101.575 - 113.375 suspected
 
 97.15 -  117.8 outliers
 
 Yes there are a handful of elements past 101.575 and some above q3 for suspected outliers
 
 There is no elements above 113.375 or below 101.575 so there are no clear outliers



# 10


In Exercise 6.1-7, lead concentrations near the
San Diego Freeway in 1976 are given. During the fall
of 1977, the weekday afternoon lead concentrations (in
μg/m3) at the measurement station near the San Diego
Freeway in Los Angeles were as follows:

```{r}
lead <- sort(c(9.5 ,10.7, 8.3 ,9.8 ,9.1 ,9.4 ,9.6 ,11.9 ,9.5 ,12.6, 10.5,
8.9 ,11.4, 12.0,12.4, 9.9,10.9,12.3, 11.0, 9.2, 9.3, 9.3,
10.5, 9.4, 9.4 ,8.2 ,10.4, 9.3, 8.7, 9.8 ,9.1 ,2.9 ,9.8,
5.7 ,8.2 ,8.1  ,8.8 ,9.7 ,8.1 ,8.8 ,10.3 ,8.6 ,10.2, 9.4,
14.8, 9.9, 9.3 ,8.2 ,9.9 ,11.6, 8.7, 5.0 ,9.9 ,6.3 ,6.5,
10.2, 8.8, 8.0 ,8.7 ,8.9 ,6.8 ,6.6 ,7.3  ,16.7))

range(lead)
length(lead)
lead
```


## (a) Construct a frequency distribution and display the
 results in the form of a histogram. Is this distribution
symmetric?

#install.package("summarytools")

frequencies: 1  2  6 37 13  3  1  1

```{r}
#library(summarytools)
#summarytools::freq(lead, order = "freq")
tomato <- hist(lead)
tomato
summary(tomato)
```

Yes fairly symmetric 

## (b) Calculate the sample mean and sample standard deviation.

$\bar{x} = \frac{1}{n}*\sum\limits_{i=1}^n x_i$

$S^2 = \frac{1}{n-1}*\sum\limits_{i=1}^n (x_i-\bar{x})^2$

```{r}
n = 64
xbar <- sum(lead)/n
potato <- (lead - xbar)^2
var <- sum(potato) / (n - 1)
# var <- (sum(lead - xbar)^2)/(n - 1)
sd <- sqrt(var)

xbar
sd
mean(lead)
sd(lead)
```





## (c) Locate x, x ± s on your histogram. How many observations lie within one standard deviation of the mean?
 How many lie within two standard deviations of the
 mean?
 
```{r}
xbar + sd
xbar - sd

xbar + 2*sd
xbar - 2*sd

```

11.50243 - 7.341317 one sd: obs: 37 + 6 + 13 = 56

13.58299 - 5.26076 two sd: obs: 37 + 6 + 13 + 2 + 3 = 61



```{r}
hist(lead,labels = TRUE ) 

# freq 1  2  6 37 13  3  1  1
```

 
## (d) Using the data from Exercise 6.1-7 and the data from
 this exercise, construct a back-to-back stem-and-leaf
 diagram with integer stems in the center and the
 leaves for 1976 going to the left and those for 1977
 going to the right.
 
 
```{r}
# 6.1 data
free <- c(6.7 ,5.4 ,5.2 ,6.0 ,8.7 ,6.0 ,6.4 ,8.3 ,5.3 ,5.9 ,7.6,
5.0 ,6.9 ,6.8 ,4.9 ,6.3 ,5.0 ,6.0 ,7.2 ,8.0 ,8.1 ,7.2,
10.9, 9.2, 8.6, 6.2, 6.1, 6.5, 7.8, 6.2, 8.5, 6.4, 8.1,
2.1 ,6.1 ,6.5 ,7.9 ,14.1, 9.5,10.6, 8.4, 8.3, 5.9, 6.0,
6.4 ,3.9 ,9.9 ,7.6 ,6.8 ,8.6 ,8.5 ,11.2, 7.0, 7.1, 6.0,
9.0 ,10.1, 8.0, 6.8, 7.3, 9.7, 9.3, 3.2, 6.4)

library(aplpack)
stem.leaf.backback(free,lead)
```
 
 
## (e) Construct box-and-whisker displays of both sets of
 data on the same graph.
 
 
```{r}
boxplot(free, lead, names = c('free', 'lead'))
```
 
 
## (f) Use your numerical and graphical results to interpret
 what you see.

The data for 1977(lead) has a higher five number summary and smaller IQR but also more outliers compare to its 1976(free) counterpart



































