---
title: 'HW: Foundations of Inference'
author: 'Jefferson Ong'
date: '`r format(Sys.time(), "%B %d, %Y @ %I:%M %p")`'
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: false
    theme: yeti
    highlight: textmate
---

**DataCamp Course Link:** [Foundations of Inference](https://www.datacamp.com/courses/foundations-of-inference)

***

> **INSTRUCTIONS:** Complete the exercises below, which are based on the activities in the DataCamp course. In each case, perform a permutation hypothesis test with 10^4^ reps. Include in your output: (1) the null and alternative hypotheses; (2) proportion summaries for both groups; (3) difference in proportions (test statistic); (4) a plot of the permutation results including a vertical line representing the test statistic; (5) p-value; and (6) Whether you reject the null hypothesis and what you conclude from the data. You should use the [R Style Guide](http://adv-r.had.co.nz/Style.html) for your document. Load all packages in the code chunk provided. The datasets have been loaded for you.

```{r}
library(tidyverse)
library(dplyr)
library(infer)
```


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment = "")
```

```{r packages&datasets}
disc <- tibble(promote = rep(c("promoted", "not_promoted"), c(35, 13)),
               sex = rep(rep(c("male", "female"), 2), c(21, 14, 3, 10)))

disc_big <- readRDS(gzcon(url("https://assets.datacamp.com/production/repositories/538/datasets/f03da8fc4a2ae50a3ddf775324f4df90c96f7f26/disc_big.rds")))

opportunity <- tibble(decision = rep(c("buyDVD", "nobuyDVD"), c(97, 53)),
                      group = rep(rep(c("control", "treatment"), 2), c(56, 41, 19, 34)))
```

Different sample size, changes the p-value since one is 10x the sample size.

Opportunity is buy or no buy, whether people will save money. 

Research question(datacamp) whether women are discriminated or not. Are men more likely to get promoted than women? In statistic, is the proportion of men greater than women in being promoted?

NULL hypothesis is that nothing is going on. There's nothing going on, everything is normal. Copy down the H_o and H_a. Formal logical process, that starts with 2 binary statement that you can only have one at a time. Court system that we assume you're innocent before guilty. 



### <span style="font-size: 100%; color: #2780E3;">&#9312;</span> Hypothesis Test 1: `disc` Data

H0: gender and promotion are unrelated variables.

HA: men are more likely to be promoted.


```{r}
# Create a contingency table summarizing the data
disc %>%
  # Count the rows by sex, promote
  count(sex, promote)

# Find proportion of each sex who were promoted
disc %>%
  # Group by sex
  group_by(sex) %>%
  # Calculate proportion promoted summary stat
  summarise(promoted_prop = mean(promote == "promoted")) 

table(disc)
```

```{r}
# Replicate the entire data frame, permuting the promote variable
disc_perm <- disc %>%
  specify(promote ~ sex, success = "promoted") %>%
  hypothesize(null = "independence") %>%
  generate(reps = 5, type = "permute")

disc_perm %>%
  # Group by replicate
  group_by(replicate) %>%
  # Count per group
  count()

disc_perm %>%
  # Calculate difference in proportion, male then female
  calculate(stat = "diff in props", order = c("male", "female"))
```
```{r}
# Calculate the observed difference in promotion rate
# diff_orig <- disc %>%
#   # Group by sex
#   group_by(sex) %>%
#   # Summarize to calculate fraction promoted
#   summarize(prop_prom = mean(promote == "promoted")) %>%
#   # Summarize to calculate difference
#   summarize(stat = diff(prop_prom)) %>% 
#   pull()
#   
# See the result
# diff_orig

# Create data frame of permuted differences in promotion rates
disc_perm <- disc %>%
  # Specify promote vs. sex
  specify(promote~sex, success = "promoted") %>%
  # Set null hypothesis as independence
  hypothesize(null = "independence") %>%
  # Generate 1000 permutations
  generate(reps = 1000, type = "permute") %>%
  # Calculate difference in proportions
  calculate(stat = "diff in props", order = c("male", "female"))

# From previous steps
 diff_orig <- disc %>%
   group_by(sex) %>%
    summarise(prop_prom = mean(promote == "promoted")) %>%
    summarise(stat = diff(prop_prom)) %>% 
    pull()
 disc_perm <- disc %>%
   specify(promote ~ sex, success = "promoted") %>%
   hypothesize(null = "independence") %>%
   generate(reps = 1000, type = "permute") %>%
   calculate(stat = "diff in props", order = c("male", "female"))
   
 # Using permutation data, plot stat
 ggplot(disc_perm, aes(x = stat)) + 
   # Add a histogram layer
   geom_histogram(binwidth = 0.01) +
   # Add a vertical line at diff_orig
   geom_vline(aes(xintercept = diff_orig), color = "red")
```

```{r}
# Visualize and calculate the p-value for the original dataset
disc_perm %>%
  visualize(obs_stat = diff_orig, direction = "greater")

disc_perm %>%
  get_p_value(obs_stat = diff_orig, direction = "greater")

```

Since the p-value is less than 5% we reject the null hypothesis and find that there is a relationship in this data with being promoted and gender. 




### <span style="font-size: 100%; color: #2780E3;">&#9313;</span> Hypothesis Test 2: `disc_large` Dataset

H0: gender and promotion are unrelated variables.

HA: men are more likely to be promoted.

```{r}
disc_big %>%
  # Count the rows by sex, promote
  count(sex, promote)

# Find proportion of each sex who were promoted
disc_big %>%
  # Group by sex
  group_by(sex) %>%
  # Calculate proportion promoted summary stat
  summarise(promoted_prop = mean(promote == "promoted")) 
```

```{r}
# Replicate the entire data frame, permuting the promote variable
disc_permW <- disc_big %>%
  specify(promote ~ sex, success = "promoted") %>%
  hypothesize(null = "independence") %>%
  generate(reps = 5, type = "permute")

disc_permW %>%
  # Group by replicate
  group_by(replicate) %>%
  # Count per group
  count()

disc_permW %>%
  # Calculate difference in proportion, male then female
  calculate(stat = "diff in props", order = c("male", "female"))
```

```{r}
# Create data frame of permuted differences in promotion rates
disc_permW <- disc_big %>%
  # Specify promote vs. sex
  specify(promote~sex, success = "promoted") %>%
  # Set null hypothesis as independence
  hypothesize(null = "independence") %>%
  # Generate 1000 permutations
  generate(reps = 1000, type = "permute") %>%
  # Calculate difference in proportions
  calculate(stat = "diff in props", order = c("male", "female"))

# From previous steps
 diff_origW <- disc_big %>%
   group_by(sex) %>%
    summarise(prop_prom = mean(promote == "promoted")) %>%
    summarise(stat = diff(prop_prom)) %>% 
    pull()
 disc_permW <- disc_big %>%
   specify(promote ~ sex, success = "promoted") %>%
   hypothesize(null = "independence") %>%
   generate(reps = 1000, type = "permute") %>%
   calculate(stat = "diff in props", order = c("male", "female"))
   
 # Using permutation data, plot stat
 ggplot(disc_permW, aes(x = stat)) + 
   # Add a histogram layer
   geom_histogram(binwidth = 0.01) +
   # Add a vertical line at diff_orig
   geom_vline(aes(xintercept = diff_origW), color = "red")
```

```{r}
disc_permW %>%
  visualize(obs_stat = diff_origW, direction = "greater")

disc_permW %>%
  get_p_value(obs_stat = diff_origW, direction = "greater")

```

We reject the null hypothesis again. Since the p-value is less than 5% we reject the null hypothesis and find that there is a relationship in this data with being promoted and gender. 




### <span style="font-size: 100%; color: #2780E3;">&#9314;</span> Hypothesis Test 3: `opportunity` Dataset

H0: Reminding students will have no impact on their spending decisions. 

HA: Reminding students will reduce the chance they continue with a purchase.

```{r}
table(opportunity)
```


```{r}
# Tabulate the data
opportunity %>%
  count(decision, group)

# Find the proportion who bought the DVD in each group
opportunity %>%
  group_by(group) %>%
  summarize(buy_prop = mean(decision == "buyDVD"))
```

```{r}
# Calculate the observed difference in purchase rate
diff_obs <- opportunity %>%
  # Group by group
  group_by(group) %>%
  # Calculate proportion deciding to buy a DVD
  summarise(prop_buy = mean(decision == "buyDVD")) %>%
  # Calculate difference between groups
  summarise(stat = diff(prop_buy)) %>% 
  pull()
    
# Review the result
diff_obs

# Create data frame of permuted differences in purchase rates
opp_perm <- opportunity %>%
  # Specify decision vs. group, where success is buying a DVD
  specify(decision ~ group, success = "buyDVD") %>%
  # Set the null hypothesis to independence
  hypothesize(null = "independence") %>%
  # Generate 1000 reps of type permute
  generate(reps = 1000, type = "permute") %>%
  # Calculate the summary stat difference in proportions
  calculate(stat = "diff in props", order = c("treatment", "control"))
    
# Review the result
opp_perm
  
# Using the permuation data, plot stat
ggplot(opp_perm, aes(x = stat)) + 
  # Add a histogram layer with binwidth 0.005
  geom_histogram(binwidth = 0.005) +
  # Add a vline layer with intercept diff_obs
  geom_vline(aes(xintercept = diff_obs), color = "red")
```
```{r}
# Visualize the statistic 
opp_perm %>%
  visualize(obs_stat = diff_orig, direction = "less")

# Calculate the p-value using `get_p_value`
opp_perm %>%
  get_p_value(obs_stat = diff_orig, direction = "less")

# Calculate the p-value using `summarize`
opp_perm %>%
  summarise(p_value = mean(stat <= diff_orig))
```

P value is greater than 5% so we do not reject the null hypothesis and that reminding students has no effect on buying.

## All the conclusions contradicted what was said in class and in datacamp so i'm not completely sure whats going on. However I just followed whether to reject or not base on the p-value. 



1: Hypothesis

2: Descriptive summary

3: Diff. in prop(test stat)

4: simulation in plot

5: p - value

diff_orig can be called observed difference in the data (2). Then show the actualy difference in proportion (3)

"randomzing gender discrimination" (3), (4)

---

What prop. of simulated results are > diff_orig? (finding p -value)

disc_perm is my sim results

diff_orig is actual data

mean(disc_perm > diff_orig), get_p_value() does this




Uses infer package

***
```{r}
sessionInfo()
```