---
title: "Inference Proficiency Problem"
subtitle: "Yankees Salary $$$ vs. Everyone Else in 2002"
author: "Jefferson Ong"
date: '`r format(Sys.time(), "%B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment = "")
```

Incorrect bootstrapping simulation. It's a separate sim from the permutation test. // The t.test() needs a one-sided p-value.  // Why did you calculate the permutation p-value twise. The first was a correct upper-tail p-value, the one you did later is an incorrect two-sided one. You also need to interpret using your percentile interval. // What are the assumptions of the t-test?

***
```{r}
# Load any packages and datasets you need here.
library(tidyverse)
library(readr)

mlb2002 <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/mlb2002.csv")
```

#### The Problem

In 2002, the New York Yankees had the highest average team salary among all Major League Baseball (MLB) teams. However, was it statistically larger than the average salary of the rest of the teams in the league combined? Use the data in the file [mlb2002.csv](https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/mlb2002.csv) to test this hypothesis. (*Technically*, this data could be a population, but we can also consider it a sample from the possible salary structures for that season and thus prone to some sampling variability.)


#### The Analysis

Add a new variable to the dataset that identifies players according to whether they played for the New York Yankees (NYY) or one of the other teams (OTH). Then `summarize` n, mean, and standard deviation for salary in each group (NYY versus OTH). Also create a boxplot to compare the NYY and OTH groups. 
```{r}
str(mlb2002)
list(mlb2002)
```

```{r}
mlb2002 <- mlb2002 %>%
  mutate(teams = ifelse(team == "New York Yankees", "NYY", "OTH"))

mlb2002 %>%
  group_by(teams) %>%
  summarize(n     = n(),
            xbar  = mean(salary, na.rm = TRUE),
            stdev = sd(salary, na.rm = TRUE)
            )


ggplot(mlb2002, aes(x = teams, y = salary)) +
  geom_boxplot()
```


Conduct a permutation resampling test to determine whether the Yankees' mean was statistically larger than the mean of the rest of the league combined. 

```{r}
#
salaryG <- mlb2002 %>% pull(salary) # Pulls out with no group info from the buckets. 

n_all <- length(salaryG)

salary_NYY <- mlb2002 %>% filter(teams == "NYY") %>% pull(salary) #makes bucket

salary_OTH <- mlb2002 %>% filter(teams == "OTH") %>% pull(salary) 
#makes other bucket
observed_diff <- mean(salary_NYY) - mean(salary_OTH)

n_NYY <- length(salary_NYY)
  
n_OTH <- length(salary_OTH)

N <- 10^4 - 1  

random_diffs <- numeric(N) 

for (i in 1:N)
{
  index <- sample(n_all, size = n_NYY, replace = FALSE) 
  random_diffs[i] <- mean(salaryG[index]) - mean(salaryG[-index])
}

ggplot() + 
  geom_histogram(aes(random_diffs), fill = "light blue") + 
  geom_vline(xintercept = observed_diff, colour = "red") +
 xlab("xbar1-xbar2")

(sum(random_diffs >= abs(observed_diff)) + 1) / (N + 1)
```

Compute a 95% bootstrap percentile interval to estimate the difference in the two group means.

```{r}
#
tgt <-  mlb2002 %>% filter(teams == "NYY") %>% pull(salary)
tgt_n = length(tgt)
wmt <- mlb2002 %>% filter(teams == "OTH") %>% pull(salary) 
wmt_n = length(wmt)

obs_diffs <- mean(tgt) - mean(wmt)

N <- 10^4
boot_diffs <- numeric(N)

for (i in 1:N) {
  boot_diffs[i] <- mean(sample(tgt, tgt_n, replace = TRUE)) - mean(sample(wmt, wmt_n, replace = TRUE))
}
quantile(boot_diffs, c(0.025, 0.975))

ggplot(NULL, aes(x = boot_diffs)) + geom_density()
ggplot(NULL, aes(sample = boot_diffs)) + 
  stat_qq() +
  stat_qq_line(color = "red")
```

A: The bootstrap percentile interval does not contains zero, so there is evidence of a difference.


Conduct a classical two-sample t-test for the same hypotheses using the `t.test` function. (Be sure to set the appropriate H~a~.)

```{r}
#
t.test(salary ~ teams, data = mlb2002, alternative = "greater")
```

Using the results of the permutation resampling test, can we conclude at the &alpha; = 0.05 level of significance that the Yankee's mean salary was statistically larger than mean salary of the rest of the teams combined? If so, how large is the difference, based on the bootstrap percentile interval? Is the difference large enough to have practical significance, in context?

**ANSWER:** 

```{r}
p_value_lower <- (sum(boot_diffs <= -abs(obs_diffs)) + 1) / (N + 1)
# p_value_upper <- (sum(random_diffs >= abs(observed_diff)) + 1) / (N + 1)
# p_value_2side <- p_value_lower + p_value_upper
p_value_lower
sprintf("The observed difference in sample means between NYY and OTH is %0.4f.", obs_diffs)
```

A: Yes, it is about double of what the average player in OTH would make compare to NYY players. 


Given the assumptions of the classical two-sample t-test, why might the p-value from the classical two-sample t-test be unreliable? Use evidence from your descriptive analysis.

**ANSWER:**

A:  Looking at the bootstrap distribution where the distribution is not normal base on the qqplot. The p-value from the t-test is unreliable since it is not a normal distribution. 


***
```{r}
sessionInfo()
```
