---
title: "Intermediate R"
author: "Jefferson Ong"
date: "4/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Conditionals and Control Flow

Equality ==

Inequality !=


< and > 


<= and >=

AND operator &
● OR operator |
● NOT operator !

```{r}
# The social data has been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)
views <- matrix(c(linkedin, facebook), nrow = 2, byrow = TRUE)

# When does views equal 13?
views == 13

# When is views less than or equal to 14?
views <= 14

# The linkedin and last variable are already defined for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
last <- tail(linkedin, 1)

# Is last under 5 or above 10?
last < 5 | last > 10

# Is last between 15 (exclusive) and 20 (inclusive)?
last > 15 & last <= 20
```


```{r}
# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Examine the if statement for medium
if (medium == "LinkedIn") {
  print("Showing LinkedIn information")
}

# Write the if statement for num_views
if (num_views > 15) {
  print("You're popular!")
}
```

```{r}
# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Control structure for medium
if (medium == "LinkedIn") {
  print("Showing LinkedIn information")
} else if (medium == "Facebook") {
  # Add code to print correct string when condition is TRUE
  print("Showing Facebook information")
} else {
  print("Unknown medium")
}

# Control structure for num_views
if (num_views > 15) {
  print("You're popular!")
} else if (num_views <= 15 & num_views > 10) {
  # Add code to print correct string when condition is TRUE
  print("Your number of views is average")
} else {
  print("Try to be more visible!")
}
```

```{r}
# Variables related to your last day of recordings
li <- 15
fb <- 9

# Code the control-flow construct
if (li >= 15 & fb >= 15) {
  sms <- 2 * (li + fb)
} else if (li < 10 & fb < 10) {
  sms <- 0.5 * (li + fb)
} else {
  sms <- li + fb
}

# Print the resulting sms to the console
sms
```




# Loops




```{r}
# Initialize the speed variable
speed <- 88

while (speed > 30) {
  print(paste("Your speed is", speed))
  
  # Break the while loop when speed exceeds 80
  if (speed > 80) {
    break
  }
  
  if (speed > 48) {
    print("Slow down big time!")
    speed <- speed - 11
  } else {
    print("Slow down!")
    speed <- speed - 6
  }
}
```

```{r}
# Initialize i as 1 
i <- 1

# Code the while loop
while (i <= 10) {
  print(3 * i)
  if ( (3 * i) %% 8 == 0) {
    break
  }
  i <- i + 1
}
```

```{r}
# The linkedin vector has already been defined for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)

# Adapt/extend the for loop
for (li in linkedin) {
  if (li > 10) {
    print("You're popular!")
  } else {
    print("Be more visible!")
  }
  
  # Add if statement with break
  if (li > 16) {
    print("This is ridiculous, I'm outta here!")
    break
  }
  
  # Add if statement with next
  if (li < 5) {
    print("This is too embarrassing!")
    next
  }
  
  print(li)
}
```

```{r}
# Pre-defined variables
rquote <- "r's internals are irrefutably intriguing"
chars <- strsplit(rquote, split = "")[[1]]

# Initialize rcount
rcount <- 0

# Finish the for loop
for (char in chars) {
  if (char == "r") {
    rcount <- rcount + 1
  }
  if (char == "u") {
    break
  }
}

# Print out rcount
rcount
```





# Functions


● Create a list: list()
● Display a variable: print()

Argument names: x, na.rm
● na.rm = FALSE
● sd(values) works?


```{r}
linkedin <- c(16, 9, 13, 5, NA, 17, 14)
facebook <- c(17, NA, 5, 16, 8, 13, 14)

# Basic average of linkedin
mean(linkedin)

# Advanced average of linkedin
mean(linkedin, na.rm = TRUE)
```
```{r}
# Create a function pow_two()
pow_two <- function(x) {
  x ^ 2
}

# Use the function
pow_two(12)

# Create a function sum_abs()
sum_abs <- function(x, y) {
  abs(x) + abs(y)
}

# Use the function
sum_abs(-2, 3)
```

```{r}
# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# The interpret() can be used inside interpret_all()
interpret <- function(num_views) {
  if (num_views > 15) {
    print("You're popular!")
    return(num_views)
  } else {
    print("Try to be more visible!")
    return(0)
  }
}

# Define the interpret_all() function
# views: vector with data to interpret
# return_sum: return total number of views on popular days?
interpret_all <- function(views, return_sum = TRUE) {
  count <- 0

  for (v in views) {
    count <- count + interpret(v)
  }

  if (return_sum) {
    return(count)
  } else {
    return(NULL)
  }
}

# Call the interpret_all() function on both linkedin and facebook
interpret_all(linkedin)
interpret_all(facebook)
```












# The apply family

sapply(), vapply(), lapply()

sort()
 print()
 identical() 
 
 
● lapply()
apply function over list or vector
output = list
● sapply()
apply function over list or vector
try to simplify list to array
● vapply()
apply function over list or vector
explicitly specify output format
 
 
 
 lapply()
● Apply function over list or vector
● Function can return R objects of different classes
● List necessary to store heterogeneous content
● However, o"en homogeneous content

```{r}
# The vector pioneers has already been created for you
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")

# Split names from birth year
split_math <- strsplit(pioneers, split = ":")

# Convert to lowercase strings: split_low
split_low <- lapply(split_math, tolower)

# Take a look at the structure of split_low
str(split_low)
```

```{r}
# Code from previous exercise:
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")
split <- strsplit(pioneers, split = ":")
split_low <- lapply(split, tolower)

# Write function select_first()
select_first <- function(x) {
  x[1]
}

# Apply select_first() over split_low: names
names <- lapply(split_low, select_first)

# Write function select_second()
select_second <- function(x) {
  x[2]
}

# Apply select_second() over split_low: years
years <- lapply(split_low, select_second)

# Generic select function
select_el <- function(x, index) {
  x[index]
}

# Use lapply() twice on split_low: names and years
names <- lapply(split_low, select_el, index = 1)
years <- lapply(split_low, select_el, index = 2)
```

```{r}
# # temp has already been defined in the workspace
# 
# # Use lapply() to find each day's minimum temperature
# lapply(temp, min)
# 
# # Use sapply() to find each day's minimum temperature
# sapply(temp, min)
# 
# # Use lapply() to find each day's maximum temperature
# lapply(temp, max)
# 
# # Use sapply() to find each day's maximum temperature
# sapply(temp, max)
```
```{r}
# # Definition of below_zero()
# below_zero <- function(x) {
#   return(x[x < 0])
# }
# 
# # Apply below_zero over temp using sapply(): freezing_s
# freezing_s <- sapply(temp, below_zero)
# 
# # Apply below_zero over temp using lapply(): freezing_l
# freezing_l <- lapply(temp, below_zero)
# 
# # Are freezing_s and freezing_l identical?
# identical(freezing_s, freezing_l)
```



```{r}
# 
# # Definition of basics()
# basics <- function(x) {
#   c(min = min(x), mean = mean(x), max = max(x))
# }
# 
# # Apply basics() over temp using vapply()
# vapply(temp, basics, numeric(3))
```














# Utilities

```{r}
# The errors vector has already been defined for you
errors <- c(1.9, -2.6, 4.0, -9.5, -3.4, 7.3)

# Sum of absolute rounded values of errors
sum(abs(round(errors)))
```

```{r}
# Don't edit these two lines
vec1 <- c(1.5, 2.5, 8.4, 3.7, 6.3)
vec2 <- rev(vec1)

# Fix the error
mean(c(abs(vec1), abs(vec2)))
```

```{r}
# The linkedin and facebook lists have already been created for you
linkedin <- list(16, 9, 13, 5, 2, 17, 14)
facebook <- list(17, 7, 5, 16, 8, 13, 14)

# Convert linkedin and facebook to a vector: li_vec and fb_vec
li_vec <- unlist(linkedin)
fb_vec <- unlist(facebook)

# Append fb_vec to li_vec: social_vec
social_vec <- append(li_vec, fb_vec)

# Sort social_vec
sort(social_vec, decreasing = TRUE)
```

```{r}
# Fix me
rep(seq(1, 7, by = 2), times = 7)

# Create first sequence: seq1
seq1 <- seq(1, 500, by = 3)

# Create second sequence: seq2
seq2 <- seq(1200, 900, by = -7)

# Calculate total sum of the sequences
sum(seq1) + sum(seq2)
```

```{r}
# Get the current date: today
today <- Sys.Date()

# See what today looks like under the hood
unclass(today)

# Get the current time: now
now <- Sys.time()

# See what now looks like under the hood
unclass(now)
```

```{r}
# Definition of character strings representing times
str1 <- "May 23, '96 hours:23 minutes:01 seconds:45"
str2 <- "2012-3-12 14:23:08"

# Convert the strings to POSIXct objects: time1, time2
time1 <- as.POSIXct(str1, format = "%B %d, '%y hours:%H minutes:%M seconds:%S")
time2 <- as.POSIXct(str2)

# Convert times to formatted strings
format(time1, "%M")
format(time2, "%I:%M %p")
```






































