---
title: 'Introduction to R Assignment'
author: 'Author: Jefferson Ong'
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

**Course Link:** [Introduction to R](https://www.datacamp.com/courses/free-introduction-to-r)

***

**Instructions**

Replicate all of the activities from the DataCamp course _Introduction to R_, including code and comments. In cases where you are told that an object is "available in your workspace" for a given activity you will have to create it for yourself here. Those sections are indicated below. You are welcome to include additional text or comments to help you recall how the code is used.

***

### Intro to Basics

How it works

```{r}
# Calculate 3 + 4
3 + 5

# Calculate 
6 + 12
```


Arithmetic with R

```{r}
# An addition
5 + 5 

# A subtraction
5 - 5 

# A multiplication
3 * 5

 # A division
(5 + 5) / 2 

# Exponentiation
2^5

# Modulo
28 %% 6
```


Variable assignment

```{r}
# Assign the value 42 to x
x <- 42

# Print out the value of the variable x
x
```


Variable assignment (2)

```{r}
# Assign the value 5 to the variable my_apples
my_apples <- 5


# Print out the value of the variable my_apples
my_apples
```


Variable assignment (3)

```{r}
# Assign a value to the variables my_apples and my_oranges

my_apples <- 5

my_oranges <- 6
# Add these two variables together
my_fruit <- my_oranges + my_apples

# Create the variable my_fruit
my_fruit
```


Apples and oranges

```{r}
# Assign a value to the variable my_apples

my_apples <- 5
# Fix the assignment of my_oranges
my_oranges <- 6

# Create the variable my_fruit and print it out
my_fruit <- my_oranges + my_apples

my_fruit
```


Basic data types in R

```{r}
# Change my_numeric to be 42

my_numeric <- 42
# Change my_character to be "universe"
my_character <- "universe"

# Change my_logical to be FALSE
my_logical <- FALSE
```


What's that data type?

```{r}
# Declare variables of different types
my_numeric <- 42
my_character <- "universe"
my_logical <- FALSE 

# Check class of my_numeric
class(my_numeric)

# Check class of my_character
class(my_character)

# Check class of my_logical
class(my_logical)
```


### Vectors

Create a vector

```{r}
# Define the variable vegas
vegas <- "Go!"
```


Create a vector (2)

```{r}

numeric_vector <- c(1, 10, 49)
character_vector <- c("a", "b", "c")
# Complete the code for boolean_vector
boolean_vector <- c(TRUE, FALSE, TRUE)
```


Create a vector (3)

```{r}
# Poker winnings from Monday to Friday

poker_vector <- c(140, -50, 20, -120, 240)
# Roulette winnings from Monday to Friday
roulette_vector <-  c(-24, -50, 100, -350, 10)
```


Naming a vector

```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday

roulette_vector <- c(-24, -50, 100, -350, 10)
# Assign days as names of poker_vector

names(poker_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
# Assign days as names of roulette_vectors
names(roulette_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
```


Naming a vector (2)

```{r}
# Poker winnings from Monday to Friday

poker_vector <- c(140, -50, 20, -120, 240)
# Roulette winnings from Monday to Friday

roulette_vector <- c(-24, -50, 100, -350, 10)
# The variable days_vector
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
 
# Assign the names of the day to roulette_vector and poker_vector
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector
```


Calculating total winnings

```{r}
A_vector <- c(1, 2, 3)
B_vector <- c(4, 5, 6)

# Take the sum of A_vector and B_vector
total_vector <- A_vector + B_vector
  
# Print out total_vector
total_vector
```


Calculating total winnings (2)

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Assign to total_daily how much you won/lost on each day
total_daily <- poker_vector + roulette_vector
```


Calculating total winnings (3)

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Total winnings with poker
total_poker <- sum(poker_vector)

# Total winnings with roulette
total_roulette <-  sum(roulette_vector)

# Total winnings overall
total_week <- total_roulette + total_poker

# Print out total_week
total_week
```


Comparing total winnings

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Calculate total gains for poker and roulette
total_poker <- sum(poker_vector)
total_roulette <- sum(roulette_vector)


# Check if you realized higher total gains in poker than in roulette 
total_poker > total_roulette
```


Vector selection: the good times

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Define a new variable based on a selection
poker_wednesday <- poker_vector[3]
poker_wednesday
```


Vector selection: the good times (2)

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Define a new variable based on a selection
poker_midweek <- poker_vector[c(2, 3, 4)]
poker_midweek
```


Vector selection: the good times (3)

```{r}
# Poker and roulette winnings from Monday to Friday:

poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector
# Define a new variable based on a selection
roulette_selection_vector <- roulette_vector[c(2:5)]
roulette_selection_vector
```


Vector selection: the good times (4)

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Select poker results for Monday, Tuesday and Wednesday
poker_start <- poker_vector[c("Monday", "Tuesday", "Wednesday")]
  
  
# Calculate the average of the elements in poker_start
mean(poker_start)
```


Selection by comparison - Step 1

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector


# Which days did you make money on poker?
selection_vector <- poker_vector > 0
  
  
# Print out selection_vector
selection_vector
```


Selection by comparison - Step 2

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector


# Which days did you make money on poker?
selection_vector <- poker_vector > 0

# Select from poker_vector these days
poker_winning_days <- poker_vector[selection_vector]
poker_winning_days
```


Advanced selection

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Which days did you make money on roulette?
selection_vector <- roulette_vector > 0

# Select from roulette_vector these days
roulette_winning_days <- roulette_vector[selection_vector]
roulette_winning_days
```


### Matrices

What's a matrix?

```{r}
# Construct a matrix with 3 rows that contain the numbers 1 up to 9
matrix(1:9, nrow = 3, byrow = TRUE)
```


Analyzing matrices, you shall

```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Create box_office
box_office <- c(new_hope, empire_strikes, return_jedi)

# Construct star_wars_matrix
star_wars_matrix <- matrix( nrow = 3, box_office, byrow = TRUE)
star_wars_matrix
```


Naming a matrix

```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Construct matrix
star_wars_matrix <- matrix(c(new_hope, empire_strikes, return_jedi), nrow = 3, byrow = TRUE)

# Vectors region and titles, used for naming
region <- c("US", "non-US")
titles <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")

# Name the columns with region
colnames(star_wars_matrix) <- region

# Name the rows with titles
rownames(star_wars_matrix) <- titles


# Print out star_wars_matrix
star_wars_matrix
```


Calculating the worldwide box office

```{r}
# Construct star_wars_matrix

box_office <- c(460.998, 314.4, 290.475, 247.900, 309.306, 165.8)
star_wars_matrix <- matrix(box_office, nrow = 3, byrow = TRUE,
                           dimnames = list(c("A New Hope", "The Empire Strikes Back", "Return of the Jedi"), 
                                           c("US", "non-US")))

# Calculate worldwide box office figures
worldwide_vector <- rowSums(star_wars_matrix)
worldwide_vector
```


Adding a column for the Worldwide box office

```{r}
# Construct star_wars_matrix
box_office <- c(460.998, 314.4, 290.475, 247.900, 309.306, 165.8)
star_wars_matrix <- matrix(box_office, nrow = 3, byrow = TRUE,
                           dimnames = list(c("A New Hope", "The Empire Strikes Back", "Return of the Jedi"), 
                                           c("US", "non-US")))

# The worldwide box office figures
worldwide_vector <- rowSums(star_wars_matrix)

# Bind the new variable worldwide_vector as a column to star_wars_matrix
all_wars_matrix <- cbind(star_wars_matrix, worldwide_vector)
all_wars_matrix
```


Adding a row

```{r}
# YOU NEED TO CREATE star_wars_matrix2 HERE.
box_office2 <- c(474.5, 552.5, 310.7, 338.7, 380.3, 468.5)
star_wars_matrix2 <- matrix(box_office2, nrow = 3, byrow = TRUE,
                           dimnames = list(c("The Phantom Menace", "Attack of the Clones", "Revenge of the Sith"), 
                                           c("US", "non-US")))
star_wars_matrix2
```

```{r}
# star_wars_matrix and star_wars_matrix2 are available in your workspace
star_wars_matrix  
star_wars_matrix2 

# Combine both Star Wars trilogies in one matrix
all_wars_matrix <- rbind(star_wars_matrix, star_wars_matrix2)
```


The total box office revenue for the entire saga

```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Total revenue for US and non-US
total_revenue_vector <- colSums(all_wars_matrix)
  
# Print out total_revenue_vector
total_revenue_vector
```


Selection of matrix elements

```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Select the non-US revenue for all movies
non_us_all <- all_wars_matrix[,2]
  
# Average non-US revenue
mean(non_us_all)
  
# Select the non-US revenue for first two movies
non_us_some <- all_wars_matrix[1:2,2]
  
# Average non-US revenue for first two movies
mean(non_us_some)
```


A little arithmetic with matrices

```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Estimate the visitors
visitors <- all_wars_matrix / 5
  
# Print the estimate to the console
visitors
```


A little arithmetic with matrices (2)

```{r}
# YOU NEED TO CREATE ticket_prices_matrix HERE.
box_office2 <- c(5, 5, 6, 6, 7, 7, 4, 4, 4.5, 4.5, 4.9, 4.9)
star_wars_matrix3 <- matrix(box_office2, nrow = 6, byrow = TRUE,
                           dimnames = list(c("A New Hope", "The Empire Strikes Back", "Return of the Jedi", "The Phantom Menace", "Attack of the Clones", "Revenge of the Sith"), 
                                           c("US", "non-US")))

ticket_prices_matrix <- cbind(star_wars_matrix3)
ticket_prices_matrix
```

```{r}
# all_wars_matrix and ticket_prices_matrix are available in your workspace
all_wars_matrix
ticket_prices_matrix

# Estimated number of visitors
visitors <- all_wars_matrix / ticket_prices_matrix

# US visitors
us_visitors <- visitors[,1]

# Average number of US visitors
mean(us_visitors)
```


### Factors

What's a factor and why would you use it?

```{r}
# Assign to the variable theory what this chapter is about!
theory <- "factors for categorical variables"
```


What's a factor and why would you use it? (2)

```{r}
# Gender vector
sex_vector <- c("Male", "Female", "Female", "Male", "Male")

# Convert gender_vector to a factor
factor_sex_vector <- factor(sex_vector)

# Print out factor_gender_vector
factor_sex_vector
```


What's a factor and why would you use it? (3)

```{r}
# Animals
animals_vector <- c("Elephant", "Giraffe", "Donkey", "Horse")
factor_animals_vector <- factor(animals_vector)
factor_animals_vector

# Temperature
temperature_vector <- c("High", "Low", "High","Low", "Medium")
factor_temperature_vector <- factor(temperature_vector, order = TRUE, levels = c("Low", "Medium", "High"))
factor_temperature_vector
```


Factor levels

```{r}
# Code to build factor_survey_vector
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)

# Specify the levels of factor_survey_vector
levels(factor_survey_vector) <- c("Female", "Male")

factor_survey_vector
```


Summarizing a factor

```{r}
# Build factor_survey_vector with clean levels

survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)
levels(factor_survey_vector) <- c("Female", "Male")
factor_survey_vector
# Generate summary for survey_vector
summary(survey_vector)

# Generate summary for factor_survey_vector
summary(factor_survey_vector)
```


Battle of the sexes

```{r}
# Build factor_survey_vector with clean levels
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)
levels(factor_survey_vector) <- c("Female", "Male")


# Male
male <- factor_survey_vector[1]

# Female
female <- factor_survey_vector[2]

# Battle of the sexes: Male 'larger' than female?
male > female
```


Ordered factors

```{r}
# Create speed_vector
speed_vector <- c("medium", "slow", "slow", "medium", "fast")
speed_vector
```


Ordered factors (2)

```{r}
# Create speed_vector
speed_vector <- c("medium", "slow", "slow", "medium", "fast")

# Convert speed_vector to ordered factor vector
factor_speed_vector <- factor(speed_vector, ordered = TRUE, levels = c("slow", "medium", "fast"))

# Print factor_speed_vector
factor_speed_vector
summary(factor_speed_vector)
```


Comparing ordered factors

```{r}
# Create factor_speed_vector
speed_vector <- c("medium", "slow", "slow", "medium", "fast")
factor_speed_vector <- factor(speed_vector, ordered = TRUE, levels = c("slow", "medium", "fast"))

# Factor value for second data analyst
da2 <- factor_speed_vector[2]

# Factor value for fifth data analyst
da5 <- factor_speed_vector[5]

# Is data analyst 2 faster than data analyst 5?
da2 > da5
```


### Data Frames

What's a data frame?

```{r}
# Print out built-in R data frame
mtcars 
```


Quick, have a look at your data set

```{r}
# Call head() on mtcars
head(mtcars)
```


Have a look at the structure

```{r}
# Investigate the structure of mtcars
str(mtcars)
```


Creating a data frame

```{r}
# Definition of vectors
name <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")
type <- c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet", 
          "Terrestrial planet", "Gas giant", "Gas giant", "Gas giant", "Gas giant")
diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)
rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)


# Create a data frame from the vectors
planets_df <- data.frame(name, type, diameter, rotation, rings)

```


Creating a data frame (2)

```{r}
# Check the structure of planets_df
str(planets_df)
```


Selection of data frame elements

```{r}
# Print out diameter of Mercury (row 1, column 3)
planets_df[1,3]

# Print out data for Mars (entire fourth row)
planets_df[4,]
```


Selection of data frame elements (2)

```{r}
# Select first 5 values of diameter column
planets_df[1:5, "diameter"]
```


Only planets with rings

```{r}
# Select the rings variable from planets_df
rings_vector <- planets_df$rings
  
# Print out rings_vector
rings_vector
```


Only planets with rings (2)

```{r}
# Adapt the code to select all columns for planets with rings
planets_df[rings_vector, ]
```


Only planets with rings but shorter

```{r}
# Select planets with diameter < 1
subset(planets_df, diameter < 1)
```


Sorting

```{r}
# Play around with the order function in the console
order(1)
```


Sorting your data frame

```{r}
# Use order() to create positions
positions <-  order(planets_df$diameter)

# Use positions to sort planets_df
planets_df[positions, ]
```


### Lists

Lists, why would you need them?

```{r}
# There is no code in this section.
```


Lists, why would you need them? (2)

```{r}
# There is no code in this section.
```


Creating a list

```{r}
# Vector with numerics from 1 up to 10
my_vector <- 1:10 

# Matrix with numerics from 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)

# First 10 elements of the built-in data frame mtcars
my_df <- mtcars[1:10,]

# Construct list with these different elements:
my_list <- list(my_vector, my_matrix, my_df)
```


Creating a named list

```{r}
# Vector with numerics from 1 up to 10
my_vector <- 1:10 

# Matrix with numerics from 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)

# First 10 elements of the built-in data frame mtcars
my_df <- mtcars[1:10,]

# Adapt list() call to give the components names
my_list <- list(my_vector, my_matrix, my_df)
my_list <- list(vec = my_vector, mat = my_matrix, df = my_df)
# Print out my_list
my_list
```


Creating a named list (2)

```{r}
# YOU NEED TO CREATE mov, act, and rev HERE
# Define the vectors mov and act
mov <- "The Shining"
act <- c("Jack Nicholson", "Shelley Duvall", "Danny Lloyd", "Scatman Crothers", "Barry Nelson")
# Define the vectors scores, sources, comments
scores <- c(4.5, 4, 5)
sources <- c("IMDb1", "IMDb2", "IMDb3")
factor_sources <- factor(sources)
levels(sources) <- c("IMDb1", "IMDb2", "IMDb3")

comments <- c("Best Horror Film I Have Ever Seen", "A truly brilliant and scary film from Stanley Kubrick", "A masterpiece of psychological horror ")
factor_comments <- factor(comments)
levels(comments) <- c("A masterpiece of psychological horror ", "A truly brilliant and scary film from Stanley Kubrick", "Best Horror Film I Have Ever Seen")

# Create a data frame rev from the vectors
rev <- cbind(scores, sources, comments)
```

```{r}
# Finish the code to build shining_list
shining_list <- list(moviename = mov, actors = act, reviews = rev)
shining_list
```


Selecting elements from a list

```{r}
# Print out the vector representing the actors
shining_list[[2]]
# Print out the vector representing the actors
shining_list$actors[2]

# Print the second element of the vector representing the actors

```


Adding more movie information to the list

```{r}
# We forgot something; add the year to shining_list
shining_list_full <- c(shining_list, year = 1980)

# Have a look at shining_list_full
str(shining_list_full)
```

***

```{r}
sessionInfo()
```
