---
title: 'Case study'
author: 'Jefferson Ong'
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document:
    theme: united
    highlight: tango
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
---

```{r, echo = FALSE}
library(tidyverse)
library(purrr)
library(Lahman)
library(resampledata)
```


# Mutating joins

● arrange()
● filter()
● select()
● mutate()
● summarise()

```{r}
# Complete the code to join artists to bands
# bands2 <- left_join(bands, artists, by = c("first", "last"))
# 
# # Examine the results
# bands2
# 
# # Fix the code to recreate bands2
# left_join(bands, artists, by = c("first", "last"))
# 
# # Finish the code below to recreate bands3 with a right join
# bands2 <- left_join(bands, artists, by = c("first", "last"))
# bands3 <- right_join(artists, bands, by = c("first", "last"))
# 
# # Check that bands3 is equal to bands2
# setequal(bands2, bands3)
```

```{r}
# Join albums to songs using inner_join()
# inner_join(songs, albums, by = "album")
# 
# # Join bands to artists using full_join()
# full_join(artists, bands, by = c("first", "last"))

# Find guitarists in bands dataset (don't change)
# temp <- left_join(bands, artists, by = c("first", "last"))
# temp <- filter(temp, instrument == "Guitar")
# select(temp, first, last, band)
# 
# # Reproduce code above using pipes
# bands %>% 
#   left_join(artists, by = c("first", "last")) %>% 
#   filter(instrument == "Guitar") %>% 
#   select(first, last, band)
```

```{r}
# Create goal2 using full_join() and inner_join() 
# goal2 <- artists %>% 
#   full_join(bands, by = c("first", "last")) %>% 
#   inner_join(songs, by = c("first", "last"))
# 
# # Check that goal and goal2 are the same
# setequal(goal, goal2)

# Create one table that combines all information
# artists %>% 
#   full_join(bands, by = c("first", "last")) %>% 
#   full_join(songs, by = c("first", "last")) %>% 
#  full_join(albums, by = c("album", "band"))
```

Summary
● left_join() add variables to the left side of a different dataset 
● right_join()
● inner_join()
● full_join()


# Filtering joins and set operations


```{r}
# albums %>% 
#   # Collect the albums made by a band
#   semi_join(bands, by = "band") %>% 
#   # Count the albums made by a band
#   nrow()
```

## anti join

```{r}
# Return rows of artists that don't have bands info
# artists %>% 
#   anti_join(bands, by = c("first", "last"))

# # Check whether album names in labels are mis-entered
# labels %>% 
#   anti_join(albums, by = "album")
```

```{r}
# aerosmith %>% 
#   # Create the new dataset using a set operation
#   union(greatest_hits) %>% 
#   # Count the total number of songs
#   nrow()
```

## Intersect

```{r}
# Create the new dataset using a set operation
# aerosmith %>% 
#   intersect(greatest_hits)
```

```{r}
# Select the song names from live
# live_songs <- live %>% select(song)
# 
# # Select the song names from greatest_hits
# greatest_songs <- greatest_hits %>% select(song)
# 
# # Create the new dataset using set operation
# live_songs %>% 
#  setdiff(greatest_songs)
```

```{r}
# # Check if same order: definitive and complete
# identical(definitive, complete)
# 
# # Check if any order: definitive and complete
# setequal(definitive, complete)
# 
# # Songs in definitive but not complete
# definitive %>% 
#   setdiff(complete)
# 
# # Songs in complete but not definitive
# complete %>% 
#   setdiff(definitive)
```


● union()
● intersect()
● setdiff()

Mutating joins
● left_join(), right_join(),
inner_join(), full_join()
Filtering joins
● semi_join(), anti_join()
Set operations
● union(), intersect(),
setdiff()
Comparisons
● setequal()



# Assembling data


Binds

● rbind()
● cbind()
● bind_rows()
● bind_cols()

data_frame()
data_frame() will not…
● Change the data type of vectors (e.g. strings to factors)
● Add row names
● Change column names
● Recycle vectors greater than length one

```{r}
# Bind side_one and side_two into a single dataset
# side_one %>% 
#   bind_rows(side_two)
# 
# jimi %>% 
#   # Bind jimi by rows, with ID column "album"
#   bind_rows(.id = "album") %>% 
#   # Make a complete data frame
#   left_join(discography)
```

```{r}
# hank_years %>% 
#   # Reorder hank_years alphabetically by song title
#   arrange(song) %>% 
#   # Select just the year column
#   select(year) %>% 
#   # Bind the year column
#   bind_cols(hank_charts) %>% 
#   # Arrange the finished dataset
#   arrange(year, song)
```

## Lists of rows (data frames)

```{r}
# Replace the first line so each album has its own rows
# bind_rows(michael, .id = "album") %>% 
#   group_by(album) %>% 
#   mutate(rank = min_rank(peak)) %>% 
#   filter(rank == 1) %>% 
#  select(-rank, -peak)
```


```{r}
# seventies %>% 
#   # Coerce seventies$year into a useful numeric
#   mutate(year = as.numeric(as.character(year))) %>% 
#   # Bind the updated version of seventies to sixties
#   bind_rows(sixties) %>% 
#   arrange(year)
```



# Advanced joining


```{r}
# stage_songs %>% 
#   # Add row names as a column named song
#   rownames_to_column("song") %>% 
#   # Left join stage_writers to stage_songs
#   left_join(stage_writers, by = "song")
```


```{r}
# Examine the result of joining singers to two_songs
# two_songs %>% inner_join(singers, by = "movie")
# 
# # Remove NA's from key before joining
# two_songs %>% 
#   filter(!is.na(movie)) %>% 
#   inner_join(singers, by = "movie")
```

```{r}
# movie_years %>% 
#   # Left join movie_studios to movie_years
#   left_join(movie_studios, by = "movie") %>% 
#   # Rename the columns: artist and studio
#   rename(artist = name.x, studio = name.y)
```

```{r}
# Identify the key columns
# movie_directors
# movie_years
# 
# movie_years %>% 
#   # Left join movie_directors to movie_years
#   left_join(movie_directors, by = c("movie" = "name")) %>% 
#   # Arrange the columns using select()
#  select(year, movie, artist = name, director, studio)
```

```{r}
# Place supergroups, more_bands, and more_artists into a list
# list(supergroups, more_bands, more_artists) %>%
#   # Use reduce to join together the contents of the list
#   reduce(left_join, by = c("first", "last"))
# 
# list(more_artists, more_bands, supergroups) %>% 
#   # Return rows of more_artists in all three datasets
#   reduce(semi_join, by = c("first", "last"))
```



# Case Study



```{r}
# # Examine lahmanNames
# lahmanNames
# 
# # Find variables in common
# reduce(lahmanNames, intersect)
```

```{r}
# lahmanNames %>%  
#   # Bind the data frames in lahmanNames
#   bind_rows(.id = "dataframe") %>%
#   # Group the result by var
#   group_by(var) %>%
#   # Tally the number of appearances
#   tally() %>%
#   # Filter the data
#   filter(n > 1) %>% 
#   # Arrange the results
#   arrange(desc(n))
```

```{r}
# lahmanNames %>% 
#   # Bind the data frames
#   bind_rows(.id = "dataframe") %>%
#   # Filter the results
#   filter(var == "playerID") %>%
#   # Extract the dataframe variable
#   `$`("dataframe")
```


```{r}
# players <- Master %>% 
#   # Return one row for each distinct player
#   distinct(playerID, nameFirst, nameLast)
# 
# 
# players %>% 
#   # Find all players who do not appear in Salaries
#   anti_join(Salaries, by = "playerID") %>%
#   # Count them
#   count()
```



```{r}
# players %>% 
#   # Find all players who do not appear in Salaries
#   anti_join(Salaries, by = "playerID") %>% 
#   # Join them to Appearances
#   left_join(Appearances, by = "playerID") %>%
#   # Calculate total_games for each player
#   group_by(playerID) %>% 
#   summarize(total_games = sum(G_all, na.rm = TRUE)) %>%
#   # Arrange in descending order by total_games
#   arrange(desc(total_games))
```


```{r}
# players %>%
#   # Find unsalaried players
#   anti_join(Salaries, by = "playerID") %>% 
#   # Join Batting to the unsalaried players
#   left_join(Batting, by = "playerID") %>% 
#   # Group by player
#   group_by(playerID) %>% 
#   # Sum at-bats for each player
#   summarize(total_at_bat = sum(AB, na.rm = TRUE)) %>% 
#   # Arrange in descending order
#  arrange(desc(total_at_bat))
```







```{r}
# Find the distinct players that appear in HallOfFame
# nominated <- HallOfFame %>% 
#   distinct(playerID)
# 
# nominated %>% 
#   # Count the number of players in nominated
#   count()
# 
# nominated_full <- nominated %>% 
#   # Join to Master
#   left_join(Master, by = "playerID") %>% 
#   # Return playerID, nameFirst, nameLast
#  select(playerID, nameFirst, nameLast)
```



```{r}
# Find distinct players in HallOfFame with inducted == "Y"
# inducted <- HallOfFame %>% 
#   filter(inducted == "Y") %>% 
#   distinct(playerID)
# 
# inducted %>% 
#   # Count the number of players in inducted
#   count()
# 
# inducted_full <- inducted %>% 
#   # Join to Master
#   left_join(Master, by = "playerID") %>% 
#   # Return playerID, nameFirst, nameLast
#   select(playerID, nameFirst, nameLast)
```

```{r}
# Tally the number of awards in AwardsPlayers by playerID
# nAwards <- AwardsPlayers %>% 
#   group_by(playerID) %>% 
#   tally()
# 
# nAwards %>% 
#   # Filter to just the players in inducted 
#   semi_join(inducted, by = "playerID") %>% 
#   # Calculate the mean number of awards per player
#   summarize(avg_n = mean(n, na.rm = TRUE))
# 
# nAwards %>% 
#   # Filter to just the players in nominated 
#   semi_join(nominated, by = "playerID") %>%
#   # Filter to players NOT in inducted 
#   anti_join(inducted, by = "playerID") %>%
#   # Calculate the mean number of awards per player
#  summarize(avg_n = mean(n, na.rm = TRUE))
```


```{r}
# Find the players who are in nominated, but not inducted
# notInducted <- nominated %>% 
#   setdiff(inducted)
# 
# Salaries %>% 
#   # Find the players who are in notInducted
#   semi_join(notInducted, by = "playerID") %>%
#   # Calculate the max salary by player
#   group_by(playerID) %>% 
#   summarize(max_salary = max(salary, na.rm = TRUE)) %>% 
#   # Calculate the average of the max salaries
#   summarize(avg_salary = mean(max_salary, na.rm = TRUE))
# 
# # Repeat for players who were inducted
# Salaries %>% 
#   semi_join(inducted, by = "playerID") %>% 
#   group_by(playerID) %>% 
#   summarize(max_salary = max(salary, na.rm = TRUE)) %>% 
#  summarize(avg_salary = mean(max_salary, na.rm = TRUE))
```



```{r}
# Appearances %>% 
#   # Filter Appearances against nominated
#   semi_join(nominated, by = "playerID") %>% 
#   # Find last year played by player
#   group_by(playerID) %>% 
#   summarize(last_year = max(yearID)) %>% 
#   # Join to full HallOfFame
#   left_join(HallOfFame, by = "playerID") %>% 
#   # Filter for unusual observations
#   filter(last_year >= yearID)
```








































































