---
title: "MAT 4310 - Linear System Capstone"
author: "Jefferson Ong"
date: "3/6/2020"
output:
  pdf_document: default
  html_document:
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)
```

https://repl.it/@ongjk/TruthfulMeanFolders

https://repl.it/@ongjk/QuarrelsomeReliableComma



# Solving Large Systems.

In a file in ASULearn there is a 50 × 50 matrix generated using random numbers
between -5 and 5, with the diagonal elements set to 200. (Thank goodness for Maple!) Also included in the file is the b-vector that causes the systems’s solution to be a vector of ones.

## (a)

Solve the system using GE SPP. What is the estimated condition number for this matrix? How many digits
are correct in the solution for GE SPP? Approximately how many operations are performed?

Formula for GE_SPP operation count: $$\frac{2}{3} n^3 + O(n^2)$$

```{r}
n = 50
n^2
2/3 * n^3
```


```{python}
import math

def GE_SPP(A, b):
  n = len(A) # if not built-in in a language,
 # read in n as the third argument
  L = [i for i in range(n)] # a quick way to initialize L
  x = [0.0 for i in range(n)]
  S = [0.0 for i in range(n)]
 #build the S vector
  for i in range(n): # find largest element in each    #row of a
    for j in range(n):
      if (abs(A[i][j])>S[i]):
        S[i] = abs(A[i][j])

#could calculate cond = the product of the two-norms of the rows of A here
  p = 1
  for i in range(n):
    s = 0
    for j in range(n):
      s += A[i][j]**2
    p = p *math.sqrt(s)

#forward elimination
  for k in range(n-1):
   # check to see whether we need to swap rows
    R = 0.0
    for i in range(k, n): #from k to n-1
      temp = abs(A[L[i]][k]/S[L[i]])
      if temp > R:
        R = temp
        index = i
 # do the swap
    temp = L[index]
    L[index] = L[k]
    L[k] = temp
 # zero out below the diagonal (update both a and b)
    for i in range(k+1,n):
      xmult = A[L[i]][k]/A[L[k]][k]
      for j in range(k+1,n):
        A[L[i]][j] = A[L[i]][j]-xmult*A[L[k]][j]
      b[L[i]] = b[L[i]]-xmult*b[L[k]]
#could divide cond by the absolute values of the pivots here

  for i in range(n):
    p = p / abs(A[L[i]][i])

#back substitution
  x[n-1] = b[L[n-1]]/A[L[n-1]][n-1]
  for k in range(n-2,-1, -1):
    sum = b[L[k]]
    for j in range(k+1,n):
      sum = sum - A[L[k]][j]*x[j]
    x[k] = sum/A[L[k]][k]

  print("the solution is:", x)
  print("the condition number estimation is", p)
  print(L)
 #also output cond!

#A= [[1, 1, 1, 1, 1], [1, 3, 3**2, 3**3, 3**4], [1, 1, 7, 7**2, 7**3], [1, 1, 1, 10, 10**2], [1,1,1,1,14] ]
#b = [22, 1, -2, 12, -2]

A = [[200, 4, 5, -1, 0, -5, -5, 0, -3, -2, -2, 3, -4, 0, 4, -1, 1, 1, 5, 0, 1, 3, 3, -2,
3, -2, 0, -2, 4, -4, -4, -2, 4, -3, 2, -1, 0, -3, 0, -4, -5, 4, 1, 0, -1, 3, 3, 4, -1,
-2], [4, 200, 4, 1, -5, 0, 0, 3, 1, 4, 2, -1, 2, 0, -5, -2, 4, -2, -2, 0, -2, 0, 4, 2, 5,
-1, 0, 5, 2, 4, -1, 0, -5, 4, 0, -1, -1, 0, 0, -3, -1, 3, -1, -5, -2, -2, 1, -2, -2, 2],
[5, 4, 200, 0, -4, 0, -3, 0, -5, -4, 3, 3, 4, 4, 1, 4, -5, 0, 0, -2, -4, -1, 4, 0, 3, 3,
4, -3, 0, 0, -2, -2, -1, 4, 1, 0, 0, 3, 1, 3, -5, 1, 4, 2, 5, -3, -2, 0, -5, 5], [-1, 1,
0, 200, 5, 0, 0, 0, 5, 1, -2, 0, -3, 4, 0, 0, 3, 4, -1, 2, 3, 0, 0, -1, 1, 4, 0, 4, -4, 0,
3, 5, 5, 5, 2, 1, -4, -3, -5, 0, -2, 0, 0, 1, 3, 0, -5, 1, 0, 5], [0, -5, -4, 5, 200, 0,
-4, -2, 0, -2, -5, 5, 0, -4, 0, 0, 3, -1, 5, 5, -4, 4, 5, 5, 0, 0, -4, -2, -1, 0, -5, 0,
4, 0, -5, 1, 3, 2, 5, 0, 0, 2, 4, -5, -3, 0, -2, -4, -3, 2], [-5, 0, 0, 0, 0, 200, 0, 0,
-4, 0, 2, -5, 0, -5, 1, 0, 2, -1, 2, 0, 0, 3, 0, -2, 1, -1, 0, 3, 2, 3, 3, 0, 5, 1, -2,
-5, -4, 0, -1, 0, -5, 0, -5, -2, 0, 2, 1, 0, 3, -3], [-5, 0, -3, 0, -4, 0, 200, 2, -1, -4,
4, 0, -2, 3, 0, -1, 0, 5, 5, 0, 4, 3, 0, 5, 2, 1, 0, 0, 0, 3, -5, 1, 3, -3, -2, -3, -2,
-5, 4, 0, -3, -2, 0, 0, 0, -2, 0, -4, 3, 3], [0, 3, 0, 0, -2, 0, 2, 200, 0, 0, 0, 2, 0, 3,
3, -1, 3, -3, 0, 1, 2, -3, 1, 0, -4, 3, 0, 5, 1, -3, 1, -5, -2, 0, -1, 0, 0, -1, -4, 1, 4,
-3, 3, 0, -3, 0, 0, 0, 3, 0], [-3, 1, -5, 5, 0, -4, -1, 0, 200, 5, 0, 4, 3, -3, 4, 0, -4,
-4, 0, 0, -4, 0, -3, 0, 0, 1, -5, 5, 1, 3, 1, 5, 1, 3, -4, 5, -2, 4, 4, 0, -4, 3, 0, 4, 0,
-3, -1, -2, 0, 0], [-2, 4, -4, 1, -2, 0, -4, 0, 5, 200, 4, 0, 0, -3, 2, 2, -3, -1, -5, 2,
2, 0, 0, -3, 4, 1, 0, 2, 1, -4, -3, 0, -3, 0, -1, -1, -1, 0, -4, 4, -3, 0, -5, 0, 3, -4,
0, 2, 2, 0], [-2, 2, 3, -2, -5, 2, 4, 0, 0, 4, 200, 0, -4, 0, 0, -5, 0, -3, -3, 1, -3, -1,
1, 2, 0, -5, -4, -5, 0, 0, -3, 3, 0, -1, -2, 2, -3, -1, 1, -1, -4, -3, 2, -2, 0, -2, 3, 0,
2, 2], [3, -1, 3, 0, 5, -5, 0, 2, 4, 0, 0, 200, 4, -5, 4, -2, 0, 3, 3, 0, -4, -3, -4, -1,
-5, -4, 0, 0, 0, -1, 0, 4, 2, -5, 0, 0, -4, 4, 0, 0, -3, 5, 3, 5, 2, 0, 0, 3, 3, -2], [-4,
2, 4, -3, 0, 0, -2, 0, 3, 0, -4, 4, 200, -5, -5, 5, -2, -3, 4, -2, 0, 0, 3, 0, 4, 0, -1,
1, 1, -2, -5, 0, 0, -3, 4, 0, 0, 0, 4, -5, -5, 0, -4, 5, -2, -4, -3, -1, 1, -3], [0, 0, 4,
4, -4, -5, 3, 3, -3, -3, 0, -5, -5, 200, 4, 3, 0, -4, 0, 0, -3, -3, 4, 0, 0, -1, 0, 4, 1,
1, 2, 0, 2, -3, 4, -1, -2, -4, 0, 5, 3, 1, -4, 4, 0, 1, 1, 0, 0, 4], [4, -5, 1, 0, 0, 1,
0, 3, 4, 2, 0, 4, -5, 4, 200, 2, 2, 0, 0, 2, -3, 0, 2, 0, 1, 0, -4, -3, -2, -4, 5, -4, 4,
0, 5, 0, -5, 4, -3, 5, -4, 0, 4, 5, -4, 0, 1, -4, 4, 0], [-1, -2, 4, 0, 0, 0, -1, -1, 0,
2, -5, -2, 5, 3, 2, 200, 5, -3, 0, 1, 0, 3, 0, -2, -5, -4, 3, 2, -4, 1, 0, -4, -4, 4, 0,
5, 0, 0, -5, -5, 0, 0, 1, 0, 0, 1, -1, 0, 0, 5], [1, 4, -5, 3, 3, 2, 0, 3, -4, -3, 0, 0,
-2, 0, 2, 5, 200, 4, 0, 0, 0, 4, -2, -1, 0, -5, -2, 4, 0, 4, 1, -1, 0, -4, 0, 3, -4, 4,
-5, 0, -4, 4, 1, 3, 0, 2, -2, 0, -4, 0], [1, -2, 0, 4, -1, -1, 5, -3, -4, -1, -3, 3, -3,
-4, 0, -3, 4, 200, 1, -4, 0, 1, 2, 0, 0, 0, 0, 0, 0, 0, 0, -2, 0, 4, 0, -2, 0, 0, -2, -3,
-2, 0, -5, 0, 1, 0, 3, 0, 0, -5], [5, -2, 0, -1, 5, 2, 5, 0, 0, -5, -3, 3, 4, 0, 0, 0, 0,
1, 200, -1, -1, -2, 5, 5, -3, 5, -4, -1, 5, 5, 5, -4, -2, -1, 0, -1, 0, 0, 3, 3, 2, 4, -5,
3, 1, -4, 0, 5, -5, 1], [0, 0, -2, 2, 5, 0, 0, 1, 0, 2, 1, 0, -2, 0, 2, 1, 0, -4, -1, 200,
0, -3, -3, 1, 1, -2, 1, -2, 0, 3, -1, -4, 0, 0, 4, 0, -2, -3, 5, 2, 4, 4, -1, 5, 0, -4, 3,
-5, 5, 0], [1, -2, -4, 3, -4, 0, 4, 2, -4, 2, -3, -4, 0, -3, -3, 0, 0, 0, -1, 0, 200, 2,
1, 3, -2, 0, 3, 3, 1, 3, -1, 0, -3, -1, 1, 0, 3, 0, -4, 3, 5, 4, 2, 4, -5, -4, 0, 0, 0,
1], [3, 0, -1, 0, 4, 3, 3, -3, 0, 0, -1, -3, 0, -3, 0, 3, 4, 1, -2, -3, 2, 200, 0, 0, 2,
-5, -3, -3, -2, 1, -4, 1, -1, 2, 1, 0, 0, -4, -1, -1, -4, 1, 0, 4, -2, 5, 0, 2, -3, -4],
[3, 4, 4, 0, 5, 0, 0, 1, -3, 0, 1, -4, 3, 4, 2, 0, -2, 2, 5, -3, 1, 0, 200, 4, -5, 5, 0,
0, -2, 2, -2, 0, 5, 1, 0, -2, 1, -4, 0, 5, 0, 1, -4, -3, -4, 4, 5, -4, -3, 5], [-2, 2, 0,
-1, 5, -2, 5, 0, 0, -3, 2, -1, 0, 0, 0, -2, -1, 0, 5, 1, 3, 0, 4, 200, 0, 3, 2, 0, 3, 4,
0, 5, 0, -3, -2, 0, 1, 3, 0, 0, -5, 1, 4, 2, 5, 0, 0, 1, 3, 0], [3, 5, 3, 1, 0, 1, 2, -4,
0, 4, 0, -5, 4, 0, 1, -5, 0, 0, -3, 1, -2, 2, -5, 0, 200, -1, -3, -4, 0, 3, 0, -5, 0, 1,
0, 0, 1, -3, 0, 1, 0, 0, 5, -5, 0, -3, 5, 4, 0, 0], [-2, -1, 3, 4, 0, -1, 1, 3, 1, 1, -5,
-4, 0, -1, 0, -4, -5, 0, 5, -2, 0, -5, 5, 3, -1, 200, 0, -5, -1, 0, -1, -4, 0, 0, -2, 2,
1, 5, 5, 2, 0, -5, 3, -5, 2, 2, -4, -5, -3, -2], [0, 0, 4, 0, -4, 0, 0, 0, -5, 0, -4, 0,
-1, 0, -4, 3, -2, 0, -4, 1, 3, -3, 0, 2, -3, 0, 200, 1, -4, 0, 0, 3, 0, 4, -1, -1, 4, 1,
1, 5, 2, -2, 0, 1, -1, -1, 0, 3, 3, 0], [-2, 5, -3, 4, -2, 3, 0, 5, 5, 2, -5, 0, 1, 4, -3,
2, 4, 0, -1, -2, 3, -3, 0, 0, -4, -5, 1, 200, 3, 0, 0, 5, 1, 4, 4, -3, -2, 4, -5, 2, -2,
-4, 0, 0, 4, 1, 0, 4, 2, 5], [4, 2, 0, -4, -1, 2, 0, 1, 1, 1, 0, 0, 1, 1, -2, -4, 0, 0, 5,
0, 1, -2, -2, 3, 0, -1, -4, 3, 200, 0, 0, 1, -2, 0, 0, 0, -4, 1, -5, 2, 0, 4, -1, 1, 1,
-4, 3, 1, -1, 2], [-4, 4, 0, 0, 0, 3, 3, -3, 3, -4, 0, -1, -2, 1, -4, 1, 4, 0, 5, 3, 3, 1,
2, 4, 3, 0, 0, 0, 0, 200, 4, 0, 0, 5, -4, 0, 2, 1, -1, 3, -1, -1, 5, 0, 2, -3, 0, -3, -2,
0], [-4, -1, -2, 3, -5, 3, -5, 1, 1, -3, -3, 0, -5, 2, 5, 0, 1, 0, 5, -1, -1, -4, -2, 0,
0, -1, 0, 0, 0, 4, 200, 4, 5, 0, -4, -5, -3, 0, 0, -1, -3, -3, 3, 0, 2, 0, -2, 5, -1, 1],
[-2, 0, -2, 5, 0, 0, 1, -5, 5, 0, 3, 4, 0, 0, -4, -4, -1, -2, -4, -4, 0, 1, 0, 5, -5, -4,
3, 5, 1, 0, 4, 200, 5, 4, 5, 3, 4, 1, -1, 0, 1, 5, -3, 3, 1, 0, 0, -3, -4, -1], [4, -5,
-1, 5, 4, 5, 3, -2, 1, -3, 0, 2, 0, 2, 4, -4, 0, 0, -2, 0, -3, -1, 5, 0, 0, 0, 0, 1, -2,
0, 5, 5, 200, -1, -2, 4, 2, 4, -5, -5, -3, 0, 4, -1, 2, 0, 0, 2, 5, 0], [-3, 4, 4, 5, 0,
1, -3, 0, 3, 0, -1, -5, -3, -3, 0, 4, -4, 4, -1, 0, -1, 2, 1, -3, 1, 0, 4, 4, 0, 5, 0, 4,
-1, 200, -3, -5, 3, 0, -3, 2, 4, -4, 1, 5, 0, -1, 2, -4, 5, -2], [2, 0, 1, 2, -5, -2, -2,
-1, -4, -1, -2, 0, 4, 4, 5, 0, 0, 0, 0, 4, 1, 1, 0, -2, 0, -2, -1, 4, 0, -4, -4, 5, -2,
-3, 200, 2, -3, -4, -2, -1, 0, 0, 4, 1, -5, -5, 3, 3, 2, -4], [-1, -1, 0, 1, 1, -5, -3, 0,
5, -1, 2, 0, 0, -1, 0, 5, 3, -2, -1, 0, 0, 0, -2, 0, 0, 2, -1, -3, 0, 0, -5, 3, 4, -5, 2,
200, 0, -1, 0, 4, 5, 0, 5, -3, 4, 0, 0, 5, -4, 1], [0, -1, 0, -4, 3, -4, -2, 0, -2, -1,
-3, -4, 0, -2, -5, 0, -4, 0, 0, -2, 3, 0, 1, 1, 1, 1, 4, -2, -4, 2, -3, 4, 2, 3, -3, 0,
200, 1, -3, 1, 3, 0, -3, -3, 5, 4, 1, -3, -5, -2], [-3, 0, 3, -3, 2, 0, -5, -1, 4, 0, -1,
4, 0, -4, 4, 0, 4, 0, 0, -3, 0, -4, -4, 3, -3, 5, 1, 4, 1, 1, 0, 1, 4, 0, -4, -1, 1, 200,
0, -3, 1, 3, 0, 3, 0, 0, -4, 0, 0, 3], [0, 0, 1, -5, 5, -1, 4, -4, 4, -4, 1, 0, 4, 0, -3,
-5, -5, -2, 3, 5, -4, -1, 0, 0, 0, 5, 1, -5, -5, -1, 0, -1, -5, -3, -2, 0, -3, 0, 200, -5,
0, -4, 0, -4, 3, 3, 1, -1, -2, 5], [-4, -3, 3, 0, 0, 0, 0, 1, 0, 4, -1, 0, -5, 5, 5, -5,
0, -3, 3, 2, 3, -1, 5, 0, 1, 2, 5, 2, 2, 3, -1, 0, -5, 2, -1, 4, 1, -3, -5, 200, -1, 0, 5,
5, 0, 0, 1, 3, 3, 5], [-5, -1, -5, -2, 0, -5, -3, 4, -4, -3, -4, -3, -5, 3, -4, 0, -4, -2,
2, 4, 5, -4, 0, -5, 0, 0, 2, -2, 0, -1, -3, 1, -3, 4, 0, 5, 3, 1, 0, -1, 200, -3, 0, -5,
4, 0, 5, 0, -1, -2], [4, 3, 1, 0, 2, 0, -2, -3, 3, 0, -3, 5, 0, 1, 0, 0, 4, 0, 4, 4, 4, 1,
1, 1, 0, -5, -2, -4, 4, -1, -3, 5, 0, -4, 0, 0, 0, 3, -4, 0, -3, 200, -1, -5, -5, -4, 1,
-1, 2, 0], [1, -1, 4, 0, 4, -5, 0, 3, 0, -5, 2, 3, -4, -4, 4, 1, 1, -5, -5, -1, 2, 0, -4,
4, 5, 3, 0, 0, -1, 5, 3, -3, 4, 1, 4, 5, -3, 0, 0, 5, 0, -1, 200, 0, -3, 0, 4, 0, 3, 3],
[0, -5, 2, 1, -5, -2, 0, 0, 4, 0, -2, 5, 5, 4, 5, 0, 3, 0, 3, 5, 4, 4, -3, 2, -5, -5, 1,
0, 1, 0, 0, 3, -1, 5, 1, -3, -3, 3, -4, 5, -5, -5, 0, 200, 2, 0, 4, 3, 5, 0], [-1, -2, 5,
3, -3, 0, 0, -3, 0, 3, 0, 2, -2, 0, -4, 0, 0, 1, 1, 0, -5, -2, -4, 5, 0, 2, -1, 4, 1, 2,
2, 1, 2, 0, -5, 4, 5, 0, 3, 0, 4, -5, -3, 2, 200, -5, 4, 0, 4, 0], [3, -2, -3, 0, 0, 2,
-2, 0, -3, -4, -2, 0, -4, 1, 0, 1, 2, 0, -4, -4, -4, 5, 4, 0, -3, 2, -1, 1, -4, -3, 0, 0,
0, -1, -5, 0, 4, 0, 3, 0, 0, -4, 0, 0, -5, 200, 0, 0, 4, -4], [3, 1, -2, -5, -2, 1, 0, 0,
-1, 0, 3, 0, -3, 1, 1, -1, -2, 3, 0, 3, 0, 0, 5, 0, 5, -4, 0, 0, 3, 0, -2, 0, 0, 2, 3, 0,
1, -4, 1, 1, 5, 1, 4, 4, 4, 0, 200, 4, -4, 5], [4, -2, 0, 1, -4, 0, -4, 0, -2, 2, 0, 3,
-1, 0, -4, 0, 0, 0, 5, -5, 0, 2, -4, 1, 4, -5, 3, 4, 1, -3, 5, -3, 2, -4, 3, 5, -3, 0, -1,
3, 0, -1, 0, 3, 0, 0, 4, 200, 0, 0], [-1, -2, -5, 0, -3, 3, 3, 3, 0, 2, 2, 3, 1, 0, 4, 0,
-4, 0, -5, 5, 0, -3, -3, 3, 0, -3, 3, 2, -1, -2, -1, -4, 5, 5, 2, -4, -5, 0, -2, 3, -1, 2,
3, 5, 4, 4, -4, 0, 200, 2], [-2, 2, 5, 5, 2, -3, 3, 0, 0, 0, 2, -2, -3, 4, 0, 5, 0, -5, 1,
0, 1, -4, 5, 0, 0, -2, 0, 5, 2, 0, 1, -1, 0, -2, -4, 1, -2, 3, 5, 5, -2, 0, 3, 0, 0, -4,
5, 0, 2, 200]]

b = [ 0.0 for i in range(50)]
for i in range(50):
   for j in range(50):
      b[i] = b[i] + A[i][j]
GE_SPP(A, b)
```

* Estimated condition number: `1.6130031646829517`

* Digits correct: `15`

* Number of operations: 83333.33 + O*2500

## (b)

Solve the system using Gauss Seidel, with max its set to to 100, and set the tolerance to the number of digits
found to be accurate for GE SPP. Approximately how many operations were performed?

Formula for GS operation count: $$5n^2 + O(n)$$

```{r}
n = 50
5 * (50)^2
```



```{python}
import math


def GS(A, b, max_its, tolerance):
  n = len(b)
  x = [0.0 for i in range(n)]
  xnew = [0.0 for i in range(n)]
  for its in range(1, max_its + 1):
    for i in range(0, n):
      s = b[i]
      for j in range(0, i):
        s -= A[i][j]*xnew[j]
      for j in range(i + 1, n):
        s -= A[i][j]*x[j]
      xnew[i] = s/A[i][i]
    print(xnew)

    s = 0
    for i in range(n):
      s += (xnew[i] - x[i])**2
    if math.sqrt(s) < tolerance:
      print("convergence", its)
      return
    else:
      for i in range(n):
        x[i] = xnew[i]


A = [[200, 4, 5, -1, 0, -5, -5, 0, -3, -2, -2, 3, -4, 0, 4, -1, 1, 1, 5, 0, 1, 3, 3, -2,
3, -2, 0, -2, 4, -4, -4, -2, 4, -3, 2, -1, 0, -3, 0, -4, -5, 4, 1, 0, -1, 3, 3, 4, -1,
-2], [4, 200, 4, 1, -5, 0, 0, 3, 1, 4, 2, -1, 2, 0, -5, -2, 4, -2, -2, 0, -2, 0, 4, 2, 5,
-1, 0, 5, 2, 4, -1, 0, -5, 4, 0, -1, -1, 0, 0, -3, -1, 3, -1, -5, -2, -2, 1, -2, -2, 2],
[5, 4, 200, 0, -4, 0, -3, 0, -5, -4, 3, 3, 4, 4, 1, 4, -5, 0, 0, -2, -4, -1, 4, 0, 3, 3,
4, -3, 0, 0, -2, -2, -1, 4, 1, 0, 0, 3, 1, 3, -5, 1, 4, 2, 5, -3, -2, 0, -5, 5], [-1, 1,
0, 200, 5, 0, 0, 0, 5, 1, -2, 0, -3, 4, 0, 0, 3, 4, -1, 2, 3, 0, 0, -1, 1, 4, 0, 4, -4, 0,
3, 5, 5, 5, 2, 1, -4, -3, -5, 0, -2, 0, 0, 1, 3, 0, -5, 1, 0, 5], [0, -5, -4, 5, 200, 0,
-4, -2, 0, -2, -5, 5, 0, -4, 0, 0, 3, -1, 5, 5, -4, 4, 5, 5, 0, 0, -4, -2, -1, 0, -5, 0,
4, 0, -5, 1, 3, 2, 5, 0, 0, 2, 4, -5, -3, 0, -2, -4, -3, 2], [-5, 0, 0, 0, 0, 200, 0, 0,
-4, 0, 2, -5, 0, -5, 1, 0, 2, -1, 2, 0, 0, 3, 0, -2, 1, -1, 0, 3, 2, 3, 3, 0, 5, 1, -2,
-5, -4, 0, -1, 0, -5, 0, -5, -2, 0, 2, 1, 0, 3, -3], [-5, 0, -3, 0, -4, 0, 200, 2, -1, -4,
4, 0, -2, 3, 0, -1, 0, 5, 5, 0, 4, 3, 0, 5, 2, 1, 0, 0, 0, 3, -5, 1, 3, -3, -2, -3, -2,
-5, 4, 0, -3, -2, 0, 0, 0, -2, 0, -4, 3, 3], [0, 3, 0, 0, -2, 0, 2, 200, 0, 0, 0, 2, 0, 3,
3, -1, 3, -3, 0, 1, 2, -3, 1, 0, -4, 3, 0, 5, 1, -3, 1, -5, -2, 0, -1, 0, 0, -1, -4, 1, 4,
-3, 3, 0, -3, 0, 0, 0, 3, 0], [-3, 1, -5, 5, 0, -4, -1, 0, 200, 5, 0, 4, 3, -3, 4, 0, -4,
-4, 0, 0, -4, 0, -3, 0, 0, 1, -5, 5, 1, 3, 1, 5, 1, 3, -4, 5, -2, 4, 4, 0, -4, 3, 0, 4, 0,
-3, -1, -2, 0, 0], [-2, 4, -4, 1, -2, 0, -4, 0, 5, 200, 4, 0, 0, -3, 2, 2, -3, -1, -5, 2,
2, 0, 0, -3, 4, 1, 0, 2, 1, -4, -3, 0, -3, 0, -1, -1, -1, 0, -4, 4, -3, 0, -5, 0, 3, -4,
0, 2, 2, 0], [-2, 2, 3, -2, -5, 2, 4, 0, 0, 4, 200, 0, -4, 0, 0, -5, 0, -3, -3, 1, -3, -1,
1, 2, 0, -5, -4, -5, 0, 0, -3, 3, 0, -1, -2, 2, -3, -1, 1, -1, -4, -3, 2, -2, 0, -2, 3, 0,
2, 2], [3, -1, 3, 0, 5, -5, 0, 2, 4, 0, 0, 200, 4, -5, 4, -2, 0, 3, 3, 0, -4, -3, -4, -1,
-5, -4, 0, 0, 0, -1, 0, 4, 2, -5, 0, 0, -4, 4, 0, 0, -3, 5, 3, 5, 2, 0, 0, 3, 3, -2], [-4,
2, 4, -3, 0, 0, -2, 0, 3, 0, -4, 4, 200, -5, -5, 5, -2, -3, 4, -2, 0, 0, 3, 0, 4, 0, -1,
1, 1, -2, -5, 0, 0, -3, 4, 0, 0, 0, 4, -5, -5, 0, -4, 5, -2, -4, -3, -1, 1, -3], [0, 0, 4,
4, -4, -5, 3, 3, -3, -3, 0, -5, -5, 200, 4, 3, 0, -4, 0, 0, -3, -3, 4, 0, 0, -1, 0, 4, 1,
1, 2, 0, 2, -3, 4, -1, -2, -4, 0, 5, 3, 1, -4, 4, 0, 1, 1, 0, 0, 4], [4, -5, 1, 0, 0, 1,
0, 3, 4, 2, 0, 4, -5, 4, 200, 2, 2, 0, 0, 2, -3, 0, 2, 0, 1, 0, -4, -3, -2, -4, 5, -4, 4,
0, 5, 0, -5, 4, -3, 5, -4, 0, 4, 5, -4, 0, 1, -4, 4, 0], [-1, -2, 4, 0, 0, 0, -1, -1, 0,
2, -5, -2, 5, 3, 2, 200, 5, -3, 0, 1, 0, 3, 0, -2, -5, -4, 3, 2, -4, 1, 0, -4, -4, 4, 0,
5, 0, 0, -5, -5, 0, 0, 1, 0, 0, 1, -1, 0, 0, 5], [1, 4, -5, 3, 3, 2, 0, 3, -4, -3, 0, 0,
-2, 0, 2, 5, 200, 4, 0, 0, 0, 4, -2, -1, 0, -5, -2, 4, 0, 4, 1, -1, 0, -4, 0, 3, -4, 4,
-5, 0, -4, 4, 1, 3, 0, 2, -2, 0, -4, 0], [1, -2, 0, 4, -1, -1, 5, -3, -4, -1, -3, 3, -3,
-4, 0, -3, 4, 200, 1, -4, 0, 1, 2, 0, 0, 0, 0, 0, 0, 0, 0, -2, 0, 4, 0, -2, 0, 0, -2, -3,
-2, 0, -5, 0, 1, 0, 3, 0, 0, -5], [5, -2, 0, -1, 5, 2, 5, 0, 0, -5, -3, 3, 4, 0, 0, 0, 0,
1, 200, -1, -1, -2, 5, 5, -3, 5, -4, -1, 5, 5, 5, -4, -2, -1, 0, -1, 0, 0, 3, 3, 2, 4, -5,
3, 1, -4, 0, 5, -5, 1], [0, 0, -2, 2, 5, 0, 0, 1, 0, 2, 1, 0, -2, 0, 2, 1, 0, -4, -1, 200,
0, -3, -3, 1, 1, -2, 1, -2, 0, 3, -1, -4, 0, 0, 4, 0, -2, -3, 5, 2, 4, 4, -1, 5, 0, -4, 3,
-5, 5, 0], [1, -2, -4, 3, -4, 0, 4, 2, -4, 2, -3, -4, 0, -3, -3, 0, 0, 0, -1, 0, 200, 2,
1, 3, -2, 0, 3, 3, 1, 3, -1, 0, -3, -1, 1, 0, 3, 0, -4, 3, 5, 4, 2, 4, -5, -4, 0, 0, 0,
1], [3, 0, -1, 0, 4, 3, 3, -3, 0, 0, -1, -3, 0, -3, 0, 3, 4, 1, -2, -3, 2, 200, 0, 0, 2,
-5, -3, -3, -2, 1, -4, 1, -1, 2, 1, 0, 0, -4, -1, -1, -4, 1, 0, 4, -2, 5, 0, 2, -3, -4],
[3, 4, 4, 0, 5, 0, 0, 1, -3, 0, 1, -4, 3, 4, 2, 0, -2, 2, 5, -3, 1, 0, 200, 4, -5, 5, 0,
0, -2, 2, -2, 0, 5, 1, 0, -2, 1, -4, 0, 5, 0, 1, -4, -3, -4, 4, 5, -4, -3, 5], [-2, 2, 0,
-1, 5, -2, 5, 0, 0, -3, 2, -1, 0, 0, 0, -2, -1, 0, 5, 1, 3, 0, 4, 200, 0, 3, 2, 0, 3, 4,
0, 5, 0, -3, -2, 0, 1, 3, 0, 0, -5, 1, 4, 2, 5, 0, 0, 1, 3, 0], [3, 5, 3, 1, 0, 1, 2, -4,
0, 4, 0, -5, 4, 0, 1, -5, 0, 0, -3, 1, -2, 2, -5, 0, 200, -1, -3, -4, 0, 3, 0, -5, 0, 1,
0, 0, 1, -3, 0, 1, 0, 0, 5, -5, 0, -3, 5, 4, 0, 0], [-2, -1, 3, 4, 0, -1, 1, 3, 1, 1, -5,
-4, 0, -1, 0, -4, -5, 0, 5, -2, 0, -5, 5, 3, -1, 200, 0, -5, -1, 0, -1, -4, 0, 0, -2, 2,
1, 5, 5, 2, 0, -5, 3, -5, 2, 2, -4, -5, -3, -2], [0, 0, 4, 0, -4, 0, 0, 0, -5, 0, -4, 0,
-1, 0, -4, 3, -2, 0, -4, 1, 3, -3, 0, 2, -3, 0, 200, 1, -4, 0, 0, 3, 0, 4, -1, -1, 4, 1,
1, 5, 2, -2, 0, 1, -1, -1, 0, 3, 3, 0], [-2, 5, -3, 4, -2, 3, 0, 5, 5, 2, -5, 0, 1, 4, -3,
2, 4, 0, -1, -2, 3, -3, 0, 0, -4, -5, 1, 200, 3, 0, 0, 5, 1, 4, 4, -3, -2, 4, -5, 2, -2,
-4, 0, 0, 4, 1, 0, 4, 2, 5], [4, 2, 0, -4, -1, 2, 0, 1, 1, 1, 0, 0, 1, 1, -2, -4, 0, 0, 5,
0, 1, -2, -2, 3, 0, -1, -4, 3, 200, 0, 0, 1, -2, 0, 0, 0, -4, 1, -5, 2, 0, 4, -1, 1, 1,
-4, 3, 1, -1, 2], [-4, 4, 0, 0, 0, 3, 3, -3, 3, -4, 0, -1, -2, 1, -4, 1, 4, 0, 5, 3, 3, 1,
2, 4, 3, 0, 0, 0, 0, 200, 4, 0, 0, 5, -4, 0, 2, 1, -1, 3, -1, -1, 5, 0, 2, -3, 0, -3, -2,
0], [-4, -1, -2, 3, -5, 3, -5, 1, 1, -3, -3, 0, -5, 2, 5, 0, 1, 0, 5, -1, -1, -4, -2, 0,
0, -1, 0, 0, 0, 4, 200, 4, 5, 0, -4, -5, -3, 0, 0, -1, -3, -3, 3, 0, 2, 0, -2, 5, -1, 1],
[-2, 0, -2, 5, 0, 0, 1, -5, 5, 0, 3, 4, 0, 0, -4, -4, -1, -2, -4, -4, 0, 1, 0, 5, -5, -4,
3, 5, 1, 0, 4, 200, 5, 4, 5, 3, 4, 1, -1, 0, 1, 5, -3, 3, 1, 0, 0, -3, -4, -1], [4, -5,
-1, 5, 4, 5, 3, -2, 1, -3, 0, 2, 0, 2, 4, -4, 0, 0, -2, 0, -3, -1, 5, 0, 0, 0, 0, 1, -2,
0, 5, 5, 200, -1, -2, 4, 2, 4, -5, -5, -3, 0, 4, -1, 2, 0, 0, 2, 5, 0], [-3, 4, 4, 5, 0,
1, -3, 0, 3, 0, -1, -5, -3, -3, 0, 4, -4, 4, -1, 0, -1, 2, 1, -3, 1, 0, 4, 4, 0, 5, 0, 4,
-1, 200, -3, -5, 3, 0, -3, 2, 4, -4, 1, 5, 0, -1, 2, -4, 5, -2], [2, 0, 1, 2, -5, -2, -2,
-1, -4, -1, -2, 0, 4, 4, 5, 0, 0, 0, 0, 4, 1, 1, 0, -2, 0, -2, -1, 4, 0, -4, -4, 5, -2,
-3, 200, 2, -3, -4, -2, -1, 0, 0, 4, 1, -5, -5, 3, 3, 2, -4], [-1, -1, 0, 1, 1, -5, -3, 0,
5, -1, 2, 0, 0, -1, 0, 5, 3, -2, -1, 0, 0, 0, -2, 0, 0, 2, -1, -3, 0, 0, -5, 3, 4, -5, 2,
200, 0, -1, 0, 4, 5, 0, 5, -3, 4, 0, 0, 5, -4, 1], [0, -1, 0, -4, 3, -4, -2, 0, -2, -1,
-3, -4, 0, -2, -5, 0, -4, 0, 0, -2, 3, 0, 1, 1, 1, 1, 4, -2, -4, 2, -3, 4, 2, 3, -3, 0,
200, 1, -3, 1, 3, 0, -3, -3, 5, 4, 1, -3, -5, -2], [-3, 0, 3, -3, 2, 0, -5, -1, 4, 0, -1,
4, 0, -4, 4, 0, 4, 0, 0, -3, 0, -4, -4, 3, -3, 5, 1, 4, 1, 1, 0, 1, 4, 0, -4, -1, 1, 200,
0, -3, 1, 3, 0, 3, 0, 0, -4, 0, 0, 3], [0, 0, 1, -5, 5, -1, 4, -4, 4, -4, 1, 0, 4, 0, -3,
-5, -5, -2, 3, 5, -4, -1, 0, 0, 0, 5, 1, -5, -5, -1, 0, -1, -5, -3, -2, 0, -3, 0, 200, -5,
0, -4, 0, -4, 3, 3, 1, -1, -2, 5], [-4, -3, 3, 0, 0, 0, 0, 1, 0, 4, -1, 0, -5, 5, 5, -5,
0, -3, 3, 2, 3, -1, 5, 0, 1, 2, 5, 2, 2, 3, -1, 0, -5, 2, -1, 4, 1, -3, -5, 200, -1, 0, 5,
5, 0, 0, 1, 3, 3, 5], [-5, -1, -5, -2, 0, -5, -3, 4, -4, -3, -4, -3, -5, 3, -4, 0, -4, -2,
2, 4, 5, -4, 0, -5, 0, 0, 2, -2, 0, -1, -3, 1, -3, 4, 0, 5, 3, 1, 0, -1, 200, -3, 0, -5,
4, 0, 5, 0, -1, -2], [4, 3, 1, 0, 2, 0, -2, -3, 3, 0, -3, 5, 0, 1, 0, 0, 4, 0, 4, 4, 4, 1,
1, 1, 0, -5, -2, -4, 4, -1, -3, 5, 0, -4, 0, 0, 0, 3, -4, 0, -3, 200, -1, -5, -5, -4, 1,
-1, 2, 0], [1, -1, 4, 0, 4, -5, 0, 3, 0, -5, 2, 3, -4, -4, 4, 1, 1, -5, -5, -1, 2, 0, -4,
4, 5, 3, 0, 0, -1, 5, 3, -3, 4, 1, 4, 5, -3, 0, 0, 5, 0, -1, 200, 0, -3, 0, 4, 0, 3, 3],
[0, -5, 2, 1, -5, -2, 0, 0, 4, 0, -2, 5, 5, 4, 5, 0, 3, 0, 3, 5, 4, 4, -3, 2, -5, -5, 1,
0, 1, 0, 0, 3, -1, 5, 1, -3, -3, 3, -4, 5, -5, -5, 0, 200, 2, 0, 4, 3, 5, 0], [-1, -2, 5,
3, -3, 0, 0, -3, 0, 3, 0, 2, -2, 0, -4, 0, 0, 1, 1, 0, -5, -2, -4, 5, 0, 2, -1, 4, 1, 2,
2, 1, 2, 0, -5, 4, 5, 0, 3, 0, 4, -5, -3, 2, 200, -5, 4, 0, 4, 0], [3, -2, -3, 0, 0, 2,
-2, 0, -3, -4, -2, 0, -4, 1, 0, 1, 2, 0, -4, -4, -4, 5, 4, 0, -3, 2, -1, 1, -4, -3, 0, 0,
0, -1, -5, 0, 4, 0, 3, 0, 0, -4, 0, 0, -5, 200, 0, 0, 4, -4], [3, 1, -2, -5, -2, 1, 0, 0,
-1, 0, 3, 0, -3, 1, 1, -1, -2, 3, 0, 3, 0, 0, 5, 0, 5, -4, 0, 0, 3, 0, -2, 0, 0, 2, 3, 0,
1, -4, 1, 1, 5, 1, 4, 4, 4, 0, 200, 4, -4, 5], [4, -2, 0, 1, -4, 0, -4, 0, -2, 2, 0, 3,
-1, 0, -4, 0, 0, 0, 5, -5, 0, 2, -4, 1, 4, -5, 3, 4, 1, -3, 5, -3, 2, -4, 3, 5, -3, 0, -1,
3, 0, -1, 0, 3, 0, 0, 4, 200, 0, 0], [-1, -2, -5, 0, -3, 3, 3, 3, 0, 2, 2, 3, 1, 0, 4, 0,
-4, 0, -5, 5, 0, -3, -3, 3, 0, -3, 3, 2, -1, -2, -1, -4, 5, 5, 2, -4, -5, 0, -2, 3, -1, 2,
3, 5, 4, 4, -4, 0, 200, 2], [-2, 2, 5, 5, 2, -3, 3, 0, 0, 0, 2, -2, -3, 4, 0, 5, 0, -5, 1,
0, 1, -4, 5, 0, 0, -2, 0, 5, 2, 0, 1, -1, 0, -2, -4, 1, -2, 3, 5, 5, -2, 0, 3, 0, 0, -4,
5, 0, 2, 200]]

b = [ 0.0 for i in range(50)]
for i in range(50):
   for j in range(50):
      b[i] = b[i] + A[i][j]
      
GS(A, b, 100, 10**-15)
```


* Number of operations: 12500 + O*50

## (c)


Repeat parts a and b after changing the diagonal elements in the matrix to 60, i.e., reducing the strictness of
the diagonal dominance. Note that you should be able to use a “find and replace all” feature to change 200
to 60 without having to find and change all of the diagonal elements.


```{python}
import math

def GE_SPP(A, b):
  n = len(A) # if not built-in in a language,
 # read in n as the third argument
  L = [i for i in range(n)] # a quick way to initialize L
  x = [0.0 for i in range(n)]
  S = [0.0 for i in range(n)]
 #build the S vector
  for i in range(n): # find largest element in each    #row of a
    for j in range(n):
      if (abs(A[i][j])>S[i]):
        S[i] = abs(A[i][j])

#could calculate cond = the product of the two-norms of the rows of A here
  p = 1
  for i in range(n):
    s = 0
    for j in range(n):
      s += A[i][j]**2
    p = p *math.sqrt(s)

#forward elimination
  for k in range(n-1):
   # check to see whether we need to swap rows
    R = 0.0
    for i in range(k, n): #from k to n-1
      temp = abs(A[L[i]][k]/S[L[i]])
      if temp > R:
        R = temp
        index = i
 # do the swap
    temp = L[index]
    L[index] = L[k]
    L[k] = temp
 # zero out below the diagonal (update both a and b)
    for i in range(k+1,n):
      xmult = A[L[i]][k]/A[L[k]][k]
      for j in range(k+1,n):
        A[L[i]][j] = A[L[i]][j]-xmult*A[L[k]][j]
      b[L[i]] = b[L[i]]-xmult*b[L[k]]
#could divide cond by the absolute values of the pivots here

  for i in range(n):
    p = p / abs(A[L[i]][i])

#back substitution
  x[n-1] = b[L[n-1]]/A[L[n-1]][n-1]
  for k in range(n-2,-1, -1):
    sum = b[L[k]]
    for j in range(k+1,n):
      sum = sum - A[L[k]][j]*x[j]
    x[k] = sum/A[L[k]][k]

  print("the solution is:", x)
  print("the condition number estimation is", p)
  print(L)
 #also output cond!

#A= [[1, 1, 1, 1, 1], [1, 3, 3**2, 3**3, 3**4], [1, 1, 7, 7**2, 7**3], [1, 1, 1, 10, 10**2], [1,1,1,1,14] ]
#b = [22, 1, -2, 12, -2]

A = [[60, 4, 5, -1, 0, -5, -5, 0, -3, -2, -2, 3, -4, 0, 4, -1, 1, 1, 5, 0, 1, 3, 3, -2,
3, -2, 0, -2, 4, -4, -4, -2, 4, -3, 2, -1, 0, -3, 0, -4, -5, 4, 1, 0, -1, 3, 3, 4, -1,
-2], [4, 60, 4, 1, -5, 0, 0, 3, 1, 4, 2, -1, 2, 0, -5, -2, 4, -2, -2, 0, -2, 0, 4, 2, 5,
-1, 0, 5, 2, 4, -1, 0, -5, 4, 0, -1, -1, 0, 0, -3, -1, 3, -1, -5, -2, -2, 1, -2, -2, 2],
[5, 4, 60, 0, -4, 0, -3, 0, -5, -4, 3, 3, 4, 4, 1, 4, -5, 0, 0, -2, -4, -1, 4, 0, 3, 3,
4, -3, 0, 0, -2, -2, -1, 4, 1, 0, 0, 3, 1, 3, -5, 1, 4, 2, 5, -3, -2, 0, -5, 5], [-1, 1,
0, 60, 5, 0, 0, 0, 5, 1, -2, 0, -3, 4, 0, 0, 3, 4, -1, 2, 3, 0, 0, -1, 1, 4, 0, 4, -4, 0,
3, 5, 5, 5, 2, 1, -4, -3, -5, 0, -2, 0, 0, 1, 3, 0, -5, 1, 0, 5], [0, -5, -4, 5, 60, 0,
-4, -2, 0, -2, -5, 5, 0, -4, 0, 0, 3, -1, 5, 5, -4, 4, 5, 5, 0, 0, -4, -2, -1, 0, -5, 0,
4, 0, -5, 1, 3, 2, 5, 0, 0, 2, 4, -5, -3, 0, -2, -4, -3, 2], [-5, 0, 0, 0, 0, 60, 0, 0,
-4, 0, 2, -5, 0, -5, 1, 0, 2, -1, 2, 0, 0, 3, 0, -2, 1, -1, 0, 3, 2, 3, 3, 0, 5, 1, -2,
-5, -4, 0, -1, 0, -5, 0, -5, -2, 0, 2, 1, 0, 3, -3], [-5, 0, -3, 0, -4, 0, 60, 2, -1, -4,
4, 0, -2, 3, 0, -1, 0, 5, 5, 0, 4, 3, 0, 5, 2, 1, 0, 0, 0, 3, -5, 1, 3, -3, -2, -3, -2,
-5, 4, 0, -3, -2, 0, 0, 0, -2, 0, -4, 3, 3], [0, 3, 0, 0, -2, 0, 2, 60, 0, 0, 0, 2, 0, 3,
3, -1, 3, -3, 0, 1, 2, -3, 1, 0, -4, 3, 0, 5, 1, -3, 1, -5, -2, 0, -1, 0, 0, -1, -4, 1, 4,
-3, 3, 0, -3, 0, 0, 0, 3, 0], [-3, 1, -5, 5, 0, -4, -1, 0, 60, 5, 0, 4, 3, -3, 4, 0, -4,
-4, 0, 0, -4, 0, -3, 0, 0, 1, -5, 5, 1, 3, 1, 5, 1, 3, -4, 5, -2, 4, 4, 0, -4, 3, 0, 4, 0,
-3, -1, -2, 0, 0], [-2, 4, -4, 1, -2, 0, -4, 0, 5, 60, 4, 0, 0, -3, 2, 2, -3, -1, -5, 2,
2, 0, 0, -3, 4, 1, 0, 2, 1, -4, -3, 0, -3, 0, -1, -1, -1, 0, -4, 4, -3, 0, -5, 0, 3, -4,
0, 2, 2, 0], [-2, 2, 3, -2, -5, 2, 4, 0, 0, 4, 60, 0, -4, 0, 0, -5, 0, -3, -3, 1, -3, -1,
1, 2, 0, -5, -4, -5, 0, 0, -3, 3, 0, -1, -2, 2, -3, -1, 1, -1, -4, -3, 2, -2, 0, -2, 3, 0,
2, 2], [3, -1, 3, 0, 5, -5, 0, 2, 4, 0, 0, 60, 4, -5, 4, -2, 0, 3, 3, 0, -4, -3, -4, -1,
-5, -4, 0, 0, 0, -1, 0, 4, 2, -5, 0, 0, -4, 4, 0, 0, -3, 5, 3, 5, 2, 0, 0, 3, 3, -2], [-4,
2, 4, -3, 0, 0, -2, 0, 3, 0, -4, 4, 60, -5, -5, 5, -2, -3, 4, -2, 0, 0, 3, 0, 4, 0, -1,
1, 1, -2, -5, 0, 0, -3, 4, 0, 0, 0, 4, -5, -5, 0, -4, 5, -2, -4, -3, -1, 1, -3], [0, 0, 4,
4, -4, -5, 3, 3, -3, -3, 0, -5, -5, 60, 4, 3, 0, -4, 0, 0, -3, -3, 4, 0, 0, -1, 0, 4, 1,
1, 2, 0, 2, -3, 4, -1, -2, -4, 0, 5, 3, 1, -4, 4, 0, 1, 1, 0, 0, 4], [4, -5, 1, 0, 0, 1,
0, 3, 4, 2, 0, 4, -5, 4, 60, 2, 2, 0, 0, 2, -3, 0, 2, 0, 1, 0, -4, -3, -2, -4, 5, -4, 4,
0, 5, 0, -5, 4, -3, 5, -4, 0, 4, 5, -4, 0, 1, -4, 4, 0], [-1, -2, 4, 0, 0, 0, -1, -1, 0,
2, -5, -2, 5, 3, 2, 60, 5, -3, 0, 1, 0, 3, 0, -2, -5, -4, 3, 2, -4, 1, 0, -4, -4, 4, 0,
5, 0, 0, -5, -5, 0, 0, 1, 0, 0, 1, -1, 0, 0, 5], [1, 4, -5, 3, 3, 2, 0, 3, -4, -3, 0, 0,
-2, 0, 2, 5, 60, 4, 0, 0, 0, 4, -2, -1, 0, -5, -2, 4, 0, 4, 1, -1, 0, -4, 0, 3, -4, 4,
-5, 0, -4, 4, 1, 3, 0, 2, -2, 0, -4, 0], [1, -2, 0, 4, -1, -1, 5, -3, -4, -1, -3, 3, -3,
-4, 0, -3, 4, 60, 1, -4, 0, 1, 2, 0, 0, 0, 0, 0, 0, 0, 0, -2, 0, 4, 0, -2, 0, 0, -2, -3,
-2, 0, -5, 0, 1, 0, 3, 0, 0, -5], [5, -2, 0, -1, 5, 2, 5, 0, 0, -5, -3, 3, 4, 0, 0, 0, 0,
1, 60, -1, -1, -2, 5, 5, -3, 5, -4, -1, 5, 5, 5, -4, -2, -1, 0, -1, 0, 0, 3, 3, 2, 4, -5,
3, 1, -4, 0, 5, -5, 1], [0, 0, -2, 2, 5, 0, 0, 1, 0, 2, 1, 0, -2, 0, 2, 1, 0, -4, -1, 60,
0, -3, -3, 1, 1, -2, 1, -2, 0, 3, -1, -4, 0, 0, 4, 0, -2, -3, 5, 2, 4, 4, -1, 5, 0, -4, 3,
-5, 5, 0], [1, -2, -4, 3, -4, 0, 4, 2, -4, 2, -3, -4, 0, -3, -3, 0, 0, 0, -1, 0, 60, 2,
1, 3, -2, 0, 3, 3, 1, 3, -1, 0, -3, -1, 1, 0, 3, 0, -4, 3, 5, 4, 2, 4, -5, -4, 0, 0, 0,
1], [3, 0, -1, 0, 4, 3, 3, -3, 0, 0, -1, -3, 0, -3, 0, 3, 4, 1, -2, -3, 2, 60, 0, 0, 2,
-5, -3, -3, -2, 1, -4, 1, -1, 2, 1, 0, 0, -4, -1, -1, -4, 1, 0, 4, -2, 5, 0, 2, -3, -4],
[3, 4, 4, 0, 5, 0, 0, 1, -3, 0, 1, -4, 3, 4, 2, 0, -2, 2, 5, -3, 1, 0, 60, 4, -5, 5, 0,
0, -2, 2, -2, 0, 5, 1, 0, -2, 1, -4, 0, 5, 0, 1, -4, -3, -4, 4, 5, -4, -3, 5], [-2, 2, 0,
-1, 5, -2, 5, 0, 0, -3, 2, -1, 0, 0, 0, -2, -1, 0, 5, 1, 3, 0, 4, 60, 0, 3, 2, 0, 3, 4,
0, 5, 0, -3, -2, 0, 1, 3, 0, 0, -5, 1, 4, 2, 5, 0, 0, 1, 3, 0], [3, 5, 3, 1, 0, 1, 2, -4,
0, 4, 0, -5, 4, 0, 1, -5, 0, 0, -3, 1, -2, 2, -5, 0, 60, -1, -3, -4, 0, 3, 0, -5, 0, 1,
0, 0, 1, -3, 0, 1, 0, 0, 5, -5, 0, -3, 5, 4, 0, 0], [-2, -1, 3, 4, 0, -1, 1, 3, 1, 1, -5,
-4, 0, -1, 0, -4, -5, 0, 5, -2, 0, -5, 5, 3, -1, 60, 0, -5, -1, 0, -1, -4, 0, 0, -2, 2,
1, 5, 5, 2, 0, -5, 3, -5, 2, 2, -4, -5, -3, -2], [0, 0, 4, 0, -4, 0, 0, 0, -5, 0, -4, 0,
-1, 0, -4, 3, -2, 0, -4, 1, 3, -3, 0, 2, -3, 0, 60, 1, -4, 0, 0, 3, 0, 4, -1, -1, 4, 1,
1, 5, 2, -2, 0, 1, -1, -1, 0, 3, 3, 0], [-2, 5, -3, 4, -2, 3, 0, 5, 5, 2, -5, 0, 1, 4, -3,
2, 4, 0, -1, -2, 3, -3, 0, 0, -4, -5, 1, 60, 3, 0, 0, 5, 1, 4, 4, -3, -2, 4, -5, 2, -2,
-4, 0, 0, 4, 1, 0, 4, 2, 5], [4, 2, 0, -4, -1, 2, 0, 1, 1, 1, 0, 0, 1, 1, -2, -4, 0, 0, 5,
0, 1, -2, -2, 3, 0, -1, -4, 3, 60, 0, 0, 1, -2, 0, 0, 0, -4, 1, -5, 2, 0, 4, -1, 1, 1,
-4, 3, 1, -1, 2], [-4, 4, 0, 0, 0, 3, 3, -3, 3, -4, 0, -1, -2, 1, -4, 1, 4, 0, 5, 3, 3, 1,
2, 4, 3, 0, 0, 0, 0, 60, 4, 0, 0, 5, -4, 0, 2, 1, -1, 3, -1, -1, 5, 0, 2, -3, 0, -3, -2,
0], [-4, -1, -2, 3, -5, 3, -5, 1, 1, -3, -3, 0, -5, 2, 5, 0, 1, 0, 5, -1, -1, -4, -2, 0,
0, -1, 0, 0, 0, 4, 60, 4, 5, 0, -4, -5, -3, 0, 0, -1, -3, -3, 3, 0, 2, 0, -2, 5, -1, 1],
[-2, 0, -2, 5, 0, 0, 1, -5, 5, 0, 3, 4, 0, 0, -4, -4, -1, -2, -4, -4, 0, 1, 0, 5, -5, -4,
3, 5, 1, 0, 4, 60, 5, 4, 5, 3, 4, 1, -1, 0, 1, 5, -3, 3, 1, 0, 0, -3, -4, -1], [4, -5,
-1, 5, 4, 5, 3, -2, 1, -3, 0, 2, 0, 2, 4, -4, 0, 0, -2, 0, -3, -1, 5, 0, 0, 0, 0, 1, -2,
0, 5, 5, 60, -1, -2, 4, 2, 4, -5, -5, -3, 0, 4, -1, 2, 0, 0, 2, 5, 0], [-3, 4, 4, 5, 0,
1, -3, 0, 3, 0, -1, -5, -3, -3, 0, 4, -4, 4, -1, 0, -1, 2, 1, -3, 1, 0, 4, 4, 0, 5, 0, 4,
-1, 60, -3, -5, 3, 0, -3, 2, 4, -4, 1, 5, 0, -1, 2, -4, 5, -2], [2, 0, 1, 2, -5, -2, -2,
-1, -4, -1, -2, 0, 4, 4, 5, 0, 0, 0, 0, 4, 1, 1, 0, -2, 0, -2, -1, 4, 0, -4, -4, 5, -2,
-3, 60, 2, -3, -4, -2, -1, 0, 0, 4, 1, -5, -5, 3, 3, 2, -4], [-1, -1, 0, 1, 1, -5, -3, 0,
5, -1, 2, 0, 0, -1, 0, 5, 3, -2, -1, 0, 0, 0, -2, 0, 0, 2, -1, -3, 0, 0, -5, 3, 4, -5, 2,
60, 0, -1, 0, 4, 5, 0, 5, -3, 4, 0, 0, 5, -4, 1], [0, -1, 0, -4, 3, -4, -2, 0, -2, -1,
-3, -4, 0, -2, -5, 0, -4, 0, 0, -2, 3, 0, 1, 1, 1, 1, 4, -2, -4, 2, -3, 4, 2, 3, -3, 0,
60, 1, -3, 1, 3, 0, -3, -3, 5, 4, 1, -3, -5, -2], [-3, 0, 3, -3, 2, 0, -5, -1, 4, 0, -1,
4, 0, -4, 4, 0, 4, 0, 0, -3, 0, -4, -4, 3, -3, 5, 1, 4, 1, 1, 0, 1, 4, 0, -4, -1, 1, 60,
0, -3, 1, 3, 0, 3, 0, 0, -4, 0, 0, 3], [0, 0, 1, -5, 5, -1, 4, -4, 4, -4, 1, 0, 4, 0, -3,
-5, -5, -2, 3, 5, -4, -1, 0, 0, 0, 5, 1, -5, -5, -1, 0, -1, -5, -3, -2, 0, -3, 0, 60, -5,
0, -4, 0, -4, 3, 3, 1, -1, -2, 5], [-4, -3, 3, 0, 0, 0, 0, 1, 0, 4, -1, 0, -5, 5, 5, -5,
0, -3, 3, 2, 3, -1, 5, 0, 1, 2, 5, 2, 2, 3, -1, 0, -5, 2, -1, 4, 1, -3, -5, 60, -1, 0, 5,
5, 0, 0, 1, 3, 3, 5], [-5, -1, -5, -2, 0, -5, -3, 4, -4, -3, -4, -3, -5, 3, -4, 0, -4, -2,
2, 4, 5, -4, 0, -5, 0, 0, 2, -2, 0, -1, -3, 1, -3, 4, 0, 5, 3, 1, 0, -1, 60, -3, 0, -5,
4, 0, 5, 0, -1, -2], [4, 3, 1, 0, 2, 0, -2, -3, 3, 0, -3, 5, 0, 1, 0, 0, 4, 0, 4, 4, 4, 1,
1, 1, 0, -5, -2, -4, 4, -1, -3, 5, 0, -4, 0, 0, 0, 3, -4, 0, -3, 60, -1, -5, -5, -4, 1,
-1, 2, 0], [1, -1, 4, 0, 4, -5, 0, 3, 0, -5, 2, 3, -4, -4, 4, 1, 1, -5, -5, -1, 2, 0, -4,
4, 5, 3, 0, 0, -1, 5, 3, -3, 4, 1, 4, 5, -3, 0, 0, 5, 0, -1, 60, 0, -3, 0, 4, 0, 3, 3],
[0, -5, 2, 1, -5, -2, 0, 0, 4, 0, -2, 5, 5, 4, 5, 0, 3, 0, 3, 5, 4, 4, -3, 2, -5, -5, 1,
0, 1, 0, 0, 3, -1, 5, 1, -3, -3, 3, -4, 5, -5, -5, 0, 60, 2, 0, 4, 3, 5, 0], [-1, -2, 5,
3, -3, 0, 0, -3, 0, 3, 0, 2, -2, 0, -4, 0, 0, 1, 1, 0, -5, -2, -4, 5, 0, 2, -1, 4, 1, 2,
2, 1, 2, 0, -5, 4, 5, 0, 3, 0, 4, -5, -3, 2, 60, -5, 4, 0, 4, 0], [3, -2, -3, 0, 0, 2,
-2, 0, -3, -4, -2, 0, -4, 1, 0, 1, 2, 0, -4, -4, -4, 5, 4, 0, -3, 2, -1, 1, -4, -3, 0, 0,
0, -1, -5, 0, 4, 0, 3, 0, 0, -4, 0, 0, -5, 60, 0, 0, 4, -4], [3, 1, -2, -5, -2, 1, 0, 0,
-1, 0, 3, 0, -3, 1, 1, -1, -2, 3, 0, 3, 0, 0, 5, 0, 5, -4, 0, 0, 3, 0, -2, 0, 0, 2, 3, 0,
1, -4, 1, 1, 5, 1, 4, 4, 4, 0, 60, 4, -4, 5], [4, -2, 0, 1, -4, 0, -4, 0, -2, 2, 0, 3,
-1, 0, -4, 0, 0, 0, 5, -5, 0, 2, -4, 1, 4, -5, 3, 4, 1, -3, 5, -3, 2, -4, 3, 5, -3, 0, -1,
3, 0, -1, 0, 3, 0, 0, 4, 60, 0, 0], [-1, -2, -5, 0, -3, 3, 3, 3, 0, 2, 2, 3, 1, 0, 4, 0,
-4, 0, -5, 5, 0, -3, -3, 3, 0, -3, 3, 2, -1, -2, -1, -4, 5, 5, 2, -4, -5, 0, -2, 3, -1, 2,
3, 5, 4, 4, -4, 0, 60, 2], [-2, 2, 5, 5, 2, -3, 3, 0, 0, 0, 2, -2, -3, 4, 0, 5, 0, -5, 1,
0, 1, -4, 5, 0, 0, -2, 0, 5, 2, 0, 1, -1, 0, -2, -4, 1, -2, 3, 5, 5, -2, 0, 3, 0, 0, -4,
5, 0, 2, 60]]

b = [ 0.0 for i in range(50)]
for i in range(50):
   for j in range(50):
      b[i] = b[i] + A[i][j]
GE_SPP(A, b)

```

```{python}
import math


def GS(A, b, max_its, tolerance):
  n = len(b)
  x = [0.0 for i in range(n)]
  xnew = [0.0 for i in range(n)]
  for its in range(1, max_its + 1):
    for i in range(0, n):
      s = b[i]
      for j in range(0, i):
        s -= A[i][j]*xnew[j]
      for j in range(i + 1, n):
        s -= A[i][j]*x[j]
      xnew[i] = s/A[i][i]
    print(xnew)

    s = 0
    for i in range(n):
      s += (xnew[i] - x[i])**2
    if math.sqrt(s) < tolerance:
      print("convergence", its)
      return
    else:
      for i in range(n):
        x[i] = xnew[i]


A = [[60, 4, 5, -1, 0, -5, -5, 0, -3, -2, -2, 3, -4, 0, 4, -1, 1, 1, 5, 0, 1, 3, 3, -2,
3, -2, 0, -2, 4, -4, -4, -2, 4, -3, 2, -1, 0, -3, 0, -4, -5, 4, 1, 0, -1, 3, 3, 4, -1,
-2], [4, 60, 4, 1, -5, 0, 0, 3, 1, 4, 2, -1, 2, 0, -5, -2, 4, -2, -2, 0, -2, 0, 4, 2, 5,
-1, 0, 5, 2, 4, -1, 0, -5, 4, 0, -1, -1, 0, 0, -3, -1, 3, -1, -5, -2, -2, 1, -2, -2, 2],
[5, 4, 60, 0, -4, 0, -3, 0, -5, -4, 3, 3, 4, 4, 1, 4, -5, 0, 0, -2, -4, -1, 4, 0, 3, 3,
4, -3, 0, 0, -2, -2, -1, 4, 1, 0, 0, 3, 1, 3, -5, 1, 4, 2, 5, -3, -2, 0, -5, 5], [-1, 1,
0, 60, 5, 0, 0, 0, 5, 1, -2, 0, -3, 4, 0, 0, 3, 4, -1, 2, 3, 0, 0, -1, 1, 4, 0, 4, -4, 0,
3, 5, 5, 5, 2, 1, -4, -3, -5, 0, -2, 0, 0, 1, 3, 0, -5, 1, 0, 5], [0, -5, -4, 5, 60, 0,
-4, -2, 0, -2, -5, 5, 0, -4, 0, 0, 3, -1, 5, 5, -4, 4, 5, 5, 0, 0, -4, -2, -1, 0, -5, 0,
4, 0, -5, 1, 3, 2, 5, 0, 0, 2, 4, -5, -3, 0, -2, -4, -3, 2], [-5, 0, 0, 0, 0, 60, 0, 0,
-4, 0, 2, -5, 0, -5, 1, 0, 2, -1, 2, 0, 0, 3, 0, -2, 1, -1, 0, 3, 2, 3, 3, 0, 5, 1, -2,
-5, -4, 0, -1, 0, -5, 0, -5, -2, 0, 2, 1, 0, 3, -3], [-5, 0, -3, 0, -4, 0, 60, 2, -1, -4,
4, 0, -2, 3, 0, -1, 0, 5, 5, 0, 4, 3, 0, 5, 2, 1, 0, 0, 0, 3, -5, 1, 3, -3, -2, -3, -2,
-5, 4, 0, -3, -2, 0, 0, 0, -2, 0, -4, 3, 3], [0, 3, 0, 0, -2, 0, 2, 60, 0, 0, 0, 2, 0, 3,
3, -1, 3, -3, 0, 1, 2, -3, 1, 0, -4, 3, 0, 5, 1, -3, 1, -5, -2, 0, -1, 0, 0, -1, -4, 1, 4,
-3, 3, 0, -3, 0, 0, 0, 3, 0], [-3, 1, -5, 5, 0, -4, -1, 0, 60, 5, 0, 4, 3, -3, 4, 0, -4,
-4, 0, 0, -4, 0, -3, 0, 0, 1, -5, 5, 1, 3, 1, 5, 1, 3, -4, 5, -2, 4, 4, 0, -4, 3, 0, 4, 0,
-3, -1, -2, 0, 0], [-2, 4, -4, 1, -2, 0, -4, 0, 5, 60, 4, 0, 0, -3, 2, 2, -3, -1, -5, 2,
2, 0, 0, -3, 4, 1, 0, 2, 1, -4, -3, 0, -3, 0, -1, -1, -1, 0, -4, 4, -3, 0, -5, 0, 3, -4,
0, 2, 2, 0], [-2, 2, 3, -2, -5, 2, 4, 0, 0, 4, 60, 0, -4, 0, 0, -5, 0, -3, -3, 1, -3, -1,
1, 2, 0, -5, -4, -5, 0, 0, -3, 3, 0, -1, -2, 2, -3, -1, 1, -1, -4, -3, 2, -2, 0, -2, 3, 0,
2, 2], [3, -1, 3, 0, 5, -5, 0, 2, 4, 0, 0, 60, 4, -5, 4, -2, 0, 3, 3, 0, -4, -3, -4, -1,
-5, -4, 0, 0, 0, -1, 0, 4, 2, -5, 0, 0, -4, 4, 0, 0, -3, 5, 3, 5, 2, 0, 0, 3, 3, -2], [-4,
2, 4, -3, 0, 0, -2, 0, 3, 0, -4, 4, 60, -5, -5, 5, -2, -3, 4, -2, 0, 0, 3, 0, 4, 0, -1,
1, 1, -2, -5, 0, 0, -3, 4, 0, 0, 0, 4, -5, -5, 0, -4, 5, -2, -4, -3, -1, 1, -3], [0, 0, 4,
4, -4, -5, 3, 3, -3, -3, 0, -5, -5, 60, 4, 3, 0, -4, 0, 0, -3, -3, 4, 0, 0, -1, 0, 4, 1,
1, 2, 0, 2, -3, 4, -1, -2, -4, 0, 5, 3, 1, -4, 4, 0, 1, 1, 0, 0, 4], [4, -5, 1, 0, 0, 1,
0, 3, 4, 2, 0, 4, -5, 4, 60, 2, 2, 0, 0, 2, -3, 0, 2, 0, 1, 0, -4, -3, -2, -4, 5, -4, 4,
0, 5, 0, -5, 4, -3, 5, -4, 0, 4, 5, -4, 0, 1, -4, 4, 0], [-1, -2, 4, 0, 0, 0, -1, -1, 0,
2, -5, -2, 5, 3, 2, 60, 5, -3, 0, 1, 0, 3, 0, -2, -5, -4, 3, 2, -4, 1, 0, -4, -4, 4, 0,
5, 0, 0, -5, -5, 0, 0, 1, 0, 0, 1, -1, 0, 0, 5], [1, 4, -5, 3, 3, 2, 0, 3, -4, -3, 0, 0,
-2, 0, 2, 5, 60, 4, 0, 0, 0, 4, -2, -1, 0, -5, -2, 4, 0, 4, 1, -1, 0, -4, 0, 3, -4, 4,
-5, 0, -4, 4, 1, 3, 0, 2, -2, 0, -4, 0], [1, -2, 0, 4, -1, -1, 5, -3, -4, -1, -3, 3, -3,
-4, 0, -3, 4, 60, 1, -4, 0, 1, 2, 0, 0, 0, 0, 0, 0, 0, 0, -2, 0, 4, 0, -2, 0, 0, -2, -3,
-2, 0, -5, 0, 1, 0, 3, 0, 0, -5], [5, -2, 0, -1, 5, 2, 5, 0, 0, -5, -3, 3, 4, 0, 0, 0, 0,
1, 60, -1, -1, -2, 5, 5, -3, 5, -4, -1, 5, 5, 5, -4, -2, -1, 0, -1, 0, 0, 3, 3, 2, 4, -5,
3, 1, -4, 0, 5, -5, 1], [0, 0, -2, 2, 5, 0, 0, 1, 0, 2, 1, 0, -2, 0, 2, 1, 0, -4, -1, 60,
0, -3, -3, 1, 1, -2, 1, -2, 0, 3, -1, -4, 0, 0, 4, 0, -2, -3, 5, 2, 4, 4, -1, 5, 0, -4, 3,
-5, 5, 0], [1, -2, -4, 3, -4, 0, 4, 2, -4, 2, -3, -4, 0, -3, -3, 0, 0, 0, -1, 0, 60, 2,
1, 3, -2, 0, 3, 3, 1, 3, -1, 0, -3, -1, 1, 0, 3, 0, -4, 3, 5, 4, 2, 4, -5, -4, 0, 0, 0,
1], [3, 0, -1, 0, 4, 3, 3, -3, 0, 0, -1, -3, 0, -3, 0, 3, 4, 1, -2, -3, 2, 60, 0, 0, 2,
-5, -3, -3, -2, 1, -4, 1, -1, 2, 1, 0, 0, -4, -1, -1, -4, 1, 0, 4, -2, 5, 0, 2, -3, -4],
[3, 4, 4, 0, 5, 0, 0, 1, -3, 0, 1, -4, 3, 4, 2, 0, -2, 2, 5, -3, 1, 0, 60, 4, -5, 5, 0,
0, -2, 2, -2, 0, 5, 1, 0, -2, 1, -4, 0, 5, 0, 1, -4, -3, -4, 4, 5, -4, -3, 5], [-2, 2, 0,
-1, 5, -2, 5, 0, 0, -3, 2, -1, 0, 0, 0, -2, -1, 0, 5, 1, 3, 0, 4, 60, 0, 3, 2, 0, 3, 4,
0, 5, 0, -3, -2, 0, 1, 3, 0, 0, -5, 1, 4, 2, 5, 0, 0, 1, 3, 0], [3, 5, 3, 1, 0, 1, 2, -4,
0, 4, 0, -5, 4, 0, 1, -5, 0, 0, -3, 1, -2, 2, -5, 0, 60, -1, -3, -4, 0, 3, 0, -5, 0, 1,
0, 0, 1, -3, 0, 1, 0, 0, 5, -5, 0, -3, 5, 4, 0, 0], [-2, -1, 3, 4, 0, -1, 1, 3, 1, 1, -5,
-4, 0, -1, 0, -4, -5, 0, 5, -2, 0, -5, 5, 3, -1, 60, 0, -5, -1, 0, -1, -4, 0, 0, -2, 2,
1, 5, 5, 2, 0, -5, 3, -5, 2, 2, -4, -5, -3, -2], [0, 0, 4, 0, -4, 0, 0, 0, -5, 0, -4, 0,
-1, 0, -4, 3, -2, 0, -4, 1, 3, -3, 0, 2, -3, 0, 60, 1, -4, 0, 0, 3, 0, 4, -1, -1, 4, 1,
1, 5, 2, -2, 0, 1, -1, -1, 0, 3, 3, 0], [-2, 5, -3, 4, -2, 3, 0, 5, 5, 2, -5, 0, 1, 4, -3,
2, 4, 0, -1, -2, 3, -3, 0, 0, -4, -5, 1, 60, 3, 0, 0, 5, 1, 4, 4, -3, -2, 4, -5, 2, -2,
-4, 0, 0, 4, 1, 0, 4, 2, 5], [4, 2, 0, -4, -1, 2, 0, 1, 1, 1, 0, 0, 1, 1, -2, -4, 0, 0, 5,
0, 1, -2, -2, 3, 0, -1, -4, 3, 60, 0, 0, 1, -2, 0, 0, 0, -4, 1, -5, 2, 0, 4, -1, 1, 1,
-4, 3, 1, -1, 2], [-4, 4, 0, 0, 0, 3, 3, -3, 3, -4, 0, -1, -2, 1, -4, 1, 4, 0, 5, 3, 3, 1,
2, 4, 3, 0, 0, 0, 0, 60, 4, 0, 0, 5, -4, 0, 2, 1, -1, 3, -1, -1, 5, 0, 2, -3, 0, -3, -2,
0], [-4, -1, -2, 3, -5, 3, -5, 1, 1, -3, -3, 0, -5, 2, 5, 0, 1, 0, 5, -1, -1, -4, -2, 0,
0, -1, 0, 0, 0, 4, 60, 4, 5, 0, -4, -5, -3, 0, 0, -1, -3, -3, 3, 0, 2, 0, -2, 5, -1, 1],
[-2, 0, -2, 5, 0, 0, 1, -5, 5, 0, 3, 4, 0, 0, -4, -4, -1, -2, -4, -4, 0, 1, 0, 5, -5, -4,
3, 5, 1, 0, 4, 60, 5, 4, 5, 3, 4, 1, -1, 0, 1, 5, -3, 3, 1, 0, 0, -3, -4, -1], [4, -5,
-1, 5, 4, 5, 3, -2, 1, -3, 0, 2, 0, 2, 4, -4, 0, 0, -2, 0, -3, -1, 5, 0, 0, 0, 0, 1, -2,
0, 5, 5, 60, -1, -2, 4, 2, 4, -5, -5, -3, 0, 4, -1, 2, 0, 0, 2, 5, 0], [-3, 4, 4, 5, 0,
1, -3, 0, 3, 0, -1, -5, -3, -3, 0, 4, -4, 4, -1, 0, -1, 2, 1, -3, 1, 0, 4, 4, 0, 5, 0, 4,
-1, 60, -3, -5, 3, 0, -3, 2, 4, -4, 1, 5, 0, -1, 2, -4, 5, -2], [2, 0, 1, 2, -5, -2, -2,
-1, -4, -1, -2, 0, 4, 4, 5, 0, 0, 0, 0, 4, 1, 1, 0, -2, 0, -2, -1, 4, 0, -4, -4, 5, -2,
-3, 60, 2, -3, -4, -2, -1, 0, 0, 4, 1, -5, -5, 3, 3, 2, -4], [-1, -1, 0, 1, 1, -5, -3, 0,
5, -1, 2, 0, 0, -1, 0, 5, 3, -2, -1, 0, 0, 0, -2, 0, 0, 2, -1, -3, 0, 0, -5, 3, 4, -5, 2,
60, 0, -1, 0, 4, 5, 0, 5, -3, 4, 0, 0, 5, -4, 1], [0, -1, 0, -4, 3, -4, -2, 0, -2, -1,
-3, -4, 0, -2, -5, 0, -4, 0, 0, -2, 3, 0, 1, 1, 1, 1, 4, -2, -4, 2, -3, 4, 2, 3, -3, 0,
60, 1, -3, 1, 3, 0, -3, -3, 5, 4, 1, -3, -5, -2], [-3, 0, 3, -3, 2, 0, -5, -1, 4, 0, -1,
4, 0, -4, 4, 0, 4, 0, 0, -3, 0, -4, -4, 3, -3, 5, 1, 4, 1, 1, 0, 1, 4, 0, -4, -1, 1, 60,
0, -3, 1, 3, 0, 3, 0, 0, -4, 0, 0, 3], [0, 0, 1, -5, 5, -1, 4, -4, 4, -4, 1, 0, 4, 0, -3,
-5, -5, -2, 3, 5, -4, -1, 0, 0, 0, 5, 1, -5, -5, -1, 0, -1, -5, -3, -2, 0, -3, 0, 60, -5,
0, -4, 0, -4, 3, 3, 1, -1, -2, 5], [-4, -3, 3, 0, 0, 0, 0, 1, 0, 4, -1, 0, -5, 5, 5, -5,
0, -3, 3, 2, 3, -1, 5, 0, 1, 2, 5, 2, 2, 3, -1, 0, -5, 2, -1, 4, 1, -3, -5, 60, -1, 0, 5,
5, 0, 0, 1, 3, 3, 5], [-5, -1, -5, -2, 0, -5, -3, 4, -4, -3, -4, -3, -5, 3, -4, 0, -4, -2,
2, 4, 5, -4, 0, -5, 0, 0, 2, -2, 0, -1, -3, 1, -3, 4, 0, 5, 3, 1, 0, -1, 60, -3, 0, -5,
4, 0, 5, 0, -1, -2], [4, 3, 1, 0, 2, 0, -2, -3, 3, 0, -3, 5, 0, 1, 0, 0, 4, 0, 4, 4, 4, 1,
1, 1, 0, -5, -2, -4, 4, -1, -3, 5, 0, -4, 0, 0, 0, 3, -4, 0, -3, 60, -1, -5, -5, -4, 1,
-1, 2, 0], [1, -1, 4, 0, 4, -5, 0, 3, 0, -5, 2, 3, -4, -4, 4, 1, 1, -5, -5, -1, 2, 0, -4,
4, 5, 3, 0, 0, -1, 5, 3, -3, 4, 1, 4, 5, -3, 0, 0, 5, 0, -1, 60, 0, -3, 0, 4, 0, 3, 3],
[0, -5, 2, 1, -5, -2, 0, 0, 4, 0, -2, 5, 5, 4, 5, 0, 3, 0, 3, 5, 4, 4, -3, 2, -5, -5, 1,
0, 1, 0, 0, 3, -1, 5, 1, -3, -3, 3, -4, 5, -5, -5, 0, 60, 2, 0, 4, 3, 5, 0], [-1, -2, 5,
3, -3, 0, 0, -3, 0, 3, 0, 2, -2, 0, -4, 0, 0, 1, 1, 0, -5, -2, -4, 5, 0, 2, -1, 4, 1, 2,
2, 1, 2, 0, -5, 4, 5, 0, 3, 0, 4, -5, -3, 2, 60, -5, 4, 0, 4, 0], [3, -2, -3, 0, 0, 2,
-2, 0, -3, -4, -2, 0, -4, 1, 0, 1, 2, 0, -4, -4, -4, 5, 4, 0, -3, 2, -1, 1, -4, -3, 0, 0,
0, -1, -5, 0, 4, 0, 3, 0, 0, -4, 0, 0, -5, 60, 0, 0, 4, -4], [3, 1, -2, -5, -2, 1, 0, 0,
-1, 0, 3, 0, -3, 1, 1, -1, -2, 3, 0, 3, 0, 0, 5, 0, 5, -4, 0, 0, 3, 0, -2, 0, 0, 2, 3, 0,
1, -4, 1, 1, 5, 1, 4, 4, 4, 0, 60, 4, -4, 5], [4, -2, 0, 1, -4, 0, -4, 0, -2, 2, 0, 3,
-1, 0, -4, 0, 0, 0, 5, -5, 0, 2, -4, 1, 4, -5, 3, 4, 1, -3, 5, -3, 2, -4, 3, 5, -3, 0, -1,
3, 0, -1, 0, 3, 0, 0, 4, 60, 0, 0], [-1, -2, -5, 0, -3, 3, 3, 3, 0, 2, 2, 3, 1, 0, 4, 0,
-4, 0, -5, 5, 0, -3, -3, 3, 0, -3, 3, 2, -1, -2, -1, -4, 5, 5, 2, -4, -5, 0, -2, 3, -1, 2,
3, 5, 4, 4, -4, 0, 60, 2], [-2, 2, 5, 5, 2, -3, 3, 0, 0, 0, 2, -2, -3, 4, 0, 5, 0, -5, 1,
0, 1, -4, 5, 0, 0, -2, 0, 5, 2, 0, 1, -1, 0, -2, -4, 1, -2, 3, 5, 5, -2, 0, 3, 0, 0, -4,
5, 0, 2, 60]]

b = [ 0.0 for i in range(50)]
for i in range(50):
   for j in range(50):
      b[i] = b[i] + A[i][j]

GS(A, b, 100, 10**-15)
```



## (d)

Repeat parts a and b after changing the diagonal elements in the matrix to 30, further reducing diagonal
dominance. Would increasing max its result in convergence?

```{python}
import math

def GE_SPP(A, b):
  n = len(A) # if not built-in in a language,
 # read in n as the third argument
  L = [i for i in range(n)] # a quick way to initialize L
  x = [0.0 for i in range(n)]
  S = [0.0 for i in range(n)]
 #build the S vector
  for i in range(n): # find largest element in each    #row of a
    for j in range(n):
      if (abs(A[i][j])>S[i]):
        S[i] = abs(A[i][j])

#could calculate cond = the product of the two-norms of the rows of A here
  p = 1
  for i in range(n):
    s = 0
    for j in range(n):
      s += A[i][j]**2
    p = p *math.sqrt(s)

#forward elimination
  for k in range(n-1):
   # check to see whether we need to swap rows
    R = 0.0
    for i in range(k, n): #from k to n-1
      temp = abs(A[L[i]][k]/S[L[i]])
      if temp > R:
        R = temp
        index = i
 # do the swap
    temp = L[index]
    L[index] = L[k]
    L[k] = temp
 # zero out below the diagonal (update both a and b)
    for i in range(k+1,n):
      xmult = A[L[i]][k]/A[L[k]][k]
      for j in range(k+1,n):
        A[L[i]][j] = A[L[i]][j]-xmult*A[L[k]][j]
      b[L[i]] = b[L[i]]-xmult*b[L[k]]
#could divide cond by the absolute values of the pivots here

  for i in range(n):
    p = p / abs(A[L[i]][i])

#back substitution
  x[n-1] = b[L[n-1]]/A[L[n-1]][n-1]
  for k in range(n-2,-1, -1):
    sum = b[L[k]]
    for j in range(k+1,n):
      sum = sum - A[L[k]][j]*x[j]
    x[k] = sum/A[L[k]][k]

  print("the solution is:", x)
  print("the condition number estimation is", p)
  print(L)
 #also output cond!

#A= [[1, 1, 1, 1, 1], [1, 3, 3**2, 3**3, 3**4], [1, 1, 7, 7**2, 7**3], [1, 1, 1, 10, 10**2], [1,1,1,1,14] ]
#b = [22, 1, -2, 12, -2]

A = [[30, 4, 5, -1, 0, -5, -5, 0, -3, -2, -2, 3, -4, 0, 4, -1, 1, 1, 5, 0, 1, 3, 3, -2,
3, -2, 0, -2, 4, -4, -4, -2, 4, -3, 2, -1, 0, -3, 0, -4, -5, 4, 1, 0, -1, 3, 3, 4, -1,
-2], [4, 30, 4, 1, -5, 0, 0, 3, 1, 4, 2, -1, 2, 0, -5, -2, 4, -2, -2, 0, -2, 0, 4, 2, 5,
-1, 0, 5, 2, 4, -1, 0, -5, 4, 0, -1, -1, 0, 0, -3, -1, 3, -1, -5, -2, -2, 1, -2, -2, 2],
[5, 4, 30, 0, -4, 0, -3, 0, -5, -4, 3, 3, 4, 4, 1, 4, -5, 0, 0, -2, -4, -1, 4, 0, 3, 3,
4, -3, 0, 0, -2, -2, -1, 4, 1, 0, 0, 3, 1, 3, -5, 1, 4, 2, 5, -3, -2, 0, -5, 5], [-1, 1,
0, 30, 5, 0, 0, 0, 5, 1, -2, 0, -3, 4, 0, 0, 3, 4, -1, 2, 3, 0, 0, -1, 1, 4, 0, 4, -4, 0,
3, 5, 5, 5, 2, 1, -4, -3, -5, 0, -2, 0, 0, 1, 3, 0, -5, 1, 0, 5], [0, -5, -4, 5, 30, 0,
-4, -2, 0, -2, -5, 5, 0, -4, 0, 0, 3, -1, 5, 5, -4, 4, 5, 5, 0, 0, -4, -2, -1, 0, -5, 0,
4, 0, -5, 1, 3, 2, 5, 0, 0, 2, 4, -5, -3, 0, -2, -4, -3, 2], [-5, 0, 0, 0, 0, 30, 0, 0,
-4, 0, 2, -5, 0, -5, 1, 0, 2, -1, 2, 0, 0, 3, 0, -2, 1, -1, 0, 3, 2, 3, 3, 0, 5, 1, -2,
-5, -4, 0, -1, 0, -5, 0, -5, -2, 0, 2, 1, 0, 3, -3], [-5, 0, -3, 0, -4, 0, 30, 2, -1, -4,
4, 0, -2, 3, 0, -1, 0, 5, 5, 0, 4, 3, 0, 5, 2, 1, 0, 0, 0, 3, -5, 1, 3, -3, -2, -3, -2,
-5, 4, 0, -3, -2, 0, 0, 0, -2, 0, -4, 3, 3], [0, 3, 0, 0, -2, 0, 2, 30, 0, 0, 0, 2, 0, 3,
3, -1, 3, -3, 0, 1, 2, -3, 1, 0, -4, 3, 0, 5, 1, -3, 1, -5, -2, 0, -1, 0, 0, -1, -4, 1, 4,
-3, 3, 0, -3, 0, 0, 0, 3, 0], [-3, 1, -5, 5, 0, -4, -1, 0, 30, 5, 0, 4, 3, -3, 4, 0, -4,
-4, 0, 0, -4, 0, -3, 0, 0, 1, -5, 5, 1, 3, 1, 5, 1, 3, -4, 5, -2, 4, 4, 0, -4, 3, 0, 4, 0,
-3, -1, -2, 0, 0], [-2, 4, -4, 1, -2, 0, -4, 0, 5, 30, 4, 0, 0, -3, 2, 2, -3, -1, -5, 2,
2, 0, 0, -3, 4, 1, 0, 2, 1, -4, -3, 0, -3, 0, -1, -1, -1, 0, -4, 4, -3, 0, -5, 0, 3, -4,
0, 2, 2, 0], [-2, 2, 3, -2, -5, 2, 4, 0, 0, 4, 30, 0, -4, 0, 0, -5, 0, -3, -3, 1, -3, -1,
1, 2, 0, -5, -4, -5, 0, 0, -3, 3, 0, -1, -2, 2, -3, -1, 1, -1, -4, -3, 2, -2, 0, -2, 3, 0,
2, 2], [3, -1, 3, 0, 5, -5, 0, 2, 4, 0, 0, 30, 4, -5, 4, -2, 0, 3, 3, 0, -4, -3, -4, -1,
-5, -4, 0, 0, 0, -1, 0, 4, 2, -5, 0, 0, -4, 4, 0, 0, -3, 5, 3, 5, 2, 0, 0, 3, 3, -2], [-4,
2, 4, -3, 0, 0, -2, 0, 3, 0, -4, 4, 30, -5, -5, 5, -2, -3, 4, -2, 0, 0, 3, 0, 4, 0, -1,
1, 1, -2, -5, 0, 0, -3, 4, 0, 0, 0, 4, -5, -5, 0, -4, 5, -2, -4, -3, -1, 1, -3], [0, 0, 4,
4, -4, -5, 3, 3, -3, -3, 0, -5, -5, 30, 4, 3, 0, -4, 0, 0, -3, -3, 4, 0, 0, -1, 0, 4, 1,
1, 2, 0, 2, -3, 4, -1, -2, -4, 0, 5, 3, 1, -4, 4, 0, 1, 1, 0, 0, 4], [4, -5, 1, 0, 0, 1,
0, 3, 4, 2, 0, 4, -5, 4, 30, 2, 2, 0, 0, 2, -3, 0, 2, 0, 1, 0, -4, -3, -2, -4, 5, -4, 4,
0, 5, 0, -5, 4, -3, 5, -4, 0, 4, 5, -4, 0, 1, -4, 4, 0], [-1, -2, 4, 0, 0, 0, -1, -1, 0,
2, -5, -2, 5, 3, 2, 30, 5, -3, 0, 1, 0, 3, 0, -2, -5, -4, 3, 2, -4, 1, 0, -4, -4, 4, 0,
5, 0, 0, -5, -5, 0, 0, 1, 0, 0, 1, -1, 0, 0, 5], [1, 4, -5, 3, 3, 2, 0, 3, -4, -3, 0, 0,
-2, 0, 2, 5, 30, 4, 0, 0, 0, 4, -2, -1, 0, -5, -2, 4, 0, 4, 1, -1, 0, -4, 0, 3, -4, 4,
-5, 0, -4, 4, 1, 3, 0, 2, -2, 0, -4, 0], [1, -2, 0, 4, -1, -1, 5, -3, -4, -1, -3, 3, -3,
-4, 0, -3, 4, 30, 1, -4, 0, 1, 2, 0, 0, 0, 0, 0, 0, 0, 0, -2, 0, 4, 0, -2, 0, 0, -2, -3,
-2, 0, -5, 0, 1, 0, 3, 0, 0, -5], [5, -2, 0, -1, 5, 2, 5, 0, 0, -5, -3, 3, 4, 0, 0, 0, 0,
1, 30, -1, -1, -2, 5, 5, -3, 5, -4, -1, 5, 5, 5, -4, -2, -1, 0, -1, 0, 0, 3, 3, 2, 4, -5,
3, 1, -4, 0, 5, -5, 1], [0, 0, -2, 2, 5, 0, 0, 1, 0, 2, 1, 0, -2, 0, 2, 1, 0, -4, -1, 30,
0, -3, -3, 1, 1, -2, 1, -2, 0, 3, -1, -4, 0, 0, 4, 0, -2, -3, 5, 2, 4, 4, -1, 5, 0, -4, 3,
-5, 5, 0], [1, -2, -4, 3, -4, 0, 4, 2, -4, 2, -3, -4, 0, -3, -3, 0, 0, 0, -1, 0, 30, 2,
1, 3, -2, 0, 3, 3, 1, 3, -1, 0, -3, -1, 1, 0, 3, 0, -4, 3, 5, 4, 2, 4, -5, -4, 0, 0, 0,
1], [3, 0, -1, 0, 4, 3, 3, -3, 0, 0, -1, -3, 0, -3, 0, 3, 4, 1, -2, -3, 2, 30, 0, 0, 2,
-5, -3, -3, -2, 1, -4, 1, -1, 2, 1, 0, 0, -4, -1, -1, -4, 1, 0, 4, -2, 5, 0, 2, -3, -4],
[3, 4, 4, 0, 5, 0, 0, 1, -3, 0, 1, -4, 3, 4, 2, 0, -2, 2, 5, -3, 1, 0, 30, 4, -5, 5, 0,
0, -2, 2, -2, 0, 5, 1, 0, -2, 1, -4, 0, 5, 0, 1, -4, -3, -4, 4, 5, -4, -3, 5], [-2, 2, 0,
-1, 5, -2, 5, 0, 0, -3, 2, -1, 0, 0, 0, -2, -1, 0, 5, 1, 3, 0, 4, 30, 0, 3, 2, 0, 3, 4,
0, 5, 0, -3, -2, 0, 1, 3, 0, 0, -5, 1, 4, 2, 5, 0, 0, 1, 3, 0], [3, 5, 3, 1, 0, 1, 2, -4,
0, 4, 0, -5, 4, 0, 1, -5, 0, 0, -3, 1, -2, 2, -5, 0, 30, -1, -3, -4, 0, 3, 0, -5, 0, 1,
0, 0, 1, -3, 0, 1, 0, 0, 5, -5, 0, -3, 5, 4, 0, 0], [-2, -1, 3, 4, 0, -1, 1, 3, 1, 1, -5,
-4, 0, -1, 0, -4, -5, 0, 5, -2, 0, -5, 5, 3, -1, 30, 0, -5, -1, 0, -1, -4, 0, 0, -2, 2,
1, 5, 5, 2, 0, -5, 3, -5, 2, 2, -4, -5, -3, -2], [0, 0, 4, 0, -4, 0, 0, 0, -5, 0, -4, 0,
-1, 0, -4, 3, -2, 0, -4, 1, 3, -3, 0, 2, -3, 0, 30, 1, -4, 0, 0, 3, 0, 4, -1, -1, 4, 1,
1, 5, 2, -2, 0, 1, -1, -1, 0, 3, 3, 0], [-2, 5, -3, 4, -2, 3, 0, 5, 5, 2, -5, 0, 1, 4, -3,
2, 4, 0, -1, -2, 3, -3, 0, 0, -4, -5, 1, 30, 3, 0, 0, 5, 1, 4, 4, -3, -2, 4, -5, 2, -2,
-4, 0, 0, 4, 1, 0, 4, 2, 5], [4, 2, 0, -4, -1, 2, 0, 1, 1, 1, 0, 0, 1, 1, -2, -4, 0, 0, 5,
0, 1, -2, -2, 3, 0, -1, -4, 3, 30, 0, 0, 1, -2, 0, 0, 0, -4, 1, -5, 2, 0, 4, -1, 1, 1,
-4, 3, 1, -1, 2], [-4, 4, 0, 0, 0, 3, 3, -3, 3, -4, 0, -1, -2, 1, -4, 1, 4, 0, 5, 3, 3, 1,
2, 4, 3, 0, 0, 0, 0, 30, 4, 0, 0, 5, -4, 0, 2, 1, -1, 3, -1, -1, 5, 0, 2, -3, 0, -3, -2,
0], [-4, -1, -2, 3, -5, 3, -5, 1, 1, -3, -3, 0, -5, 2, 5, 0, 1, 0, 5, -1, -1, -4, -2, 0,
0, -1, 0, 0, 0, 4, 30, 4, 5, 0, -4, -5, -3, 0, 0, -1, -3, -3, 3, 0, 2, 0, -2, 5, -1, 1],
[-2, 0, -2, 5, 0, 0, 1, -5, 5, 0, 3, 4, 0, 0, -4, -4, -1, -2, -4, -4, 0, 1, 0, 5, -5, -4,
3, 5, 1, 0, 4, 30, 5, 4, 5, 3, 4, 1, -1, 0, 1, 5, -3, 3, 1, 0, 0, -3, -4, -1], [4, -5,
-1, 5, 4, 5, 3, -2, 1, -3, 0, 2, 0, 2, 4, -4, 0, 0, -2, 0, -3, -1, 5, 0, 0, 0, 0, 1, -2,
0, 5, 5, 30, -1, -2, 4, 2, 4, -5, -5, -3, 0, 4, -1, 2, 0, 0, 2, 5, 0], [-3, 4, 4, 5, 0,
1, -3, 0, 3, 0, -1, -5, -3, -3, 0, 4, -4, 4, -1, 0, -1, 2, 1, -3, 1, 0, 4, 4, 0, 5, 0, 4,
-1, 30, -3, -5, 3, 0, -3, 2, 4, -4, 1, 5, 0, -1, 2, -4, 5, -2], [2, 0, 1, 2, -5, -2, -2,
-1, -4, -1, -2, 0, 4, 4, 5, 0, 0, 0, 0, 4, 1, 1, 0, -2, 0, -2, -1, 4, 0, -4, -4, 5, -2,
-3, 30, 2, -3, -4, -2, -1, 0, 0, 4, 1, -5, -5, 3, 3, 2, -4], [-1, -1, 0, 1, 1, -5, -3, 0,
5, -1, 2, 0, 0, -1, 0, 5, 3, -2, -1, 0, 0, 0, -2, 0, 0, 2, -1, -3, 0, 0, -5, 3, 4, -5, 2,
30, 0, -1, 0, 4, 5, 0, 5, -3, 4, 0, 0, 5, -4, 1], [0, -1, 0, -4, 3, -4, -2, 0, -2, -1,
-3, -4, 0, -2, -5, 0, -4, 0, 0, -2, 3, 0, 1, 1, 1, 1, 4, -2, -4, 2, -3, 4, 2, 3, -3, 0,
30, 1, -3, 1, 3, 0, -3, -3, 5, 4, 1, -3, -5, -2], [-3, 0, 3, -3, 2, 0, -5, -1, 4, 0, -1,
4, 0, -4, 4, 0, 4, 0, 0, -3, 0, -4, -4, 3, -3, 5, 1, 4, 1, 1, 0, 1, 4, 0, -4, -1, 1, 30,
0, -3, 1, 3, 0, 3, 0, 0, -4, 0, 0, 3], [0, 0, 1, -5, 5, -1, 4, -4, 4, -4, 1, 0, 4, 0, -3,
-5, -5, -2, 3, 5, -4, -1, 0, 0, 0, 5, 1, -5, -5, -1, 0, -1, -5, -3, -2, 0, -3, 0, 30, -5,
0, -4, 0, -4, 3, 3, 1, -1, -2, 5], [-4, -3, 3, 0, 0, 0, 0, 1, 0, 4, -1, 0, -5, 5, 5, -5,
0, -3, 3, 2, 3, -1, 5, 0, 1, 2, 5, 2, 2, 3, -1, 0, -5, 2, -1, 4, 1, -3, -5, 30, -1, 0, 5,
5, 0, 0, 1, 3, 3, 5], [-5, -1, -5, -2, 0, -5, -3, 4, -4, -3, -4, -3, -5, 3, -4, 0, -4, -2,
2, 4, 5, -4, 0, -5, 0, 0, 2, -2, 0, -1, -3, 1, -3, 4, 0, 5, 3, 1, 0, -1, 30, -3, 0, -5,
4, 0, 5, 0, -1, -2], [4, 3, 1, 0, 2, 0, -2, -3, 3, 0, -3, 5, 0, 1, 0, 0, 4, 0, 4, 4, 4, 1,
1, 1, 0, -5, -2, -4, 4, -1, -3, 5, 0, -4, 0, 0, 0, 3, -4, 0, -3, 30, -1, -5, -5, -4, 1,
-1, 2, 0], [1, -1, 4, 0, 4, -5, 0, 3, 0, -5, 2, 3, -4, -4, 4, 1, 1, -5, -5, -1, 2, 0, -4,
4, 5, 3, 0, 0, -1, 5, 3, -3, 4, 1, 4, 5, -3, 0, 0, 5, 0, -1, 30, 0, -3, 0, 4, 0, 3, 3],
[0, -5, 2, 1, -5, -2, 0, 0, 4, 0, -2, 5, 5, 4, 5, 0, 3, 0, 3, 5, 4, 4, -3, 2, -5, -5, 1,
0, 1, 0, 0, 3, -1, 5, 1, -3, -3, 3, -4, 5, -5, -5, 0, 30, 2, 0, 4, 3, 5, 0], [-1, -2, 5,
3, -3, 0, 0, -3, 0, 3, 0, 2, -2, 0, -4, 0, 0, 1, 1, 0, -5, -2, -4, 5, 0, 2, -1, 4, 1, 2,
2, 1, 2, 0, -5, 4, 5, 0, 3, 0, 4, -5, -3, 2, 30, -5, 4, 0, 4, 0], [3, -2, -3, 0, 0, 2,
-2, 0, -3, -4, -2, 0, -4, 1, 0, 1, 2, 0, -4, -4, -4, 5, 4, 0, -3, 2, -1, 1, -4, -3, 0, 0,
0, -1, -5, 0, 4, 0, 3, 0, 0, -4, 0, 0, -5, 30, 0, 0, 4, -4], [3, 1, -2, -5, -2, 1, 0, 0,
-1, 0, 3, 0, -3, 1, 1, -1, -2, 3, 0, 3, 0, 0, 5, 0, 5, -4, 0, 0, 3, 0, -2, 0, 0, 2, 3, 0,
1, -4, 1, 1, 5, 1, 4, 4, 4, 0, 30, 4, -4, 5], [4, -2, 0, 1, -4, 0, -4, 0, -2, 2, 0, 3,
-1, 0, -4, 0, 0, 0, 5, -5, 0, 2, -4, 1, 4, -5, 3, 4, 1, -3, 5, -3, 2, -4, 3, 5, -3, 0, -1,
3, 0, -1, 0, 3, 0, 0, 4, 30, 0, 0], [-1, -2, -5, 0, -3, 3, 3, 3, 0, 2, 2, 3, 1, 0, 4, 0,
-4, 0, -5, 5, 0, -3, -3, 3, 0, -3, 3, 2, -1, -2, -1, -4, 5, 5, 2, -4, -5, 0, -2, 3, -1, 2,
3, 5, 4, 4, -4, 0, 30, 2], [-2, 2, 5, 5, 2, -3, 3, 0, 0, 0, 2, -2, -3, 4, 0, 5, 0, -5, 1,
0, 1, -4, 5, 0, 0, -2, 0, 5, 2, 0, 1, -1, 0, -2, -4, 1, -2, 3, 5, 5, -2, 0, 3, 0, 0, -4,
5, 0, 2, 30]]

b = [ 0.0 for i in range(50)]
for i in range(50):
   for j in range(50):
      b[i] = b[i] + A[i][j]
GE_SPP(A, b)


```

```{python}
import math


def GS(A, b, max_its, tolerance):
  n = len(b)
  x = [0.0 for i in range(n)]
  xnew = [0.0 for i in range(n)]
  for its in range(1, max_its + 1):
    for i in range(0, n):
      s = b[i]
      for j in range(0, i):
        s -= A[i][j]*xnew[j]
      for j in range(i + 1, n):
        s -= A[i][j]*x[j]
      xnew[i] = s/A[i][i]
    print(xnew)

    s = 0
    for i in range(n):
      s += (xnew[i] - x[i])**2
    if math.sqrt(s) < tolerance:
      print("convergence", its)
      return
    else:
      for i in range(n):
        x[i] = xnew[i]


A = [[30, 4, 5, -1, 0, -5, -5, 0, -3, -2, -2, 3, -4, 0, 4, -1, 1, 1, 5, 0, 1, 3, 3, -2,
3, -2, 0, -2, 4, -4, -4, -2, 4, -3, 2, -1, 0, -3, 0, -4, -5, 4, 1, 0, -1, 3, 3, 4, -1,
-2], [4, 30, 4, 1, -5, 0, 0, 3, 1, 4, 2, -1, 2, 0, -5, -2, 4, -2, -2, 0, -2, 0, 4, 2, 5,
-1, 0, 5, 2, 4, -1, 0, -5, 4, 0, -1, -1, 0, 0, -3, -1, 3, -1, -5, -2, -2, 1, -2, -2, 2],
[5, 4, 30, 0, -4, 0, -3, 0, -5, -4, 3, 3, 4, 4, 1, 4, -5, 0, 0, -2, -4, -1, 4, 0, 3, 3,
4, -3, 0, 0, -2, -2, -1, 4, 1, 0, 0, 3, 1, 3, -5, 1, 4, 2, 5, -3, -2, 0, -5, 5], [-1, 1,
0, 30, 5, 0, 0, 0, 5, 1, -2, 0, -3, 4, 0, 0, 3, 4, -1, 2, 3, 0, 0, -1, 1, 4, 0, 4, -4, 0,
3, 5, 5, 5, 2, 1, -4, -3, -5, 0, -2, 0, 0, 1, 3, 0, -5, 1, 0, 5], [0, -5, -4, 5, 30, 0,
-4, -2, 0, -2, -5, 5, 0, -4, 0, 0, 3, -1, 5, 5, -4, 4, 5, 5, 0, 0, -4, -2, -1, 0, -5, 0,
4, 0, -5, 1, 3, 2, 5, 0, 0, 2, 4, -5, -3, 0, -2, -4, -3, 2], [-5, 0, 0, 0, 0, 30, 0, 0,
-4, 0, 2, -5, 0, -5, 1, 0, 2, -1, 2, 0, 0, 3, 0, -2, 1, -1, 0, 3, 2, 3, 3, 0, 5, 1, -2,
-5, -4, 0, -1, 0, -5, 0, -5, -2, 0, 2, 1, 0, 3, -3], [-5, 0, -3, 0, -4, 0, 30, 2, -1, -4,
4, 0, -2, 3, 0, -1, 0, 5, 5, 0, 4, 3, 0, 5, 2, 1, 0, 0, 0, 3, -5, 1, 3, -3, -2, -3, -2,
-5, 4, 0, -3, -2, 0, 0, 0, -2, 0, -4, 3, 3], [0, 3, 0, 0, -2, 0, 2, 30, 0, 0, 0, 2, 0, 3,
3, -1, 3, -3, 0, 1, 2, -3, 1, 0, -4, 3, 0, 5, 1, -3, 1, -5, -2, 0, -1, 0, 0, -1, -4, 1, 4,
-3, 3, 0, -3, 0, 0, 0, 3, 0], [-3, 1, -5, 5, 0, -4, -1, 0, 30, 5, 0, 4, 3, -3, 4, 0, -4,
-4, 0, 0, -4, 0, -3, 0, 0, 1, -5, 5, 1, 3, 1, 5, 1, 3, -4, 5, -2, 4, 4, 0, -4, 3, 0, 4, 0,
-3, -1, -2, 0, 0], [-2, 4, -4, 1, -2, 0, -4, 0, 5, 30, 4, 0, 0, -3, 2, 2, -3, -1, -5, 2,
2, 0, 0, -3, 4, 1, 0, 2, 1, -4, -3, 0, -3, 0, -1, -1, -1, 0, -4, 4, -3, 0, -5, 0, 3, -4,
0, 2, 2, 0], [-2, 2, 3, -2, -5, 2, 4, 0, 0, 4, 30, 0, -4, 0, 0, -5, 0, -3, -3, 1, -3, -1,
1, 2, 0, -5, -4, -5, 0, 0, -3, 3, 0, -1, -2, 2, -3, -1, 1, -1, -4, -3, 2, -2, 0, -2, 3, 0,
2, 2], [3, -1, 3, 0, 5, -5, 0, 2, 4, 0, 0, 30, 4, -5, 4, -2, 0, 3, 3, 0, -4, -3, -4, -1,
-5, -4, 0, 0, 0, -1, 0, 4, 2, -5, 0, 0, -4, 4, 0, 0, -3, 5, 3, 5, 2, 0, 0, 3, 3, -2], [-4,
2, 4, -3, 0, 0, -2, 0, 3, 0, -4, 4, 30, -5, -5, 5, -2, -3, 4, -2, 0, 0, 3, 0, 4, 0, -1,
1, 1, -2, -5, 0, 0, -3, 4, 0, 0, 0, 4, -5, -5, 0, -4, 5, -2, -4, -3, -1, 1, -3], [0, 0, 4,
4, -4, -5, 3, 3, -3, -3, 0, -5, -5, 30, 4, 3, 0, -4, 0, 0, -3, -3, 4, 0, 0, -1, 0, 4, 1,
1, 2, 0, 2, -3, 4, -1, -2, -4, 0, 5, 3, 1, -4, 4, 0, 1, 1, 0, 0, 4], [4, -5, 1, 0, 0, 1,
0, 3, 4, 2, 0, 4, -5, 4, 30, 2, 2, 0, 0, 2, -3, 0, 2, 0, 1, 0, -4, -3, -2, -4, 5, -4, 4,
0, 5, 0, -5, 4, -3, 5, -4, 0, 4, 5, -4, 0, 1, -4, 4, 0], [-1, -2, 4, 0, 0, 0, -1, -1, 0,
2, -5, -2, 5, 3, 2, 30, 5, -3, 0, 1, 0, 3, 0, -2, -5, -4, 3, 2, -4, 1, 0, -4, -4, 4, 0,
5, 0, 0, -5, -5, 0, 0, 1, 0, 0, 1, -1, 0, 0, 5], [1, 4, -5, 3, 3, 2, 0, 3, -4, -3, 0, 0,
-2, 0, 2, 5, 30, 4, 0, 0, 0, 4, -2, -1, 0, -5, -2, 4, 0, 4, 1, -1, 0, -4, 0, 3, -4, 4,
-5, 0, -4, 4, 1, 3, 0, 2, -2, 0, -4, 0], [1, -2, 0, 4, -1, -1, 5, -3, -4, -1, -3, 3, -3,
-4, 0, -3, 4, 30, 1, -4, 0, 1, 2, 0, 0, 0, 0, 0, 0, 0, 0, -2, 0, 4, 0, -2, 0, 0, -2, -3,
-2, 0, -5, 0, 1, 0, 3, 0, 0, -5], [5, -2, 0, -1, 5, 2, 5, 0, 0, -5, -3, 3, 4, 0, 0, 0, 0,
1, 30, -1, -1, -2, 5, 5, -3, 5, -4, -1, 5, 5, 5, -4, -2, -1, 0, -1, 0, 0, 3, 3, 2, 4, -5,
3, 1, -4, 0, 5, -5, 1], [0, 0, -2, 2, 5, 0, 0, 1, 0, 2, 1, 0, -2, 0, 2, 1, 0, -4, -1, 30,
0, -3, -3, 1, 1, -2, 1, -2, 0, 3, -1, -4, 0, 0, 4, 0, -2, -3, 5, 2, 4, 4, -1, 5, 0, -4, 3,
-5, 5, 0], [1, -2, -4, 3, -4, 0, 4, 2, -4, 2, -3, -4, 0, -3, -3, 0, 0, 0, -1, 0, 30, 2,
1, 3, -2, 0, 3, 3, 1, 3, -1, 0, -3, -1, 1, 0, 3, 0, -4, 3, 5, 4, 2, 4, -5, -4, 0, 0, 0,
1], [3, 0, -1, 0, 4, 3, 3, -3, 0, 0, -1, -3, 0, -3, 0, 3, 4, 1, -2, -3, 2, 30, 0, 0, 2,
-5, -3, -3, -2, 1, -4, 1, -1, 2, 1, 0, 0, -4, -1, -1, -4, 1, 0, 4, -2, 5, 0, 2, -3, -4],
[3, 4, 4, 0, 5, 0, 0, 1, -3, 0, 1, -4, 3, 4, 2, 0, -2, 2, 5, -3, 1, 0, 30, 4, -5, 5, 0,
0, -2, 2, -2, 0, 5, 1, 0, -2, 1, -4, 0, 5, 0, 1, -4, -3, -4, 4, 5, -4, -3, 5], [-2, 2, 0,
-1, 5, -2, 5, 0, 0, -3, 2, -1, 0, 0, 0, -2, -1, 0, 5, 1, 3, 0, 4, 30, 0, 3, 2, 0, 3, 4,
0, 5, 0, -3, -2, 0, 1, 3, 0, 0, -5, 1, 4, 2, 5, 0, 0, 1, 3, 0], [3, 5, 3, 1, 0, 1, 2, -4,
0, 4, 0, -5, 4, 0, 1, -5, 0, 0, -3, 1, -2, 2, -5, 0, 30, -1, -3, -4, 0, 3, 0, -5, 0, 1,
0, 0, 1, -3, 0, 1, 0, 0, 5, -5, 0, -3, 5, 4, 0, 0], [-2, -1, 3, 4, 0, -1, 1, 3, 1, 1, -5,
-4, 0, -1, 0, -4, -5, 0, 5, -2, 0, -5, 5, 3, -1, 30, 0, -5, -1, 0, -1, -4, 0, 0, -2, 2,
1, 5, 5, 2, 0, -5, 3, -5, 2, 2, -4, -5, -3, -2], [0, 0, 4, 0, -4, 0, 0, 0, -5, 0, -4, 0,
-1, 0, -4, 3, -2, 0, -4, 1, 3, -3, 0, 2, -3, 0, 30, 1, -4, 0, 0, 3, 0, 4, -1, -1, 4, 1,
1, 5, 2, -2, 0, 1, -1, -1, 0, 3, 3, 0], [-2, 5, -3, 4, -2, 3, 0, 5, 5, 2, -5, 0, 1, 4, -3,
2, 4, 0, -1, -2, 3, -3, 0, 0, -4, -5, 1, 30, 3, 0, 0, 5, 1, 4, 4, -3, -2, 4, -5, 2, -2,
-4, 0, 0, 4, 1, 0, 4, 2, 5], [4, 2, 0, -4, -1, 2, 0, 1, 1, 1, 0, 0, 1, 1, -2, -4, 0, 0, 5,
0, 1, -2, -2, 3, 0, -1, -4, 3, 30, 0, 0, 1, -2, 0, 0, 0, -4, 1, -5, 2, 0, 4, -1, 1, 1,
-4, 3, 1, -1, 2], [-4, 4, 0, 0, 0, 3, 3, -3, 3, -4, 0, -1, -2, 1, -4, 1, 4, 0, 5, 3, 3, 1,
2, 4, 3, 0, 0, 0, 0, 30, 4, 0, 0, 5, -4, 0, 2, 1, -1, 3, -1, -1, 5, 0, 2, -3, 0, -3, -2,
0], [-4, -1, -2, 3, -5, 3, -5, 1, 1, -3, -3, 0, -5, 2, 5, 0, 1, 0, 5, -1, -1, -4, -2, 0,
0, -1, 0, 0, 0, 4, 30, 4, 5, 0, -4, -5, -3, 0, 0, -1, -3, -3, 3, 0, 2, 0, -2, 5, -1, 1],
[-2, 0, -2, 5, 0, 0, 1, -5, 5, 0, 3, 4, 0, 0, -4, -4, -1, -2, -4, -4, 0, 1, 0, 5, -5, -4,
3, 5, 1, 0, 4, 30, 5, 4, 5, 3, 4, 1, -1, 0, 1, 5, -3, 3, 1, 0, 0, -3, -4, -1], [4, -5,
-1, 5, 4, 5, 3, -2, 1, -3, 0, 2, 0, 2, 4, -4, 0, 0, -2, 0, -3, -1, 5, 0, 0, 0, 0, 1, -2,
0, 5, 5, 30, -1, -2, 4, 2, 4, -5, -5, -3, 0, 4, -1, 2, 0, 0, 2, 5, 0], [-3, 4, 4, 5, 0,
1, -3, 0, 3, 0, -1, -5, -3, -3, 0, 4, -4, 4, -1, 0, -1, 2, 1, -3, 1, 0, 4, 4, 0, 5, 0, 4,
-1, 30, -3, -5, 3, 0, -3, 2, 4, -4, 1, 5, 0, -1, 2, -4, 5, -2], [2, 0, 1, 2, -5, -2, -2,
-1, -4, -1, -2, 0, 4, 4, 5, 0, 0, 0, 0, 4, 1, 1, 0, -2, 0, -2, -1, 4, 0, -4, -4, 5, -2,
-3, 30, 2, -3, -4, -2, -1, 0, 0, 4, 1, -5, -5, 3, 3, 2, -4], [-1, -1, 0, 1, 1, -5, -3, 0,
5, -1, 2, 0, 0, -1, 0, 5, 3, -2, -1, 0, 0, 0, -2, 0, 0, 2, -1, -3, 0, 0, -5, 3, 4, -5, 2,
30, 0, -1, 0, 4, 5, 0, 5, -3, 4, 0, 0, 5, -4, 1], [0, -1, 0, -4, 3, -4, -2, 0, -2, -1,
-3, -4, 0, -2, -5, 0, -4, 0, 0, -2, 3, 0, 1, 1, 1, 1, 4, -2, -4, 2, -3, 4, 2, 3, -3, 0,
30, 1, -3, 1, 3, 0, -3, -3, 5, 4, 1, -3, -5, -2], [-3, 0, 3, -3, 2, 0, -5, -1, 4, 0, -1,
4, 0, -4, 4, 0, 4, 0, 0, -3, 0, -4, -4, 3, -3, 5, 1, 4, 1, 1, 0, 1, 4, 0, -4, -1, 1, 30,
0, -3, 1, 3, 0, 3, 0, 0, -4, 0, 0, 3], [0, 0, 1, -5, 5, -1, 4, -4, 4, -4, 1, 0, 4, 0, -3,
-5, -5, -2, 3, 5, -4, -1, 0, 0, 0, 5, 1, -5, -5, -1, 0, -1, -5, -3, -2, 0, -3, 0, 30, -5,
0, -4, 0, -4, 3, 3, 1, -1, -2, 5], [-4, -3, 3, 0, 0, 0, 0, 1, 0, 4, -1, 0, -5, 5, 5, -5,
0, -3, 3, 2, 3, -1, 5, 0, 1, 2, 5, 2, 2, 3, -1, 0, -5, 2, -1, 4, 1, -3, -5, 30, -1, 0, 5,
5, 0, 0, 1, 3, 3, 5], [-5, -1, -5, -2, 0, -5, -3, 4, -4, -3, -4, -3, -5, 3, -4, 0, -4, -2,
2, 4, 5, -4, 0, -5, 0, 0, 2, -2, 0, -1, -3, 1, -3, 4, 0, 5, 3, 1, 0, -1, 30, -3, 0, -5,
4, 0, 5, 0, -1, -2], [4, 3, 1, 0, 2, 0, -2, -3, 3, 0, -3, 5, 0, 1, 0, 0, 4, 0, 4, 4, 4, 1,
1, 1, 0, -5, -2, -4, 4, -1, -3, 5, 0, -4, 0, 0, 0, 3, -4, 0, -3, 30, -1, -5, -5, -4, 1,
-1, 2, 0], [1, -1, 4, 0, 4, -5, 0, 3, 0, -5, 2, 3, -4, -4, 4, 1, 1, -5, -5, -1, 2, 0, -4,
4, 5, 3, 0, 0, -1, 5, 3, -3, 4, 1, 4, 5, -3, 0, 0, 5, 0, -1, 30, 0, -3, 0, 4, 0, 3, 3],
[0, -5, 2, 1, -5, -2, 0, 0, 4, 0, -2, 5, 5, 4, 5, 0, 3, 0, 3, 5, 4, 4, -3, 2, -5, -5, 1,
0, 1, 0, 0, 3, -1, 5, 1, -3, -3, 3, -4, 5, -5, -5, 0, 30, 2, 0, 4, 3, 5, 0], [-1, -2, 5,
3, -3, 0, 0, -3, 0, 3, 0, 2, -2, 0, -4, 0, 0, 1, 1, 0, -5, -2, -4, 5, 0, 2, -1, 4, 1, 2,
2, 1, 2, 0, -5, 4, 5, 0, 3, 0, 4, -5, -3, 2, 30, -5, 4, 0, 4, 0], [3, -2, -3, 0, 0, 2,
-2, 0, -3, -4, -2, 0, -4, 1, 0, 1, 2, 0, -4, -4, -4, 5, 4, 0, -3, 2, -1, 1, -4, -3, 0, 0,
0, -1, -5, 0, 4, 0, 3, 0, 0, -4, 0, 0, -5, 30, 0, 0, 4, -4], [3, 1, -2, -5, -2, 1, 0, 0,
-1, 0, 3, 0, -3, 1, 1, -1, -2, 3, 0, 3, 0, 0, 5, 0, 5, -4, 0, 0, 3, 0, -2, 0, 0, 2, 3, 0,
1, -4, 1, 1, 5, 1, 4, 4, 4, 0, 30, 4, -4, 5], [4, -2, 0, 1, -4, 0, -4, 0, -2, 2, 0, 3,
-1, 0, -4, 0, 0, 0, 5, -5, 0, 2, -4, 1, 4, -5, 3, 4, 1, -3, 5, -3, 2, -4, 3, 5, -3, 0, -1,
3, 0, -1, 0, 3, 0, 0, 4, 30, 0, 0], [-1, -2, -5, 0, -3, 3, 3, 3, 0, 2, 2, 3, 1, 0, 4, 0,
-4, 0, -5, 5, 0, -3, -3, 3, 0, -3, 3, 2, -1, -2, -1, -4, 5, 5, 2, -4, -5, 0, -2, 3, -1, 2,
3, 5, 4, 4, -4, 0, 30, 2], [-2, 2, 5, 5, 2, -3, 3, 0, 0, 0, 2, -2, -3, 4, 0, 5, 0, -5, 1,
0, 1, -4, 5, 0, 0, -2, 0, 5, 2, 0, 1, -1, 0, -2, -4, 1, -2, 3, 5, 5, -2, 0, 3, 0, 0, -4,
5, 0, 2, 30]]

b = [ 0.0 for i in range(50)]
for i in range(50):
   for j in range(50):
      b[i] = b[i] + A[i][j]

GS(A, b, 100, 10**-14)
```

* Would increasing max its result in convergence? It results in overflow as the numbers do not converge and increase towards infinity. I've tested this with a max_its of 10000.

## Conclusion

Discuss your findings in the form of a general recommendation regarding which algorithm to use when

* When dealing with larger and larger n values, its going to be better using an iterative technique since the order of magnitude of a GE_SPP model is a cubic for the number of operations while A GS or even a Jacobi would only be a squared term. In cases of small n, it can be the case that an iterative technique would have more operations but these aren't issues with small n values. 

* We can see that as we reduce the diagonal dominance of the matrix, both models start to get worst in their results. For the GE_SPP for example, its condition number starts to increase several magnitude higher than the previous. This condition number means that the error adjusted for those numbers will skew the result. It's the case that we convergence with a diagonal of 30 but its condition number is too high to assess anything meaning from the result.

* The Iteration technique on the otherhand could not even converge for a diagonal of 30, but had no issues with 200 or 60. Those converging in less than 100 iteration. A downside of the iterative technique is the need to add in the maximum iteration and a tolerance compare to the GE_SPP that attempts to just be precise as possible. The user would need the scope of the solution in the iterative case. Aside from that and the fact that GE_SPP converge while GS did not for 30, while not a meaning answer, at the very least does not result in an overflow. GS is often the case, the model one would want to use. 

# Special Cases of Linear Systems: DE Boundary Value Problems.

Some common applications of linear
systems result in a system of equations in which the matrix of coefficients is “tri-diagonal,” i.e., the only non-zero
elements are the diagonal elements and the adjacent elements on either side of the diagonal:



$$\begin{bmatrix}
d_0     & c_0     &      &  &                &\\
a_0     & d_1     & c_1 &         &  &\\
  & a_1 & d_2 & c_3 \\
 &    &      \ddots      & \ddots & \ddots &\\
          &             &      & a_{n-3} & d_{n-2} & c_{n-2}\\
          &          &           &    & a_{n-2} & d_{n-1}
\end{bmatrix}$$

If the diagonal elements are larger than the off-diagonal elements (again, often the case), we could get by with a whittled down version of naive Gaussian elimination that only stores the vectors $\vec a, \vec d, \vec c, \vec b$ and does many fewer operations because of the large number of zeros in the matrix (see page 105 in the text!). You will first write this greatly simplified ``Tridiagonal'' code.

In particular, this tridiagonal format arises in the solution of a second order two point linear boundary value problem that is linear in the dependent variable, $x$ such as:
$$ x''(t) = x(t) + x'(t) + e^t, x(0) = 0, x(2) = 0.$$

We can solve for values of $x$ between $t=0$ and $t=2$, by breaking the interval into evenly spaced points. Let


$$h = \frac{b-a}{n}$$ 
$$t_i = a + i h, \hskip0.03in 0 \le i \le n$$

Replace the derivatives with derivative approximations (more about these in a couple of weeks!):
$$x'' = \frac{x_{i-1} - 2x_i + x_{i+1}}{h^2} \hskip0.1in  x' = \frac{x_{i+1} - x_{i-1}}{2h}$$

Section 11.2 in our book goes through the algebra needed to determine the formulas for the coefficients and right hand sides of the system of equations:
$$d_i = 2 + h^2, 0 \le i \le n-2$$
$$a_i = - \left( 1+ \frac{h}{2} \right); c_i =  - \left( 1- \frac{h}{2} \right), 0 \le i \le n-3$$
$$b_i = -h^2 e^{t_{i+1}}, 0 \le i \le n-2$$
$$ x_0 = x_{n} = 0$$

Write code to build the vectors $a$, $c$, $d$, $b$, and use your tridiagonal code to solve the matrix equation,
dividing the interval (0,2) into $n=8$ equal subintervals (i.e., $h = 0.25$ with 9 points numbered 0 through 8).


```{python}
def tri(a, b, c, d):
  n = len(b)
  x = [0.0 for i in range(n)]

  for i in range(1, n):
    xmult = a[i - 1] / d[i - 1]
    d[i] = d[i] - xmult * c[i - 1]
    b[i] = b[i] - xmult * b[i - 1]
  x[n-1] = b[n-1] / d[n-1]
  for i in range(n - 2, -1, -1):
    x[i] = (b[i] - c[i] * x[i + 1]) / d[i]
  print(x)
  
d = [4,4,4,4,4]
a = [1,1,1,1]
# a = [[4,1,0,0,0], [1,4,1,0,0],[0,1,4,1,0],[0,0,1,4,1],[0,0,0,1,4]]
c = [1,1,1,1]
b = [5,6,6,6,5]

#main idea. I have 8 steps between 0 and 2 
# t array, .25,.5,.75.,...,1.75
#d = 7, a = 6, b = h^2 *e^t ?

# q1: control F, search and replace, replit:explain what happens gauss seidel.
tri(a,b,c,d)

```

```{python}
b = 2
a = 0
n = 8
h = (b - a) / n

import math
a = - (1 + h / 2)
c = - (1 - h / 2)
b = - h ** 2
d = 2 + h**2

a = [a, a, a, a, a, a]
b = [b * math.exp(.25), b * math.exp(.5), b * math.exp(.75), b * math.exp(1),b * math.exp(1.25), b * math.exp(1.5), b * math.exp(1.75)]
c = [c, c, c, c, c, c]
d = [d, d, d,d,d,d,d]

def tri(a, b, c, d):
  n = len(b)
  x = [0.0 for i in range(n)]

  for i in range(1, n):
    xmult = a[i - 1] / d[i - 1]
    d[i] = d[i] - xmult * c[i - 1]
    b[i] = b[i] - xmult * b[i - 1]
  x[n-1] = b[n-1] / d[n-1]
  for i in range(n - 2, -1, -1):
    x[i] = (b[i] - c[i] * x[i + 1]) / d[i]
  print(x)

tri(a, b ,c, d)
```


Method:

* The pseudocode for the tridiagonal was taken from the book and most of the issue is converting the range index. Its a simplied version of a GE_SPP algorithm but the magnitude of its operations is just n rather than n^3 because we're ignoring the zeros along the other matrix position that is not a, c, d. We're trying to solve $Ax = b$, where x is our target and we're only looking at the main diagonal of A and its two adjacent diagonals, respectively c, d. So we first find the length of the main diagonal as we will be looping through it, then create an index for X to put our final answer. Then we start our for loop at 1 and ending at n - 1, here we calculate a multiplier for the diagonals xmult. We also calculate new values for d, b, and x. Where the diagonal is being replace, this is the forward substitution step. We now start from the bottom of the x values and work our way backwards to compute the final xi, this is the back substitution step. 

* Once the code is done, we want to be sure that it's running properly so we create a problem and a known solution, where its output should be just a list of 1. Then for the particular problem we compute everything normally with the given formulas. The only thing different is our b values where we want to be sure its precise, so opting to use the `math.exp()` value for the full 15 digit accuracy. Our results for the x vectors are negative, as each iteration its trying to find the unique solution in as little operations as possible. How good or stable of these answers, not completely sure. 






