---
title: "Modeling Proficiency Problem"
subtitle: "WCS Inclement Weather Days vs. Total Snowfall"
author: "Jefferson Ong"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment = "")
```

***
```{r}
# Load any packages you need here.
library(tibble)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
library(moderndive)
SchoolWed <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/wcsweatherdays.csv")
```


#### The Problem

All K-12 schools in North Carolina are required to complete a minimum number of instructional days per year. Watauga County Schools starts the school year with calendar that has 180 student days. However, the system typically has several cancelations every year as a result of inclement weather. These days must be made up either during the year (e.g., Saturday school or moving teacher workdays) or by adding days to the end of the year. Even with makeups, the number of days students miss is of concern to many people because they feel canceled days disrupt student learning, especially considering the pressure to perform well on standardized tests and the NC School Report Cards. ([school closing data](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/wcsweatherdays.html))

#### The Analysis

Compute the correlation between each month's number of canceled days and total snowfall. Assuming all of the relationships are linear, which month is most strongly correlated with total snowfall? Which is most weakly correlated?



There is a function for correlation, and you can do them all as one summarize(). You cannot group my month because each month is its own column, you'd have to gather() first. Also, September is not the weakest corr. // "Intercept of 5.7 can be taught of as the baseline number of days that will have snow." No--snow is the predictor, the y-intercept reflects a value of y. Interpretation of slope? Months do not directly come into play in this model since we are looking at total days. // Residuals too far from what? You're also missing a plot. 



```{r}
##
 # filter(!is.na(height)) %>%
  
SchoolWed_gathered <- SchoolWed %>% 
  gather(key = "Months", value = "days_number", aug_days:may_days)


Aug <- SchoolWed_gathered %>%
    filter(Months == "aug_days") %>%
    filter(!is.na(total_snow)) %>%
    get_correlation(formula = total_snow ~ days_number)

Sep <- SchoolWed_gathered %>%
    filter(Months == "sep_days") %>%
    filter(!is.na(total_snow)) %>%
    get_correlation(formula = total_snow ~ days_number)
 
Oct <-  SchoolWed_gathered %>%
    filter(Months == "oct_days") %>%
    filter(!is.na(total_snow)) %>%
    get_correlation(formula = total_snow ~ days_number)
  
Nov <-   SchoolWed_gathered %>%
    filter(Months == "nov_days") %>%
    filter(!is.na(total_snow)) %>%
    get_correlation(formula = total_snow ~ days_number)
   
Dec <-  SchoolWed_gathered %>%
    filter(Months == "dec_days") %>%
    filter(!is.na(total_snow)) %>%
    get_correlation(formula = total_snow ~ days_number)
  

Jan <- SchoolWed_gathered %>%
    filter(Months == "jan_days") %>%
    filter(!is.na(total_snow)) %>%
    get_correlation(formula = total_snow ~ days_number)
 
Feb <-  SchoolWed_gathered %>%
    filter(Months == "feb_days") %>%
    filter(!is.na(total_snow)) %>%
    get_correlation(formula = total_snow ~ days_number)
  
Mar <-   SchoolWed_gathered %>%
    filter(Months == "mar_days") %>%
    filter(!is.na(total_snow)) %>%
    get_correlation(formula = total_snow ~ days_number)
   
Apr <-  SchoolWed_gathered %>%
    filter(Months == "apr_days") %>%
    filter(!is.na(total_snow)) %>%
    get_correlation(formula = total_snow ~ days_number)
  
May <-  SchoolWed_gathered %>%
    filter(Months == "may_days") %>%
    filter(!is.na(total_snow)) %>%
    get_correlation(formula = total_snow ~ days_number)
```

```{r}

Aug
Sep
Oct
Nov
Dec
Jan
Feb
Mar
Apr
May
```


ANSWER: Febuary has the greatest correlation. May is the weakest correlation.

Fit a least squares linear model to examine the relationship between total number of inclement weather days (response variable) and total snowfall (explanatory variable). Display the table of coefficients as well as the model summaries using functions from the `moderndive` package. Interpret the coefficients and R^2^ in the context of the problem. 

```{r}
SchoolWed_gathered_model <- lm(total_days ~ total_snow, data = SchoolWed_gathered)

get_regression_table(SchoolWed_gathered_model)

summary(SchoolWed_gathered_model)
```

ANSWER: total snow is predictor variable. total days is the dependent variable. The summary statistic on the residuals will show a smaller average depending if the model fits the data better. The coefficients are weights to minimize the error created by the model with the residual squared shows how well it fits the model. In terms of the problem, its a terrible model considering the range of the residual. 

Create plots to assess the residuals. Do you see anything potentially problematic?

```{r}
#
get_regression_points(SchoolWed_gathered_model)

ggplot(get_regression_points(SchoolWed_gathered_model), aes(x = total_snow, y = residual)) +
  geom_point() +
  geom_hline(yintercept = 0, col = "blue")

ggplot(get_regression_points(SchoolWed_gathered_model), aes(x = residual)) +
  geom_histogram()
```

ANSWER: Yes, theres residuals that are too far(from zero blue line, where there is no residual difference.) that could be omitted and have a better calculation. Looking at the histogram the residuals are skewed right and not normal. 

Does total snowfall in a given year seem to be the primary reason for the total number of WCS inclement weather cancelations? Support your answer using relevant evidence from your analysis. 

ANSWER: Ehh..hmmm.. not really. They're correlated but it isn't all that strong of a correlation, theres even some months where they are negatively correlated. I believe the cause of the cancellation has to do more with the school and how they take precautions rather than the actual amount of snow that happens in the school year. 

***
```{r}
sessionInfo()
```
