---
title: "MSRR CH02, Sections 1-3 HW"
author: "Jefferson Ong"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment = "")
```

***

**Instructions**

Complete the following problems from _Mathematical Statistics with Resampling and R_. You can access the textbook datasets on our [Class Data Files](https://stat-jet-asu.github.io/Datasets/DatasetList.html) page or from the `R` package `resampledata`. The textbook examples use base `R`code for plotting, but you `ggplot2` for visualizations. You may use `table()` to make tables, but you should use `dplyr` for other summaries. Refer to the case studies in Chapter 1 to add informative titles and axis labels to your plots and to provide context for your answers.

Load the packages and datasets you need here.

```{r}
library(readr)

library(resampledata)

library(ggplot2)
library(dplyr)
```



***
#### Problem 2.1

Note: Test for equality using code, not just visual inspection.

```{r}
x_bar <- c(3, 5, 8, 15, 20, 21)

mean(x_bar)
median(x_bar)

mean(log(x_bar))

median(log(x_bar))

log(mean(x_bar)) == mean(log(x_bar))

log(median(x_bar)) == median(log(x_bar))
```


***
#### Problem 2.2

Note: Test for equality using code, not just visual inspection.

```{r}
x_bar2 <- c(1, 2, 4, 5, 6, 8, 11, 15)

mean(x_bar2)
median(x_bar2)

sqrt(x_bar2)

mean(sqrt(x_bar2))
median(sqrt(x_bar2))

mean(sqrt(x_bar2)) == sqrt(mean(x_bar2))

median(sqrt(x_bar2)) == sqrt(median(x_bar2))
```


***
#### Problem 2.5

Note: In (c), create the table of proportions you need to answer (d) as well as a table of counts.

Part (a)

```{r}
str(GSS2002)

names(GSS2002)
```

```{r}
table(GSS2002$DeathPenalty)

GSS2002 %>%
  ggplot( aes(x = DeathPenalty)) +
  geom_bar()
```






Part (b)

```{r}
table(GSS2002$OwnGun)
summary(GSS2002$OwnGun)
```

ANSWER: The summary shows the NA values while the table does not.

Part (c)

```{r}
table(GSS2002$OwnGun, GSS2002$DeathPenalty)

243 / (243 + 59)

199 / (375 + 199)
```

Part (d)

ANSWER: 80%, yes those who do not own guns don't favor the death penalty, reduced to only ~35%


***
#### Problem 2.6

Note: In (b), create a table of proportions as well as a table of counts to answer the question. In (c), also compute the upper fence for each group to show the upper cutoff for outliers.

Part (a)

```{r}
names( Recidivism)
summary(Recidivism)

str(Recidivism)
```

```{r}
table(Recidivism$Recid)

Recidivism %>%
  ggplot( aes(x = Recid)) +
  geom_bar()
```


Part (b)

```{r}
table(Recidivism$Recid, Recidivism$Age25)

1123 / (1123 + 1954)

4263 / (4263 + 9679)
```

~36 were sent back to prison of those that were under 25. THose voer 25 only ~ 31% were sent back. 

Part (c)

```{r}
ggplot(Recidivism, aes(x = Offense, y = Days)) +
  geom_boxplot()
```

Comparison 1: The felony has a longer whiskers meaning that it has a bit more outliers.

Comparison 2: The median of felony is lower than misdemeanor.

Comparison 3: The Q1 and Q3 of Misdemeanor is much bigger so the data is more spread out. 

Part (d)

```{r}
quantile(Recidivism$Days, na.rm = TRUE)
```

Part (e)

```{r}
x <- rnorm(Recidivism$Age)
plot.ecdf(x)
plot.ecdf(table(Recidivism$Days, Recidivism$Age25)) 
Underage <- Recidivism %>%
  filter(Age == "Under 25")

```

ANSWER: I don't know how to break up Age and how it works with ecdf. Not completely sure what rnorm is doing to it. I remember it just has to do with distributions. I'm guessing the 0, 2, 4 are the different factors in Age which is why theres none in -2 or -4. so at 0 would be everyone under 25 so its at about .4 so 40%, the rest is 60%. Unsure how to look into the 400 days. 


***

```{r}
sessionInfo()
```

