---
title: "MSRR CH02, Sections 4-6 HW"
author: "SOLUTIONS / Jill E. Thomley"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment = "")
```

***

**Instructions**

Complete the following problems from _Mathematical Statistics with Resampling and R_ and additional example problems using the normal distribution functions in `R`. You can access the textbook datasets on our [Class Data Files](https://stat-jet-asu.github.io/Datasets/DatasetList.html) page or from the `R` package `resampledata`. The textbook examples use base `R`code for plotting, but you `ggplot2` for visualizations. You may use `table()` to make tables, but you should use `dplyr` for other summaries. Refer to the case studies in Chapter 1 to add informative titles and axis labels to your plots and to provide context for your answers.

Load the packages and datasets you need here.

```{r}
library(resampledata)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(psych)
```


***
#### Problem 2.7

Part (a)

Note: For summaries, include sample size, mean, standard deviation, five-number summary, skewness, and kurtosis.

```{r}
str(Spruce)

SpruceHt <- Spruce %>%
  mutate(Htchange = Height5 - Height0)
describe(SpruceHt$Htchange)
```

```{r}
SpruceHt <- Spruce %>%
  mutate(Htchange = Height5 - Height0)
Sprucesum <- describe(SpruceHt$Htchange)

Sprucesum
str(Sprucesum)
str(SpruceHt)
```

Part (b)

Note: Also make a density plot and ecdf plot. As we learned to do in class, include theoretical normal reference curves.

```{r}
Sprucedat <- SpruceHt %>%
  mutate(htmean = mean(Htchange), htsd = sd(Htchange))

ggplot(Sprucedat , aes(x = Htchange)) + 
  geom_density() +
  stat_function(fun = dnorm, 
                args = list(Sprucedat$htmean, Sprucedat$htsd),   
                color = "red")

ggplot(Sprucedat, aes(x = Htchange)) + 
  stat_ecdf() +
  stat_function(fun = pnorm, 
                args = list(Sprucedat$htmean, Sprucedat$htsd), 
                color = "red")
```



ANSWER: Yep, looks fairly normal

Part (c)

```{r}
str(Sprucedat)
```


```{r}


ggplot(Sprucedat, aes(x = Fertilizer, y = Di.change)) +
  geom_boxplot()
```

Part (d)

Note: For summaries, include sample size, mean, standard deviation, five-number summary, skewness, and kurtosis. Use `dplyr` functions rather than `tapply()`.


```{r}
Spruce %>%
  summarize(meand = mean(Di.change))

# Spruce %>%
  # group_by(Fertilizer)
  # summarize(n     = n(Di.change)
                      # xbar  = mean(Di.change, na.rm = TRUE),
                     #  stdev = sd(Di.change, na.rm = TRUE),
                     #  min   = fivenum(Di.change)[1],
                     # Q1    = fivenum(Di.change)[2],
                   #   med   = fivenum(Di.change)[3],                  
                    #  Q3    = fivenum(Di.change)[4],
                   #   max   = fivenum(Di.change)[5],
                   #   skew  = skewness(Di.change, na.rm = TRUE),
                   #   kurt  = kurtosis(Di.change, na.rm = TRUE))

# Code doesn't work, says Di.change not found. Not sure why when top code works
```





Part (e)

```{r}
ggplot(Sprucedat, aes(x = Ht.change, y = Di.change)) +
  geom_point()
```

ANSWER: Looks postively linear


***
#### Problem 2.8

Note: Where the problem asks you to "describe the distribution," discuss the shape and possible outliers. Choose an appropriate number of bins for the histograms.

```{r}
str(MobileAds)
```


Part (a)

```{r}
ggplot(MobileAds, aes(x = m.cpc_pre)) +
  geom_histogram()

ggplot(MobileAds, aes(x = m.cpc_post)) +
  geom_histogram()
```

ANSWER: Both are right skewed

Part (b)

Note: Use `mutate` and name the new variable `m.cpc_diff`.

```{r}
MobileAdsB <- MobileAds %>%
  mutate(m.cpc_diff = m.cpc_post - m.cpc_pre)

ggplot(MobileAdsB, aes(x = m.cpc_diff)) +
  geom_histogram()
```

ANSWER: Looks like a noraml distribution

Part (c)

```{r}
ggplot(MobileAdsB, aes(sample = m.cpc_diff)) + 
  stat_qq() +
  stat_qq_line(color = "red")
```

ANSWER: No since it looks like half the data is below and the other is above when it should be more scattered. 


***
#### `R` Problem 1

Let the random variable X have a standard normal distribution.

Part (a) : What is P(X < 1.5)?

```{r}
pnorm( 1.5, 0, 1)
```

Part (b) : What is P(-2.0 < X < 1.5)?

```{r}
pnorm(1.5, 0, 1) - pnorm(-2, 0, 1)  # P(-2 <= X <= -1)
```

Part (c) : What is P(X < -1.75 or X > 1.75)?

```{r}
pnorm(q = -1.75, 0, 1) + pnorm(1.75, 0, 1, lower.tail = F)
```

Part (d) : What value of x satisfies P(X < x) = 0.45?

```{r}
qnorm(.45, 0, 1)
```

Part (e) : What value of x satisfies P(X ≥ x) = 0.35?

```{r}
qnorm(1 - .35, 0, 1)
```

Part (f) : The middle 50% of the area under a standard normal curve lies between what values?

```{r}
x_1 = qnorm(p = .25, mean = 0, sd = 1)

x_2 = qnorm(p = .75, mean = 0, sd = 1)

x_1 
x_2
```


***
#### `R` Problem 2

In Hottsylvania, the mean July midday temperature is 32 °C with a standard deviation of 3 °C. Assume that midday temperatures can be modeled with an approximately normal distribution. July has 31 days.

Part (a) : About how many days should be less than 29.44 °C (about 85 °F)?

```{r}
pnorm(29.44, 32, 3)

0.1967372 * 31
```

A: About 6 days

Part (b) : How many days should have a midday temperature between 35 °C and 38 °C?

```{r}
pnorm(38, 32, 3) - pnorm(35, 32, 3)

0.1359051 * 31
```

A: About 4 days

Part (c) : The midday temperature of the hottest 30% of days should be greater than what temperature?

```{r}
qnorm(.3, 32, 3)
```

A: 30.4268

***
#### `R` Problem 3

Generate a random sample of size 100 from a standard normal distribution. Create a density plot and a normal probability plot with appropriate theoretical reference curves.

```{r}
rand_dat <- tibble(unif_sample = runif(100, 0, 1))

rand_stat <- rand_dat %>%
  summarize(randmean = mean(unif_sample), randsd = sd(unif_sample))

ggplot(rand_dat , aes(x = unif_sample)) + 
  geom_density() +
  stat_function(fun = dnorm, 
                args = list(rand_stat$randmean, rand_stat$randsd),   
                color = "red")

ggplot(rand_dat, aes(x = unif_sample)) + 
  stat_ecdf() +
  stat_function(fun = pnorm, 
                args = list(rand_stat$randmean, rand_stat$randsd), 
                color = "red")
```



***

```{r}
sessionInfo()
```
