---
title: "Normal Proficiency Problem"
subtitle: "Properties of the Theoretical Normal Model"
author: "Jefferson Ong"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment = "")
```

	
Can you show mean = median using theoretical calcs rather than simulation? Consider: how much area is above and below the median in any distribution? For your plot, you simulated uniform data (runif), not normal data.

***
```{r}
# Load any packages and datasets you need here.
library(tibble)
library(tidyverse)
library(dplyr)
library(ggplot2)
```

#### The Problem

The normal distribution with parameters $\mu$ and $\sigma$ is unimodal and perfectly symmetric. 

The Empirical Rule states that the normal distribution also has the following properties.

* approximately 68% of area under the curve is between $\mu \pm 1\sigma$ 
* approximately 95% of area under the curve is between $\mu \pm 2\sigma$ 
* approximately 99.7% of area under the curve is between $\mu \pm 3\sigma$ 

Use the normal distribution functions provided in `R` to answer the following questions.

#### The Analysis

Verify the three components of the Empirical Rule for the standard normal distribution.

```{r}
#
pnorm(1, 0, 1) - pnorm(-1, 0, 1)  # P(-1 <= X <= 1)
pnorm(2, 0, 1) - pnorm(-2, 0, 1)  # P(-2 <= X <= 2)
pnorm(3, 0, 1) - pnorm(-3, 0, 1)  # P(-3 <= X <= 3)
```

In a perfectly symmetric distribution, the median and mean are equal. Verify this for the standard normal distribution. 

```{r}
#
# rand_dat <- tibble(unif_sample = runif(10000, 0, 1))
# 
# rand_stat <- rand_dat %>%
#   summarize(randmean = mean(unif_sample), randsd = sd(unif_sample), randmedian = median(unif_sample))
# 
# rand_stat

n <- 30
min <- 0
max <- 1
X <- runif(n, min, max)


mean(X)


median(X)


```

Calculate the lower and upper boundaries for the middle 90%, 95%, and 99% of the area under a standard normal distribution.

```{r}
#
qnorm(p = .05, mean = 0, sd = 1)
qnorm(p = .05, mean = 0, sd = 1, lower.tail = F)

qnorm(p = .025, mean = 0, sd = 1)
qnorm(p = .025, mean = 0, sd = 1, lower.tail = F)

qnorm(p = .005, mean = 0, sd = 1)
qnorm(p = .005, mean = 0, sd = 1, lower.tail = F)

```

Sample 10,000 values from a standard normal distribution and create a density plot that includes a theoretical reference curve.

```{r}
#
rand_dat <- tibble(rnorm_sample = rnorm(10000, 0, 1))

rand_stat <- rand_dat %>%
  summarize(randmean = mean(rnorm_sample), randsd = sd(rnorm_sample), randmedian = median(rnorm_sample))

ggplot(rand_dat , aes(x = rnorm_sample)) + 
  geom_density() +
  stat_function(fun = dnorm, 
                args = list(rand_stat$randmean, rand_stat$randsd),   
                color = "red")
```

***
```{r}
sessionInfo()
```
