---
title: "Normal Proficiency Problem"
subtitle: "Properties of the Theoretical Normal Model"
author: "Jefferson Ong"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment = "")
```

	
Can you show mean = median using theoretical calcs rather than simulation? Consider: how much area is above and below the median in any distribution? For your plot, you simulated uniform data (runif), not normal data.

***
```{r}
# Load any packages and datasets you need here.
library(tibble)
library(tidyverse)
library(dplyr)
library(ggplot2)
```


In a perfectly symmetric distribution, the median and mean are equal. Verify this for the standard normal distribution. 

```{r}
#
# rand_dat <- tibble(unif_sample = runif(10000, 0, 1))
# 
# rand_stat <- rand_dat %>%
#   summarize(randmean = mean(unif_sample), randsd = sd(unif_sample), randmedian = median(unif_sample))
# 
# rand_stat

n <- 10^4
min <- 0
max <- 1
X <- runif(n, min, max)


mean(X)


median(X)


```


Sample 10,000 values from a standard normal distribution and create a density plot that includes a theoretical reference curve.

```{r}
#
rand_dat <- tibble(rnorm_sample = rnorm(10000, 0, 1))

rand_stat <- rand_dat %>%
  summarize(randmean = mean(rnorm_sample), randsd = sd(rnorm_sample), randmedian = median(rnorm_sample))

ggplot(rand_dat , aes(x = rnorm_sample)) + 
  geom_density() +
  stat_function(fun = dnorm, 
                args = list(rand_stat$randmean, rand_stat$randsd),   
                color = "red")
```

***
```{r}
sessionInfo()
```
