---
title: "quiz1"
author: "Jefferson Ong"
date: "2/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readr)
```
# 1: 

```{r}
weather <- read_csv("weather.csv")
```
# 2

```{r}
str(weather)
```

```{r}
ggplot(weather, aes(x = month, y = hitemp)) +
  geom_boxplot()

boxplot(hitemp ~ month, weather)
```

* This plot tells me that the hitemp is higher for June compare to May but May itself has more variability on how the temperature is.


# 3 

```{r}
str(weather)
```

```{r}
pairs(hitemp ~ day + yestavg + yesthi, weather)

pairs(hitemp ~ yestlo + yestrain + yestwind + yesthumid,weather)
```

```{r}
library("PerformanceAnalytics")
chart.Correlation(weather[c(3, 2:9)], histogram=TRUE, pch=19)
```
Explain the scatterplots: We can see here that yestavg and yesthi are good predictors for hitemp looking at the scatterplot and eventually using the correlation plot.

# 4:

```{r}
sapply(weather[,2:9], function(x) (c(mean(x), median(x), range(x))))
```


This shows that the mean on the first row, median on the second, and the lower and upper(range) with the respective 3rd and 4th row

# 5:

```{r}
names(weather)
```


```{r}
model1 <- lm(hitemp ~ yesthi, weather)
summary(model1)
```

This shows that the model is statisticall significant with the t statistic having a small P value and looking at the R squared this model explains about 39% of the variability


y =   28.7939 + 0.6211(yesthi)

# 6:

```{r}
model2 <- lm(hitemp ~ yesthi + yestlo, weather)
summary(model2)
```

This shows that the model is statistically significant with the t statistic having a small P value for yesthi but *not* yestlo with a p value of .37. This means that its not a good estimator for hitemp, and looking at the R squared this model explains about 40% of the variability. Looking at the F statistic its not too high but still having a small p value, it means the the model is statistically significant. 

The prediction equation : 30.79716 + 0.64926 (yesthi) + -0.07046 (yestlo) + epsilon


# 7:

Using model 1 we want to predict the hitemp 


y =   28.7939 + 0.6211(yesthi)

yesthi = 75

so the answer is  75.3764


```{r}
28.7939 + 0.6211*(75)
```

We chose this model over the multiple one since yestlo was not a significant individual predictor. 
















