---
title: "R for Data Science CH03--CH05"
author: "Jefferson Ong"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(comment = "")
```

***

Read [Chapter 3](https://r4ds.had.co.nz/data-visualisation.html), [Chapter 4](https://r4ds.had.co.nz/workflow-basics.html), and [Chapter 5](https://r4ds.had.co.nz/transform.html) of _R for Data Science_. Work through the activities below and complete this accompanying Markdown document. Knit to HTML and submit your work as instructed. 

```{r, message = FALSE}
# load all packages here
library(ggplot2)

library(dplyr)

library(nycflights13)
library(tidyverse)

```


### 3.5.1 Exercises

1) What happens if you facet on a continuous variable?

ANSWER: It will facet for each value. AS in a bunch fo different graphs


2) Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other options control the layout of the individual panels? Why doesn’t `facet_grid()` have `nrow` and `ncol` arguments? Read the documentation [here](https://ggplot2.tidyverse.org/reference/facet_wrap.html) and [here](https://ggplot2.tidyverse.org/reference/facet_grid.html).

ANSWER: nrow is how many rows of faceted graphs do you want, same with ncol. it treats individual cells as a graph in a larger matrix in some sense. Theres a bunch fo panel commands, scaling, facet.by on specificity. facet grid uses the variables to figure out the x and y, vertical ~ horizontal.


### 3.6.1 Exercises

6) Recreate the R code necessary to generate the following graphs (see textbook).

```{r}
# show your code here (plot1)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(aes(size = 2)) +
  geom_smooth(se = FALSE)
```

```{r}
# show your code here (plot2)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, fill = drv)) + 
  geom_point(aes( size = 2)) +
  geom_smooth(se = FALSE)
```


```{r}
# show your code here (plot3)
names(mpg)
str(mpg)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, fill = drv, color = drv)) + 
  geom_point(aes(size = 2)) +
  geom_smooth(se = FALSE)
```

```{r}
# show your code here (plot4)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(aes(color = drv, size = 2)) +
  geom_smooth(se = FALSE)
```

```{r}
# show your code here (plot5)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, group = drv)) + 
  geom_point(aes(color = drv, size = 2)) +
  geom_smooth(aes(linetype = drv), se = FALSE)
```

```{r}
# show your code here (plot6)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(aes(color = drv, size = 2))
```


### 3.9.1 Exercises

2) What does `labs()` do? Read the [documentation](https://ggplot2.tidyverse.org/reference/labs.html).

ANSWER: Puts abels on the graph with titles, captions, subtitles, x, y, etc. 


3) What’s the difference between `coord_quickmap()` and `coord_map()`? Read the [documentation](https://ggplot2.tidyverse.org/reference/coord_map.html).

ANSWER: coord map is the more precise version of it. quickmaps estimates it and works better for smaller maps closer to equator. Both project the earth on 2d.


4) What does the plot (see textbook) tell you about the relationship between city and highway mpg? Why is `coord_fixed()` important for this plot? What does `geom_abline()` do on this plot? Read the documentation [here](https://ggplot2.tidyverse.org/reference/coord_fixed.html) and [here](https://ggplot2.tidyverse.org/reference/geom_abline.html).

ANSWER: SHows a positive linear relationship. Coord fix changes the aspect ratio of the graph, it can be skewed to manipulate the relationship. geom abline shows the slope witht he y intercept, it would be the line in thsi case. 


### 5.2.4 Exercises

1) Find all flights that...

a) Had an arrival delay of two or more hours

```{r}
# show your code here
filter(flights, arr_delay >= 120)

```



b) Flew to Houston (IAH or HOU)

```{r}
# show your code here
filter(flights, dest %in% c('IAH', 'HOU'))
```


c) Were operated by United, American, or Delta

```{r}
# show your code here
joined_flight <- left_join(flights, airlines)

filter(joined_flight, name %in% c('United Air Lines Inc.','American Airlines Inc.','Delta Air Lines Inc.'))
```



d) Departed in summer (July, August, and September)

```{r}
# show your code here
filter(flights, month %in% c(7,8,9))
```


e) Arrived more than two hours late, but didn’t leave late

```{r}
# show your code here
filter(flights, arr_time > 120 & dep_delay < 0)
```


f) Were delayed by at least an hour, but made up over 30 minutes in flight

```{r}
# show your code here
filter(flights, dep_delay >= 60, sched_arr_time > -30)
```

g) Departed between midnight and 6am (inclusive)

```{r}
# show your code here
filter(flights, dep_time >= 2400 | dep_time <= 600)
```
```{r}
names(flights)
# summary(flights)
str(flights)
glimpse(flights)
```


3) How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

ANS: dep_delay = 8255, arr_time = 8713, arr_delay = 9430, dep_time = 8255. seems liek they represent the zeros and just counted as na. 

```{r}
# show your code here
sum(!complete.cases(flights$dep_time))
sum(!complete.cases(flights))
summary(flights)
```


### 5.3.1 Exercises

1) How could you use arrange() to sort all missing values to the start? (Hint: use `is.na()`).

```{r}
# show your code here
df <- tibble(x = c(1, NA, 3))
arrange(df, desc(is.na(x)))
```


2) Sort flights to find the most delayed flights. Find the flights that left earliest.

```{r}
# show your code here
arrange(flights, dep_delay)
arrange(flights, desc(dep_delay))
```


3) Sort flights to find the fastest flights.

```{r}
# show your code here
arrange(flights, air_time)
```


4) Which flights travelled the longest? Which travelled the shortest?

```{r}
# show your code here
flights %>%
  arrange(air_time) %>%
  select(carrier, flight, air_time)

flights %>%
  arrange(-air_time) %>%
  select(carrier, flight, air_time)
```


### 5.5.2 Exercises

1) Currently `dep_time` and `sched_dep_time` are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
# show your code here

mutate(flights,
       dep_time = (dep_time %/% 100) * 60 + (dep_time %% 100),
       sched_dep_time = (sched_dep_time %/% 100) * 60 + (sched_dep_time %% 100))
```


2) Compare `dep_time`, `sched_dep_time`, and `dep_delay`. How would you expect those three numbers to be related? Do they meet this expectation?

```{r}
# show your code here
flights_deptim <-
  mutate(flights,
    dep_time_min = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
    sched_dep_time_min = (sched_dep_time %/% 100 * 60 +
      sched_dep_time %% 100) %% 1440,
    dep_delay_diff = dep_delay - dep_time_min + sched_dep_time_min
  )
flights_deptim
```

ANSWER:


### 5.7.1 Exercises

2) Which plane (tailnum) has the worst on-time record?

```{r}
# show your code here
flights %>%
  group_by(tailnum) %>%
  summarise(arr_delay = mean(arr_delay)) %>%
  filter(min_rank(desc(arr_delay)) <= 1)
```

ANSWER: N844MH


3) What time of day should you fly if you want to avoid delays as much as possible?

```{r}
# show your code here
flights %>%
  group_by(hour) %>%
  summarise(arr_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(arr_delay)
```

ANSWER: Morning flights


***

```{r}
sessionInfo()
```
