---
title: "RDS CH09-CH11"
author: "Jefferson Ong"
date: "3/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```



# 11.2.2 #1

A: read_delim()



# 11.2.2 #2

```{r}
union(names(formals(read_csv)), names(formals(read_tsv)))
```
A: 
  `col_names` and `col_types` are used to specify the column names and how to parse the columns
  `locale` is important for determining things like the encoding and whether "." or "," is used as a decimal mark.
   `na` and `quoted_na` control which strings are treated as missing values when parsing vectors
   `trim_ws` trims whitespace before and after cells before parsing
   `n_max` sets how many rows to read
   `guess_max` sets how many rows to use when guessing the column type
   `progress` determines whether a progress bar is shown.
   
   
   
   

# 11.3.5 #5

A: read_cvs2() uses a semi-colon to separate data while read_csv() uses a comma.



# 11.3.5 #6


A: UTF-8 is standard
    
-   Western European Latin script languages: ISO-8859-1, Windows-1250 (also CP-1250 for code-point)
-   Eastern European Latin script languages: ISO-8859-2, Windows-1252
-   Greek: ISO-8859-7
-   Turkish: ISO-8859-9, Windows-1254
-   Hebrew: ISO-8859-8, IBM424, Windows 1255
-   Russian: Windows 1251
-   Japanese: Shift JIS, ISO-2022-JP, EUC-JP
-   Korean: ISO-2022-KR, EUC-KR
-   Chinese: GB18030, ISO-2022-CN (Simplified), Big5 (Traditional)
-   Arabic: ISO-8859-6, IBM420, Windows 1256


# 11.3.5 #7

```{r}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014
t1 <- "1705"
t2 <- "11:15:10.12 PM"

# ANSWER: 
parse_date(d1, "%B %d, %Y")
parse_date(d2, "%Y-%b-%d")
parse_date(d3, "%d-%b-%Y")
parse_date(d4, "%B %d (%Y)")
parse_date(d5, "%m/%d/%y")
parse_time(t1, "%H%M")
```

```{r}
parse_time(t2, "%H:%M:%OS %p")
```



# EXTRA: What is an RDS file?

A: R built in format. It allows storing in working environment, it can also be used as extensions.

# EXTRA: What is JSON?


A: JavaScript Object Notation. A minimal, readable format for structuring data. It is used primarily to transmit data between a server and web application. 