---
title: "Winterization Activity"
author: "Jefferson Ong"
date: "1/29/2019"
output: html_document
---


## Dataset Structure

A dataset is a collection of data values for the observations and variables in a study (e.g., survey or experiment). 

-An observational unit (or observation) is a single individual, object, or entity about which we have collected data for one or underlying attributes (variables).

## `tidy` Data

* Each column is a variable

* Each row is an observation

* Each type of observational unit forms a "table"

* Untidy data is sometimes called wide (versus long) format


## Types of Variables

* categorical / qualitative

These are nominal or ordinal.

* numerical / quantitative

These are interval or ratio. ALso classified as discrete ro continuous. 



## Reading Data into R


The `read.csv()` function converts a `.csv `file into a data frame. The `read_csv()` function in the` readr `package coverts a `.csv` file into an enhanced data frame known as a `tibble`.

```{r}
library(readr)
draft <- read_csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/vietnamdraft.csv"))

draft_wide <- read_csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/vietnamdraft_wide.csv"))
```

How to specify in dataframes

dataset_name$column_name

list_name$list_component

## Some Useful Functions

str()

glimpse() # from dplyr

attributes()

names()

class()

dim()

nrow()

ncol()

head()

tail()

summary()

## Missing Values

```{r}
sum(complete.cases(draft)) # this is two nested functions
sum(!complete.cases(draft)) # what does the ! do?


```


```{r}
index <- which(is.na(draft$draftnumber) == TRUE)
draft[index,]
draft[index, 1:3]
draft[index, "day"]
draft[index, c("month", "day", "draftyear")]
```

```{r}
library(tidyr)
draft_wide_gathered <- draft_wide %>% 
  gather(key = "draftyear", value = "draftnumber", num1970:num1973)


draft_wide_gathered
```

```{r}
library(tidyr)
draft_spread <- draft %>% 
  spread(key = "draftyear", value = "draftnumber")

draft_spread
```

```{r, warning=FALSE, echo=FALSE}
library(gapminder)
library(tidyverse)

(my_gap <- gapminder)
```


```{r}
## let output print to screen, but do not store
my_gap %>% filter(country == "Canada")

## store the output as an R object
my_precious <- my_gap %>% filter(country == "Canada")
```

```{r}
my_gap %>%
  mutate(gdp = pop * gdpPercap)
```


```{r}
ctib <- my_gap %>%
  filter(country == "Canada")
## this is a semi-dangerous way to add this variable
## I'd prefer to join on year, but we haven't covered joins yet
my_gap <- my_gap %>%
  mutate(tmp = rep(ctib$gdpPercap, nlevels(country)),
         gdpPercapRel = gdpPercap / tmp,
         tmp = NULL)

my_gap
```

```{r}
my_gap %>% 
  filter(country == "Canada") %>% 
  select(country, year, gdpPercapRel)
```

```{r}
summary(my_gap$gdpPercapRel)
```


```{r}
my_gap %>%
  arrange(year, country)
```

```{r}
my_gap %>%
  filter(year == 2007) %>%
  arrange(lifeExp)
```

```{r}
my_gap %>%
  filter(year == 2007) %>%
  arrange(desc(lifeExp))
```



```{r}
my_gap %>%
  rename(life_exp = lifeExp,
         gdp_percap = gdpPercap,
         gdp_percap_rel = gdpPercapRel)
```


```{r}
my_gap %>%
  filter(country == "Burundi", year > 1996) %>% 
  select(yr = year, lifeExp, gdpPercap) %>% 
  select(gdpPercap, everything())
```


## dplyr 

* `group_by()` adds extra structure to your dataset – grouping information – which lays the groundwork for computations within the groups.

* `summarize()` takes a dataset with n observations, computes requested summaries, and returns a dataset with 1 observation.

* Window functions take a dataset with n observations and return a dataset with n observations.

* `mutate()` and `summarize()` will honor groups.

* You can also do very general computations on your groups with `do()`, though elsewhere in this course, I advocate for other approaches that I find more intuitive, using the purrr package.



```{r}
my_gap %>%
  group_by(continent) %>%
  summarize(n = n())
```


```{r}
table(gapminder$continent)
```

```{r}
str(table(gapminder$continent))
```

```{r}
str(table(gapminder$continent))
```


```{r}
my_gap %>% 
  count(continent)
```

```{r}
my_gap %>%
  group_by(continent) %>%
  summarize(n = n(),
            n_countries = n_distinct(country))
```


## General summarization

The functions you’ll apply within `summarize()` include classical statistical summaries, like `mean()`, `median()`,  `var()`, `sd()`, `mad()`, `IQR()`, `min()`, and `max()`. 

```{r}
my_gap %>%
  group_by(continent) %>%
  summarize(avg_lifeExp = mean(lifeExp))
```


```{r}
my_gap %>%
  filter(year %in% c(1952, 2007)) %>%
  group_by(continent, year) %>%
  summarize_at(vars(lifeExp, gdpPercap), funs(mean, median))
```

```{r}
my_gap %>%
  filter(continent == "Asia") %>%
  group_by(year) %>%
  summarize(min_lifeExp = min(lifeExp), max_lifeExp = max(lifeExp))
```

```{r}
my_gap %>% 
  group_by(country) %>% 
  select(country, year, lifeExp) %>% 
  mutate(lifeExp_gain = lifeExp - first(lifeExp)) %>% 
  filter(year < 1963)
```

```{r}
my_gap %>%
  filter(continent == "Asia") %>%
  select(year, country, lifeExp) %>%
  group_by(year) %>%
  filter(min_rank(desc(lifeExp)) < 2 | min_rank(lifeExp) < 2) %>% 
  arrange(year) %>%
  print(n = Inf)
```


```{r}
asia <- my_gap %>%
  filter(continent == "Asia") %>%
  select(year, country, lifeExp) %>%
  group_by(year)
asia
```


```{r}
asia %>%
  mutate(le_rank = min_rank(lifeExp),
         le_desc_rank = min_rank(desc(lifeExp))) %>% 
  filter(country %in% c("Afghanistan", "Japan", "Thailand"), year > 1995)
```

```{r}

# filter(min_rank(desc(lifeExp)) < 2 | min_rank(lifeExp) < 2)

my_gap %>%
  filter(continent == "Asia") %>%
  select(year, country, lifeExp) %>%
  arrange(year) %>%
  group_by(year) %>%
  #top_n(1, wt = lifeExp)        ## gets the min
  top_n(1, wt = desc(lifeExp)) ## gets the max
```


```{r}
my_gap %>%
  select(country, year, continent, lifeExp) %>%
  group_by(continent, country) %>%
  ## within country, take (lifeExp in year i) - (lifeExp in year i - 1)
  ## positive means lifeExp went up, negative means it went down
  mutate(le_delta = lifeExp - lag(lifeExp)) %>% 
  ## within country, retain the worst lifeExp change = smallest or most negative
  summarize(worst_le_delta = min(le_delta, na.rm = TRUE)) %>% 
  ## within continent, retain the row with the lowest worst_le_delta
  top_n(-1, wt = worst_le_delta) %>% 
  arrange(worst_le_delta)
```





























